### 2013-09-27 papers/2012/tags/nss2012final/NSS.tex
add a 'papers' directory with 2012 NSS paper, and CHEP start

\documentclass[journal]{IEEEtran}
\usepackage{graphicx}
\usepackage{subfigure}

\usepackage{hyperref}
\usepackage{multirow}
\usepackage[all]{hypcap} % fix links to figs/tabs

\renewcommand{\arraystretch}{1.1} % make tables more readable

% disable the running header on all pages following the title page
\pagestyle{empty}
%
\begin{document}
% paper title
\title{Applications of GPUs to Online Track Reconstruction in HEP Experiments}
% Author list
\author{S.~Amerio, R.~Ammendola, A.~Biagioni, D.~Bastieri, D.~Benjamin, 
O.~Frezza, S.~Gelain, W.~Ketchum, Y.~K.~Kim, F.~Lo~Cicero, A.~Lonardo, T.~Liu, D.~Lucchesi, P.~S.~Paolucci, S.~Poprocki, D.~Rossetti, F.~Simula, L.~Tosoratto, 
G.~Urso, P.~Vicini, and P.~Wittich
\thanks{Manuscript received November 16, 2012.
This work was supported by the National Science Foundation, the
Department of Energy Office of Science and INFN.}% <-this % stops a space
\thanks{W.~Ketchum and Y.~K.~Kim are with University of Chicago.}%
\thanks{T.~Liu is with Fermi National Accelerator Laboratory.}%
\thanks{S.~Poprocki and P.~Wittich are with Cornell University.}%
\thanks{S.~Amerio is with INFN Padova.}% 
\thanks{R.~Ammendola is with INFN Tor Vergata.}%
\thanks{A.~Biagion, O.~Frezza,  F.~ Lo Cicero, A.~Lonardo, P.~S.~Paolucci, D.~Rossetti,  F.~Simula, L.~Tosoratto and P.~Vicini are with INFN Roma}% 
\thanks{D.~Bastieri, S.~Gelain and D.~Lucchesi are with University of Padova and INFN.}%
\thanks{G.~Urso is with ORMA Software.}%
\thanks{D.~Benjamin is with Duke University.}%
}
% make the title area
\maketitle
% after \maketitle, add the following command to remove the header on the title page
\thispagestyle{empty}
% abstract usually in passive voice
\begin{abstract}
  One of the most important issues that particle physics experiments
  at hadron colliders have to solve is real-time selection of the most
  interesting events. Typical collision frequencies do not allow all
  events to be written to tape for offline analysis, and in most
  cases, only a small fraction can be saved. The most commonly used
  strategy is based on two or three selection levels, with the low
  level ones usually exploiting dedicated hardware to decide within a
  few to ten microseconds if the event should be kept or not. This
  strict time requirement has made the usage of commercial devices 
  inadequate, but recent improvements to Graphics Processing Units (GPUs)
  have substantially
  changed the conditions. Thanks to their highly parallel,
  multi-threaded, multicore architecture with remarkable computational
  power and high memory bandwidth, these commercial devices may be
  used in scientific applications, among which the 
  event selection system (trigger) in particular may benefit, 
  even at low levels. This paper describes the results of an R\&D
  project to study the performance of GPU technology for
  low latency applications, such as HEP fast tracking trigger
  algorithms.  On two different setups, we measure the latency to
  transfer data to/from the GPU, exploring the timing
  of different I/O technologies on different GPU models. We then
  describe the implementation and the performance of a track fitting
  algorithm which mimics the CDF Silicon Vertex Tracker.  These
  studies provide performance benchmarks to investigate the potential
  and limitations of GPUs for future real-time applications in HEP
  experiments.
\end{abstract}

\section{Introduction}
\IEEEPARstart{R}{eal} time event reconstruction plays a fundamental
role in High Energy Physics (HEP) experiments at hadron colliders.
Reducing the rate of data to be saved on tape from millions to
hundreds of events per second is critical. To increase the purity of
the collected samples, rate reduction has to be coupled with an initial
selection of the most interesting events.  In a typical hadron collider experiment, the
event rate has to be reduced from tens of MHz to a few kHz.  The selection system
(trigger) is usually organized in multiple levels, each capable of
performing a finer selection on more complex physics objects
describing the event. Trigger 
systems usually comprise a first level based on custom hardware,
followed by one or two levels usually based on farms of general
purpose processors.  The possibility of also using commercial devices
at a low trigger level is very appealing: they are subject
to continuous performance improvements driven by the consumer market, are
less expensive than dedicated hardware, and are easier to support.  Among
the commercial devices, Graphic Processing Units (GPUs) are of
particular interest for online selections given their impressive
computing power (the latest NVIDIA~\cite{bib_nvidia} GPU architecture,
Kepler, exceeds Teraflop computing power); moreover, high-level
programming architectures based on C/C++ such as
\textsc{cuda}~\cite{bib_cuda} and \textsc{opencl}~\cite{bib_OpenCL}
make these devices more accessible to the general user.  However, GPUs
are not designed for low-latency response, which is a critical
requirement in a trigger system.

We showed initial studies on GPU performance in low-latency
environments ($\sim$100\,$\mu$s) in a previous
paper~\cite{bib_GPU_TIPP2011}.  In this paper we complement those
studies measuring the timing performance of different GPU cards in
different data I/O environments. The algorithm run on the GPU is a
simplified version of the fast track fitting algorithm of the Silicon
Vertex Tracker (SVT) system at CDF~\cite{bib_SVT1}. It has been implemented
and tested on two commercial GPUs (NVIDIA GTX 285 and GTX 590) and on an
NVIDIA Tesla M2075, optimized for scientific and technical
computing. To transfer data to the GPU, we have compared the
performance of S-LINK~\cite{bib_slink} --- 
used in the second level of the CDF trigger system and which we have prior experience with --- and InfiniBand (IB), a more modern protocol.

The goal of this study is to investigate the strengths and weaknesses of
GPUs when applied in a low latency environment, with particular
emphasis on the data transfer latency to/from the GPU and the
algorithm latency for processing on the GPU in a manner similar to
a typical HEP trigger application.

\section{Experimental setup and data flow}
The studies are performed on two different experimental setups,
equipped with different data transfer links and GPU cards.  In both
setups two PCs are used, one acting as a transmitter (TX) and the other 
as a receiver (RX).  Data are transferred from TX to RX, processed on the GPU, and sent back to the receiver (see
Fig.~\ref{fig_data_flow}).  The latency for a complete loop is measured
on the transmitter using the time stamp counter register.

In this setup, the transmitter can represent the detector, as
the source of the data, or an upstream trigger processor, as
the ultimate sink of the data, while the the receiver is the
trigger system: the time to transfer data to the receiver is thus a
rough estimate of the latency to transfer the data from the detector
front-end to the trigger system.

The input data is a set of 32-bit words, each representing a set of
hits from the detector. When the track fitting algorithm is run, the
output data is composed of 32-bit words as well, representing the
results of the fit; when the input data is not processed by the fitting algorithm, the output
data is a copy of the input words.

\begin{figure}[!h]
\centering
\includegraphics[width=3.5in]{figures/SetUp-general}
\caption{Data flow. Data is sent from the transmitter PC to the
  receiver PC, where it is processed by the GPU before being returned
  to the transmitter PC. The transmitter plays the role of the
  detector as the source of the data and as an upstream trigger
  processor as the data's ultimate sink. The receiver PC plays the
  role of a component in the trigger system. We consider different I/O
  protocols (S-LINK and InfiniBand) and both consumer (GTX) and scientific computing (Tesla) GPUs. }
\label{fig_data_flow}
\end{figure}

\subsection{S-LINK and GTX cards setup}

Both TX and RX PCs are equipped with a CERN FILAR (Four Input Links
for ATLAS Readout)~\cite{bib_filar} and an S32PCI64~\cite{bib_solar}
SOLAR (Single Output Link for ATLAS Readout) PCI interface card for
I/O communication over S-LINK~\cite{bib_slink}.  The TX PC houses two
2.4 GHz AMD Opteron 250 (single core) processors while the RX PC
contains a 3.07 GHz Intel Core i7 CPU 950 (quad core with
hyperthreading) processor.  The RX PC also contains either an NVIDIA GeForce GTX 285 or GTX 590 GPU in a PCIe slot for the GPU computations. A separate GPU handles display operations.

%SILVIA: I propose to remove the following lines as this is repeated 
%before the Data Transfer paragraph.
%The time stamp counter register on the TX PC is used for timing measurements 
%since it is consistent with but has a higher resolution than the high 
%precision Linux timers on this system. Long tails in timing 
%distributions were removed by locking the program to a single CPU core 
%and moving all system interrupts to the remaining CPUs, and running 
%at the highest system scheduler priority, on both PCs. 
%

\subsection{InfiniBand and Tesla cards setup}
The second setup is a computing cluster composed of 12 identical nodes. 
Each node contains a Intel Xeon E6520 2.4 GHz CPU and two Tesla M2075 
GPU cards. The nodes are connected by InfiniBand communication 
links using Connect-X2 Mellanox or Apenet+ adapters. 
Apenet+ is an FPGA-based PCIe board supporting peer-to-peer communication with 
Tesla and Kepler cards~\cite{bib_Apenet}.

\begin{table}[!t]
  \centering
  \begin{tabular}{|l|c|c|c|}
    \hline
    & GTX 285 & GTX 590 & Tesla M2075 \\
    \hline
    \hline
    %Multiprocessors & & & 14 \\
    CUDA cores & 240 & 1024 & 448  \\
    %Peak Single Precision floating point performance (GFlops) & & & 1030 \\
    Memory interface & GDDR3 & GDDR5 & GDDR5 \\
    Memory size (GB) & 1 & 3 & 6  \\
    Memory bandwidth  (GB/s) & 159 & 327 & 150 \\   
    GPUDirect support & No & No & yes \\
    \hline
  \end{tabular}
  \caption{Capabilities of the GPUs used in this study, according to the manufacturer's specifications.}
  \label{tab_GPUcards}
\end{table}

\section{Latency measurements}
The data flow can be divided into separate steps, each adding its own 
contribution to the total latency. First, data are transferred from the 
transmitter to the receiver. Once on the receiver, data
are copied to the GPU and then are processed by the GPU. 
%Finally, the output  from the GPU is typically copied back to the receiver PC, and then sent back to the transmitter.

We perform three different sets or measurements:
\begin{enumerate}
\item[A)] Latency for data transfer only:\\
TX $\to$ RX system memory $\to$ TX;
\item[B)] Latency for data transfer and copy to/from the GPU:\\
TX $\to$ RX system memory $\to$ GPU memory $\to$ RX system memory $\to$ TX;
\item[C)] Latency for data transfer, copy to/from the GPU and processing  
in the GPU, where the data flow is the same as above, except a benchmark algorithm is performed in the GPU.
\end{enumerate}

The latency of each round-trip is measured on the transmitter using
the \textsc{rdtsc} time stamp function 
which was found to be consistent with --- but having a higher resolution than --- the high precision Linux timers on our system. 
Each test is repeated 10,000
times and we show the mean and RMS of the repeated measurements. 
To minimize interrupts we use two strategies first developed during the
CDF L2 studies. We use \textsc{cpu affinity} to tie our code to a
single processor, change the scheduler priority of the process to the
highest value (99) and set the scheduling algorithm to
\textsc{sched\_fifo} to ensure that our algorithm does not get removed
from the CPU queue. We also use \textsc{irq affinity} to mask off the
CPU selected above from all interrupts. Fig.~\ref{fig:affinity} shows that 
this setup reduces the long tails in the timing distributions.
\begin{figure}[tbp]
\centering
\includegraphics[width=3.5in]{figures/affinity_32_20121017_irqaffinity}
\caption{Comparison of the timing distributions for default running (solid black)
and for applying CPU \& IRQ affinities and scheduler priority (dashed
red). We see a reduction in the long tails, though some remain. }
\label{fig:affinity}
\end{figure}
However, the timing distributions typically exhibit features such as
multiple humps and long non-Gaussian tails. Preliminarily, we suspect
these features are due to interactions between the CPU and GPU, or
possibly power-saving modes on the GPUs, and further studies are
forthcoming.

\subsection{Data transfer}
Packets of 32-bit words, from a few bytes to 2 MB, are sent from the receiver
to the transmitter and back. No GPU is involved in this measurement.
Due to hardware limitations, the maximum size in input to the S-LINK 
setup is 4 kB.

The latency as a function of the packet size is shown in 
Fig.~\ref{fig_DT_SLINK_IB} and in Tab.~\ref{tab_DT}: 
IB shows a linear increase in latency for 
packets up to 2 MB, where 1.4 ms are needed for a complete loop.
A comparison with S-LINK for packets up to 2 kB is presented in the 
inset of the same figure: IB shows lower latencies and a 
smaller slope with increasing packet size. About 13 $\mu$s are needed to 
transfer (round-trip) 4 kB of data.

\begin{table}[!t]
  \centering
  \begin{tabular}{|r||r|r|}
    \hline

    Input  & \multicolumn{2}{|c|}{Data transfer latency ($\mu$s)}  \\

    (Bytes) & \multicolumn{1}{|c|}{S-LINK}   & \multicolumn{1}{|c|}{IB}\\

    \hline
    \hline
    8    & 12.1 $\pm$ 0.4 & 3.3    $\pm$ 0.3 \\
    %\hline
    32   & 12.6 $\pm$ 0.3 & 3.4    $\pm$ 0.3 \\
    %\hline
    128  & 13.2 $\pm$ 0.4 & 6.1    $\pm$ 0.3 \\
    %\hline
    512  & 16.6 $\pm$ 0.4 & 7.0    $\pm$ 0.2 \\
    %\hline
    1024 & 20.8 $\pm$ 0.2 & 8.3    $\pm$ 0.2 \\
    %\hline
    4096 & 40.3 $\pm$ 0.3 & 13.4   $\pm$ 0.3 \\
    %\hline
    32k  &                & 34.7   $\pm$ 1.8 \\
    %\hline
    65k  &                & 56.0   $\pm$ 1.8 \\
    %\hline
    204k &                & 148.3  $\pm$ 1.1\\
    %\hline
    512k &                & 350.4  $\pm$ 1.8\\
    %\hline
    2M   &                & 1358.1 $\pm$ 1.8\\

    \hline
  \end{tabular}
\caption{Data transfer latency. This time measures the round trip 
from the transmitter to the receiver, without interaction to the GPU. }
\label{tab_DT}
\end{table}

\begin{figure}[!t]
\centering
\includegraphics[width=3.5in]{figures/DT_SLINK_IB_inset}
% where an .eps filename suffix will be assumed under latex, 
% and a .pdf suffix will be assumed for pdflatex; or what has been declared
% via \DeclareGraphicsExtensions.
\caption{Data transfer latency as a function of input packets from 32 kB to 
2 MB and for packets from 1 B to 4 kB (inset, logarithmic scale).}
\label{fig_DT_SLINK_IB}
\end{figure}

%\begin{figure}[!h]
%\centering
%\includegraphics[width=3.5in]{figures/IB_DT_high}
%\caption{Data transfer latency for input packets from 32 kB to 2 MB.}
%\label{fig_DT_IB_HIGH}
%\end{figure}
%
\subsection{Data copy to/from the GPU}

The two experimental setups allow us to test different data transfer 
strategies to the GPU. The standard data transfer strategy is via the 
system memory, where the 
PCIe adapter card and the GPU allocate \emph{separate} buffers on the
system memory for the copy (as shown in Fig.~\ref{fig_standardDT}).
This is inefficient, as the data are copied twice in the 
system memory before being transferred to the GPU/PCIe card. 
Data may also be transferred using Direct Memory Access (DMA, 
GPUDirect\cite{bib_GPUDirect}) to the CPU memory:
the PCIe card and the GPU share the \textit{same} buffer on the CPU memory; as 
a result the data are copied only once in the CPU memory 
(Fig.~\ref{fig_GPUDirectV1}). 
GPUDirect is the standard data transfer mechanism used in the IB-Tesla setup. 
While the GTX GPUs cannot use GPUDirect functionalities, the firmware for S-LINK 
PCI cards allows us to specify an address in system memory where the data is deposited, 
thus achieving a GPUDirect-like setup that avoids an additional copy of the data in system memory.
In the IB-Tesla setup we can test two additional copy 
strategies, which are the results of different levels of 
optimization of the GPUDirect protocol:
\begin{itemize}
\item GPU-Aware MPI, where the copy latency is further reduced
by automatically allocating the buffer on the CPU memory;
\item peer-to-peer (P2P) strategy, when data are transferred 
directly to the GPU, without any  
intermediate copy to the CPU (Fig.~\ref{fig_GPUDirectV2}).
\end{itemize}
%
\begin{figure*}[!t]
\centering
\subfigure[]
{\label{fig_standardDT}
\includegraphics[width=2in]{figures/noGPUDirect}}
\hspace{1mm}
\subfigure[]
{\label{fig_GPUDirectV1}
\includegraphics[width=2in]{figures/GPUDirect}}
\subfigure[]
{\label{fig_GPUDirectV2}
\includegraphics[width=2in]{figures/GPUDirect-APE}}
\caption{Standard data transfer (a), via GPUDirect (b) and via
  GPUDirect with P2P support (c). In (a), two buffers are required in
  the main memory. In GPUDirect (b), one of the main memory buffers is
  eliminated. In GPUDirect with P2P support, data is sent directly
  from the APEnet+ transceiver to the GPU memory.}
\end{figure*}
% No spurious paragraph breaks
As in the previous measurements, packets of varying size are sent to
the receiver, copied from the system memory to the GPU and back, and
finally sent back to the transmitter.  The total round-trip time
(Tab.~\ref{tab_MC}) comprises the data transfer plus copy latency.
Measurements done with the P2P protocol are available only for packets
in the range 32 B -- 4 kB.

For small packet sizes (see Fig.~\ref{fig_DT_MC_SLINK_IB_APE_low}), the 
S-LINK+GTX and IB+Tesla setups show similar performance, about 
30 $\mu$s to transfer 100~B of data. For over 100~B of transferred data,  
the S-LINK latency increases with a much larger slope than IB.
P2P is the best strategy in terms of latency over the whole considered range 
(32 B -- 4 kB).
A comparison between standard GPUDirect and GPU-aware MPI strategy is presented
in Fig.~\ref{fig_DT_MC_GPUDirect_high}. We find that the GPU-Aware MPI 
strategy gives consistent reduction only above 200 kB.

\begin{table}[!t]
  \centering
  \begin{tabular}{|r||r|r|r|r|}
    \hline
    \multicolumn{1}{|c||}{\multirow{2}{*}{Input}} & \multicolumn{4}{|c|}{Data transfer + copy latency  ($\mu$s)}  \\\cline{2-5}
    \multirow{2}{*}{(Bytes)}
& \multicolumn{2}{|c|}{GPUDirect}
%& \multicolumn{1}{|c|}{GPUDirect}
& \multicolumn{1}{|c|}{GPU-Aware MPI}
& \multicolumn{1}{|c|}{P2P} \\
& \multicolumn{1}{|c|}{(S-LINK)}
& \multicolumn{1}{|c|}{(IB)}
& \multicolumn{1}{|c|}{(IB)}
& \multicolumn{1}{|c|}{(Apenet+)} \\

    \hline
    \hline
     8    & 29.0 $\pm$ 0.3 & 27.1 $\pm$ 2.0   & 27.3 $\pm$ 0.9 &  \\
     %\hline
     32   & 29.3 $\pm$ 0.1 & 27.5 $\pm$ 0.7   & 27.5 $\pm$ 2.4 &  18.1 $\pm$ 1.2 \\
     %\hline                               
     128  & 32.6 $\pm$ 1.0 & 29.6 $\pm$ 0.7   & 28.9 $\pm$ 1.9 &  20.6 $\pm$ 3.0\\
     %\hline                               
     512  & 38.6 $\pm$ 0.9 & 30.8 $\pm$ 0.7   & 31.5 $\pm$ 0.8 &  20.8 $\pm$ 2.4\\
     %\hline                               
     1024 & 46.1 $\pm$ 1.0 & 32.6 $\pm$ 1.7   & 33.5 $\pm$ 2.3 &  21.2 $\pm$ 2.6\\
     %\hline                               
     4096 & 86.3 $\pm$ 0.7 & 39.1 $\pm$ 2.2   & 40.5 $\pm$ 1.1  &  25.6 $\pm$ 3.0\\
     %\hline                               
     32k  &                & 66.4 $\pm$ 2.3   & 84.9 $\pm$ 2.7 &\\
     %\hline                               
     65k  &                & 98.9 $\pm$ 2.1   &  115.4 $\pm$ 2.8 &\\
    %\hline
     204k &                & 230.3 $\pm$ 2.4  &  231.0 $\pm$ 4.1 &\\
     %\hline
     512k &                & 511.6 $\pm$ 3.3  &  413.0 $\pm$ 3.5 &\\
     %\hline
     2M   &                & 1917.8 $\pm$ 6.2 &  1330.5 $\pm$ 4.5 &\\

    \hline
  \end{tabular}
\caption{Latency for data transfer plus copy to/from the GPU. The S-LINK measurements here are taken using the GTX 590 GPU, while the IB and Apenet+ measurements use the Tesla M2075 GPU.}
\label{tab_MC}
\end{table}

\begin{figure}[!t]
\centering
\includegraphics[width=3.5in]{figures/DT_MC_SLINK_IB_APE_low}
\caption{Data transfer plus copy latency for GPUDirect standard, 
GPU-Aware MPI and P2P. Input packet size from 1 B (32 B for Apenet+) to 4 kB.}
\label{fig_DT_MC_SLINK_IB_APE_low}
\end{figure}

\begin{figure}[!t]
\centering
\includegraphics[width=3.5in]{figures/MC_IB_high}
\caption{Data transfer plus copy latency for GPUDirect standard 
 and GPU-Aware MPI up to 2 MB.}
\label{fig_DT_MC_GPUDirect_high}
\end{figure}

\subsection{Data processing on the GPU}

Many tasks performed by trigger systems may benefit from the 
parallelization available in a GPU (e.g. jet clustering and track finding). 
For this study, we choose as a benchmark a simplified fast track-fitting
algorithm which was used in CDF's Silicon Vertex Trigger 
(SVT)~\cite{bib_SVT1}~\cite{bib_SVT2}.
This algorithm uses a linearized approximation to track-fitting as 
implemented in hardware (described in greater detail in~\cite{bib_SVT3}). 
With the SVT approach, the determination of the track parameters 
($p_i$) is reduced to a simple scalar product:
\[
p_i = \vec{f_i} \cdot \vec{x_i} + q_i,
\]
where $\vec{x_i}$ are input silicon hits, and $\vec{f_i}$ and $q_i$ are 
pre-defined constant sets. For each set of hits, the algorithm
computes the impact parameter $d_0$, the azimuthal angle $\phi$, 
the transverse momentum $p_\mathrm{T}$ , and the $\chi^2$ of the
fitted track by using simple operations such as memory lookup and 
integer addition and multiplication. In
our testing of the track-fitting algorithm, each word in input is treated to
represent a set of silicon hits. While the track-fitting algorithm 
itself is simple, it must be performed on
many combinations of hits in a given event, especially in high-occupancy 
environments. This makes it a good 
benchmark for testing performance of a GPU, using massive parallelization at low latencies.

In Figs.~\ref{fig_DT_MC_TF_SLINK} and \ref{fig_DT_MC_TF_IB_low}
and Tabs.~\ref{tab_TF_SLINK} and \ref{tab_TF_IB},
the total latency for data transfer, 
copy to the GPU, and processing in the GPU is shown. For comparison, 
 the distributions for data transfer and data transfer plus copy are 
superimposed.

\begin{table}[!t]
  \centering
  \begin{tabular}{|r||r|r|r|}
    \hline

    \multirow{2}{*}{Input (Bytes)} & \multicolumn{3}{|c|}{ Latency  ($\mu$s)}  \\

                                   & \multicolumn{1}{|c|}{Data transfer}
                                   & \multicolumn{1}{|c|}{+ copy}
                                   & \multicolumn{1}{|c|}{+ calculations} \\

    \hline
    \hline
    128 & 13.2 $\pm$ 0.4 & 32.6 $\pm$ 1.0 & 39.5 $\pm$ 1.4 \\ 
    256 & 14.4 $\pm$ 0.4 & 33.9 $\pm$ 0.8 & 41.6 $\pm$ 1.5 \\ 
    512 & 16.6 $\pm$ 0.4 & 38.6 $\pm$ 0.9 & 46.8 $\pm$ 1.4 \\ 
    1024 & 20.9 $\pm$ 0.2 & 46.1 $\pm$ 1.0 & 53.9 $\pm$ 1.4 \\ 
    2048 & 27.7 $\pm$ 0.3 & 59.4 $\pm$ 0.8 & 67.1 $\pm$ 1.2 \\ 
    4096 & 40.3 $\pm$ 0.2 & 86.3 $\pm$ 0.7 & 93.9 $\pm$ 1.0 \\      
    \hline
  \end{tabular}
\caption{Data transfer + copy + calculation latency (S-LINK + GTX 590).}
\label{tab_TF_SLINK}
\end{table}

\begin{table}[!t]
  \centering
  \begin{tabular}{|r||r|r|r|}
    \hline

    \multirow{2}{*}{Input (Bytes)} & \multicolumn{3}{|c|}{ Latency  ($\mu$s)}  \\

                                   & \multicolumn{1}{|c|}{Data transfer}
                                   & \multicolumn{1}{|c|}{+ copy}
                                   & \multicolumn{1}{|c|}{+ calculations} \\

    \hline
    \hline
     8    & 3.3 $\pm$ 0.3   & 27.1 $\pm$ 2.0 & 33.3 $\pm$ 2.1\\
     %\hline
     32   & 3.4 $\pm$ 0.3   & 27.5 $\pm$ 0.7 & 33.2 $\pm$ 2.0\\
     %\hline
     128  & 6.1 $\pm$ 0.3  & 29.6 $\pm$ 0.7 & 35.7 $\pm$ 2.0\\
     %\hline
     512  & 7.0 $\pm$ 0.2  & 30.8 $\pm$ 0.7 & 36.5 $\pm$ 2.6\\
     %\hline
     1024 & 8.3 $\pm$ 0.2  & 32.6 $\pm$ 1.7 & 37.9 $\pm$ 2.7\\
     %\hline
     4096 &  13.4 $\pm$ 0.3 &  39.1 $\pm$ 2.2 & 45.8 $\pm$ 2.1\\
%     %\hline
%     32k  &  34.7 $\pm$ 1.8 &   66.4 $\pm$ 2.3 & 75.4 $\pm$ 3.9\\
%     %\hline
%     262k & 187.6 $\pm$ 1.1 & 280.2 $\pm$ 2.5 & 298.1 $\pm$ 2.7\\
%     %\hline
%     512k & 350.4 $\pm$ 1.8 & 511.7 $\pm$ 3.3 &   534.2 $\pm$ 4.4\\
%     %\hline
%     2M   &  1358.1 $\pm$ 1.8 & 1917.8 $\pm$ 6.2 & 2001.8 $\pm$ 6.8\\
%     
    \hline
  \end{tabular}
\caption{Data transfer + copy + calculation latency (IB + Tesla M2075).}
\label{tab_TF_IB}
\end{table}

\begin{figure}[!t]
  \centering
  \subfigure[]
{\label{fig_DT_MC_TF_SLINK}
\includegraphics[width=3.5in]{figures/DT_MC_TF_SLINK}}
\hspace{1mm}
\subfigure[]
{\label{fig_DT_MC_TF_IB_low}
 \includegraphics[width=3.5in]{figures/DT_MC_TF_IB_low}}
\caption{Total latency for data tranfer, copy to/from the GPU and data 
processing on the GPU, for S-LINK and GTX 590 card (a) and for IB and Tesla M2075 card (b). Note the $y$ axis goes up to 90 $\mu$s in (a) and 50 $\mu$s in (b).} 
\end{figure}

%\begin{figure}[!h]
%  \centering
%  \includegraphics[width=3.5in]{figures/DT_MC_TF_IB_low}
%  \caption{Total latency for data tranfer, copy to/from the GPU and data 
%processing on the GPU, for IB and Tesla card.}
%  \label{fig_DT_MC_TF_IB_low}
%\end{figure}
% 
Adding the track fitting algorithm on the GPU, the latency increases of
about 8~$\mu$s for the GTX 590 card and 6 $\mu$s for the Tesla. 
The two cards have similar performance in terms of calculation latency. 
The increase in the total latency is fixed as a function of 
the input data (\textit{i.e.}, the number of calculations since each input 
word is a track to be fitted), due to the parallel nature of the 
calculations inside the GPU.

This is visible also in Fig.~\ref{fig_TF_SLINK_IB_fixedSize}, where the total 
latency is measured for fixed input (2 kB) and output (256 kB) size, and 
only the number of fits is varied, from 32 to 16,384. The latency 
distribution is flat up to 1,000 fits and shows a slight slope 
for larger numbers of fits. The slope is largest for the GTX 285, as 
expected, due to it having the fewest number of CUDA cores.
In this plot the latency is dominated by the time needed to transfer 
the output back to the transmitter.
Moreover, each point is the mean 
over 10 different runs, and the error is the relative error on the mean: 
the smaller this error, the smaller the run-by-run spread --- a highly 
desirable feature in any real-time event selection system. In both experimental
setups we find that the run-by-run spread is in the range of 1.5 -- 3 $\mu$s.

\begin{figure}[!t]
  \centering
  \includegraphics[width=3.5in]{figures/TF_SLINK_IB_fixedSize}
  \caption{Total latency (data transfer, copy and track fitting algorithm) for the three different GPU cards used in this study.}
  \label{fig_TF_SLINK_IB_fixedSize}
\end{figure}

Figure~\ref{fig_TF_SLINK_IB_fixedSize_CPUvsGPU}
shows the latency for the same SVT-like track fitting
algorithm run on the CPU of the experimental setup, compared to the
performance on the GPU.  We stress that the algorithm is run on a
single core of the CPU and it is not optimized for the CPU.  The data
transfer time has been subtracted, so the latency is due only to the
calculations and to the copy to/from the GPU.  The GPU has much better
performance for a large number of fits, while it is surpassed by the
CPU for small number of calculations, as would be expected due to the
excellent multi-threaded processing of GPUs.

\begin{figure}[!h]
  \centering
  \includegraphics[width=3.5in]{figures/TF_SLINK_IB_fixedSize_CPUvsGPU_noDT}
  \caption{Comparison of the algorithm latency on the different GPUs 
to the total latency for the same algorithm run on the CPU (IB 
experimental setup only), as a function of the number of threads.  As would
be expected, the GPU performs better for a large number of fits.}
  \label{fig_TF_SLINK_IB_fixedSize_CPUvsGPU}
\end{figure}

%\begin{figure}[!h]
%  \centering
%  \subfigure[]
%{\label{fig_TF_SLINK_fixedSize_CPUvsGPU}
%\includegraphics[width=3.5in]{figures/TF_SLINK_fixedSize_CPUvsGPU_noDT}}
%\hspace{1mm}
%\subfigure[]
%{\label{fig_TF_IB_fixedSize_CPUvsGPU}
% \includegraphics[width=3.5in]{figures/TF_IB_fixedSize_CPUvsGPU_noDT}}
%\caption{Comparison of the algorithm latency on GPU to the total latency 
%for the same algorithm run on the CPU of the S-LINK (a) and IB (b) 
%experimental setups, as a function of the number of threads.  As would
%be expected, the GPU performs better for a large number of fits. } 
%\end{figure}
%
\subsection{Conclusions}

Following up on initial studies of GPUs to low-latency applications in HEP experiments, 
we have performed additional benchmark measurements for a variety of GPUs and data transfer 
strategies. We see improvements in latency for the more modern InfiniBand communication link 
protocol, compared to the S-LINK protocol. More importantly, we see smaller latencies for the time
it takes to copy memory into the GPU when using some of NVIDIA's GPUDirect utilities. Using GPU-Aware MPI
leads to a decrease in that memory transfer time for large data sizes, while using peer-to-peer communication
strategies shows a strong boost to performance for even small data sizes. As shown in 
Tabs.~\ref{tab_TF_SLINK} and \ref{tab_TF_IB}, the time to copy memory into and out of the GPU represents a significant overhead 
for the total latency. These and other strategies for improving the memory transfer speed, along with a smarter implementation of 
algorithms which reduces the amount of memory transfers, will make GPUs more viable for low-latency applications.

We have also conducted a comparison of different GPUs, showing latency 
measurements for two commercial GPUs (the GTX 285 and GTX 590 from NVIDIA) 
and a GPU optimized for scientific and technical computing (the Tesla M2075, 
also manufactured by NVIDIA). The performance in the InfiniBand+Tesla setup is 
better, but much of this is from the aforementioned memory transfer improvements.
When looking only at the additional latency when performing the track-fitting algorithm, 
the performance of the GTX and Tesla class GPUs are comparable in the region we explored.

Finally, we performed tests while holding the input and output data sizes fixed, only varying the number of 
calculation threads on the GPU. While there is some increase in latency with an increase in the number of
calculation threads, that increase is rather small, especially compared to running the same number of 
calculations on a single CPU thread. If the overhead times can be controlled or even further reduced, GPUs 
may offer a flexible and cost-effective way to achieving a high level of parallelization in applications that demand low latency.

%%%% IEEE wants placement as !t
%\begin{figure}[!t]
%\centering
%%\includegraphics[width=3.5in]{myFigure}
%% where an .eps filename suffix will be assumed under latex, 
%% and a .pdf suffix will be assumed for pdflatex; or what has been declared
%% via \DeclareGraphicsExtensions.
%\caption{Daily abstract submission rate of the 2012 NSS-MIC. }
%\label{fig_sim}
%\end{figure}
%

% use section* for acknowledgement
\section*{Acknowledgment}
The authors would like to thank the Fermilab staff and the FTK group at the 
University of Chicago for their support. This work was supported by the
U.S. Department of Energy and National Science Foundation and the Italian
Istituto Nazionale di Fisica Nucleare.

\begin{thebibliography}{99}
\bibitem{bib_nvidia} \url{http://www.nvidia.com}
\bibitem{bib_cuda}\url{http://www.nvidia.com/object/cuda_home_new.html}
\bibitem{bib_OpenCL}\url{http://www.khronos.org/opencl/}
\bibitem{bib_GPU_TIPP2011} W.~Ketchum \textit{et al.}, \emph{Performance Study of GPUs in Real-Time Trigger Applications for HEP Experiments},  Physics Procedia, Volume 37, 2012, 1965-1972.
\bibitem{bib_filar} W. Iwanski \emph{et al.}, \url{http://hsi.web.cern.ch/HSI/s-link/devices/filar/Welcome.html.}
\bibitem{bib_solar} W. Iwanski \emph{et al.}, \url{http://hsi.web.cern.ch/HSI/s-link/devices/s32pci64/.}
\bibitem{bib_slink} E. van der Bij \emph{et al.}, \emph{S-LINK, A data link interface specification for the LHC era}, IEEE Trans.Nucl.Sci. 44 (1997) 398–402,
\url{http://hsi.web.cern.ch/HSI/s-link/}. \href{http://dx.doi.org/10.1109/23.603679}{doi:10.1109/23.603679}.
\bibitem{bib_Apenet} R.~Ammendola \textit{et al.}, \textit{APEnet+: a 3D 
toroidal network enabling Petaflops scale Lattice QCD simulations on 
commodity clusters},  PoS(Lattice 2010)022.
\bibitem{bib_GPUDirect} \url{https://developer.nvidia.com/gpudirect}
\bibitem{bib_SVT1} B.~Ashmanskas \emph{et al.}, \emph{The CDF Silicon
    Vertex Trigger}, Nucl.~Instrum.~Meth. A \textbf{518} (2004) 532–536. \href{http://dx.doi.org/10.1016/j.nima.2003.11.078}{doi:10.1016/j.nima.2003.11.078}
\bibitem{bib_SVT2} J.~A.~Adelman \emph{et al.}, \emph{The Silicon
    Vertex Trigger upgrade at CDF}, Nucl.~Instrum.~Meth. A \textbf{572} (2007)
  361–364. 
\href{http://dx.doi.org/10.1016/j.nima.2006.10.383}{doi:10.1016/j.nima.2006.10.383}
\bibitem{bib_SVT3} S. Amerio \emph{et al.}, \emph{The GigaFitter: Performance at CDF and perspectives for future applications}, J.Phys.Conf.Ser. 219 (2010)
022001.
\end{thebibliography}

\end{document}

### 2013-09-27 papers/2013/CHEP/paper.tex
add a 'papers' directory with 2012 NSS paper, and CHEP start

--- 2013-09-27 papers/2012/tags/nss2012final/NSS.tex
@@
-\documentclass[journal]{IEEEtran}
+\documentclass[letterpaper]{jpconf}
 \usepackage{graphicx}
-\usepackage{subfigure}
+\usepackage{iopams}
+\usepackage[pdftex,
+      colorlinks=true,
+      urlcolor=blue,       % \href{...}{...} external (URL)
+      filecolor=green,     % \href{...} local file
+      linkcolor=red,       % \ref{...} and \pageref{...}
+      pdftitle={Papers by Peter Wittich},
+      pdfauthor={Peter Wittich},
+      pdfsubject={Just a test},
+      pdfkeywords={test testing testable},
+      pagebackref,
+      pdfpagemode=UseNone,
+       bookmarksopen=true]{hyperref}
 
-\usepackage{hyperref}
-\usepackage{multirow}
-\usepackage[all]{hypcap} % fix links to figs/tabs
-
-\renewcommand{\arraystretch}{1.1} % make tables more readable
-
-% disable the running header on all pages following the title page
-\pagestyle{empty}
-%
 \begin{document}
@@
 % Author list
-\author{S.~Amerio, R.~Ammendola, A.~Biagioni, D.~Bastieri, D.~Benjamin, 
-O.~Frezza, S.~Gelain, W.~Ketchum, Y.~K.~Kim, F.~Lo~Cicero, A.~Lonardo, T.~Liu, D.~Lucchesi, P.~S.~Paolucci, S.~Poprocki, D.~Rossetti, F.~Simula, L.~Tosoratto, 
-G.~Urso, P.~Vicini, and P.~Wittich
-\thanks{Manuscript received November 16, 2012.
-This work was supported by the National Science Foundation, the
-Department of Energy Office of Science and INFN.}% <-this % stops a space
-\thanks{W.~Ketchum and Y.~K.~Kim are with University of Chicago.}%
-\thanks{T.~Liu is with Fermi National Accelerator Laboratory.}%
-\thanks{S.~Poprocki and P.~Wittich are with Cornell University.}%
-\thanks{S.~Amerio is with INFN Padova.}% 
-\thanks{R.~Ammendola is with INFN Tor Vergata.}%
-\thanks{A.~Biagion, O.~Frezza,  F.~ Lo Cicero, A.~Lonardo, P.~S.~Paolucci, D.~Rossetti,  F.~Simula, L.~Tosoratto and P.~Vicini are with INFN Roma}% 
-\thanks{D.~Bastieri, S.~Gelain and D.~Lucchesi are with University of Padova and INFN.}%
-\thanks{G.~Urso is with ORMA Software.}%
-\thanks{D.~Benjamin is with Duke University.}%
+\author{S.~Amerio$^1$, 
+  D.~Bastieri$^1$, 
+  M.~Corvo$^1$, 
+  A.~Gianelle$^1$, 
+  D.~Lucchesi$^1$,
+  T.~Liu$^2$, 
+  R.~Rivera$^2$, 
+  W.~Ketchum$^3$,
+  S.~Poprocki$^4$, P.~Wittich$^4$,
+  A.~Lonardo$^5$, and P.~Vicini$^5$
 }
-% make the title area
-\maketitle
-% after \maketitle, add the following command to remove the header on the title page
-\thispagestyle{empty}
-% abstract usually in passive voice
+\address{$^1$ INFN and University of Padova, Italy}
+\address{$^2$ Fermi National Accelerator Laboratory, Illinois, USA}
+\address{$^3$ Los Alamos National Laboratory, New Mexico, USA}
+\address{$^4$ Cornell University, New York, USA}
+\address{$^5$ INFN Roma, Italy}
+
+\ead{silvia.amerio@pd.infn.it}
+
 \begin{abstract}
+  %%% NSS PAPER abstract
   One of the most important issues that particle physics experiments
@@
   strict time requirement has made the usage of commercial devices
-  inadequate, but recent improvements to Graphics Processing Units (GPUs)
-  have substantially
-  changed the conditions. Thanks to their highly parallel,
-  multi-threaded, multicore architecture with remarkable computational
-  power and high memory bandwidth, these commercial devices may be
-  used in scientific applications, among which the 
-  event selection system (trigger) in particular may benefit, 
-  even at low levels. This paper describes the results of an R\&D
-  project to study the performance of GPU technology for
-  low latency applications, such as HEP fast tracking trigger
-  algorithms.  On two different setups, we measure the latency to
-  transfer data to/from the GPU, exploring the timing
-  of different I/O technologies on different GPU models. We then
-  describe the implementation and the performance of a track fitting
-  algorithm which mimics the CDF Silicon Vertex Tracker.  These
-  studies provide performance benchmarks to investigate the potential
-  and limitations of GPUs for future real-time applications in HEP
-  experiments.
+  inadequate, but recent improvements to Graphics Processing Units
+  (GPUs) have substantially changed the conditions. Thanks to their
+  highly parallel, multi-threaded, multicore architecture with
+  remarkable computational power and high memory bandwidth, these
+  commercial devices may be used in scientific applications, among
+  which the event selection system (trigger) in particular may
+  benefit, even at low levels. This paper describes the results of an
+  R\&D project to study the performance of GPU technology for low
+  latency applications, such as HEP fast tracking trigger algorithms.
+  On two different setups, we measure the latency to transfer data
+  to/from the GPU, exploring the timing of different I/O technologies
+  on different GPU models. We then describe the implementation and the
+  performance of a track fitting algorithm which mimics the CDF
+  Silicon Vertex Tracker.  These studies provide performance
+  benchmarks to investigate the potential and limitations of GPUs for
+  future real-time applications in HEP experiments.
 \end{abstract}
@@
 \section{Introduction}
-\IEEEPARstart{R}{eal} time event reconstruction plays a fundamental
-role in High Energy Physics (HEP) experiments at hadron colliders.
-Reducing the rate of data to be saved on tape from millions to
-hundreds of events per second is critical. To increase the purity of
-the collected samples, rate reduction has to be coupled with an initial
-selection of the most interesting events.  In a typical hadron collider experiment, the
-event rate has to be reduced from tens of MHz to a few kHz.  The selection system
-(trigger) is usually organized in multiple levels, each capable of
-performing a finer selection on more complex physics objects
-describing the event. Trigger 
-systems usually comprise a first level based on custom hardware,
-followed by one or two levels usually based on farms of general
-purpose processors.  The possibility of also using commercial devices
-at a low trigger level is very appealing: they are subject
-to continuous performance improvements driven by the consumer market, are
-less expensive than dedicated hardware, and are easier to support.  Among
-the commercial devices, Graphic Processing Units (GPUs) are of
-particular interest for online selections given their impressive
-computing power (the latest NVIDIA~\cite{bib_nvidia} GPU architecture,
-Kepler, exceeds Teraflop computing power); moreover, high-level
-programming architectures based on C/C++ such as
-\textsc{cuda}~\cite{bib_cuda} and \textsc{opencl}~\cite{bib_OpenCL}
-make these devices more accessible to the general user.  However, GPUs
-are not designed for low-latency response, which is a critical
-requirement in a trigger system.
+%%% Verbatim from NSS paper
+Real-time event reconstruction plays a fundamental role in High Energy
+Physics (HEP) experiments at hadron colliders.  Reducing the rate of
+data to be saved on tape from millions to hundreds of events per
+second is critical. To increase the purity of the collected samples,
+rate reduction has to be coupled with an initial selection of the most
+interesting events.  In a typical hadron collider experiment, the
+event rate has to be reduced from tens of MHz to a few kHz.  The
+selection system (trigger) is usually organized in multiple levels,
+each capable of performing a finer selection on more complex physics
+objects describing the event. Trigger systems usually comprise a first
+level based on custom hardware, followed by one or two levels usually
+based on farms of general purpose processors.  The possibility of also
+using commercial devices at a low trigger level is very appealing:
+they are subject to continuous performance improvements driven by the
+consumer market, are less expensive than dedicated hardware, and are
+easier to support.  Among the commercial devices, Graphic Processing
+Units (GPUs) are of particular interest for online selections given
+their impressive computing power (the latest NVIDIA~\cite{bib_nvidia}
+GPU architecture, Kepler, exceeds Teraflop computing power); moreover,
+high-level programming architectures based on C/C++ such as
+\textsc{cuda} and \textsc{opencl} make these devices more accessible
+to the general user.  However, GPUs are not designed for low-latency
+response, which is a critical requirement in a trigger system.
 
 We showed initial studies on GPU performance in low-latency
-environments ($\sim$100\,$\mu$s) in a previous
-paper~\cite{bib_GPU_TIPP2011}.  In this paper we complement those
-studies measuring the timing performance of different GPU cards in
-different data I/O environments. The algorithm run on the GPU is a
-simplified version of the fast track fitting algorithm of the Silicon
-Vertex Tracker (SVT) system at CDF~\cite{bib_SVT1}. It has been implemented
-and tested on two commercial GPUs (NVIDIA GTX 285 and GTX 590) and on an
-NVIDIA Tesla M2075, optimized for scientific and technical
-computing. To transfer data to the GPU, we have compared the
-performance of S-LINK~\cite{bib_slink} --- 
-used in the second level of the CDF trigger system and which we have prior experience with --- and InfiniBand (IB), a more modern protocol.
+environments ($\sim$100\,$\mu$s) in a previous paper~\cite{TIPP2011,
+  NSS2012}.  In this paper we extend those studies measuring the
+timing performance of different GPU cards in different data I/O
+environments. The algorithm run on the GPU is a complete version of
+the fast track fitting algorithm of the Silicon Vertex Tracker (SVT)
+system at CDF~\cite{SVT1}.
 
-The goal of this study is to investigate the strengths and weaknesses of
-GPUs when applied in a low latency environment, with particular
+The goal of this study is to investigate the strengths and weaknesses
+of GPUs when applied in a low latency environment, with particular
 emphasis on the data transfer latency to/from the GPU and the
-algorithm latency for processing on the GPU in a manner similar to
-a typical HEP trigger application.
+algorithm latency for processing on the GPU in a manner similar to a
+typical HEP trigger application.
 
@@
 equipped with different data transfer links and GPU cards.  In both
-setups two PCs are used, one acting as a transmitter (TX) and the other 
-as a receiver (RX).  Data are transferred from TX to RX, processed on the GPU, and sent back to the receiver (see
-Fig.~\ref{fig_data_flow}).  The latency for a complete loop is measured
-on the transmitter using the time stamp counter register.
+setups two PCs are used, one acting as a transmitter (TX) and the
+other as a receiver (RX).  Data are transferred from TX to RX,
+processed on the GPU, and sent back to the receiver (see
+Fig.~\ref{fig_data_flow}).  The latency for a complete loop is
+measured on the transmitter using the time stamp counter register.
 
@@
 output data is composed of 32-bit words as well, representing the
-results of the fit; when the input data is not processed by the fitting algorithm, the output
-data is a copy of the input words.
+results of the fit; when the input data is not processed by the
+fitting algorithm, the output data is a ....
 
-\begin{figure}[!h]
-\centering
-\includegraphics[width=3.5in]{figures/SetUp-general}
-\caption{Data flow. Data is sent from the transmitter PC to the
-  receiver PC, where it is processed by the GPU before being returned
-  to the transmitter PC. The transmitter plays the role of the
-  detector as the source of the data and as an upstream trigger
-  processor as the data's ultimate sink. The receiver PC plays the
-  role of a component in the trigger system. We consider different I/O
-  protocols (S-LINK and InfiniBand) and both consumer (GTX) and scientific computing (Tesla) GPUs. }
-\label{fig_data_flow}
-\end{figure}
-
-\subsection{S-LINK and GTX cards setup}
-
-Both TX and RX PCs are equipped with a CERN FILAR (Four Input Links
-for ATLAS Readout)~\cite{bib_filar} and an S32PCI64~\cite{bib_solar}
-SOLAR (Single Output Link for ATLAS Readout) PCI interface card for
-I/O communication over S-LINK~\cite{bib_slink}.  The TX PC houses two
-2.4 GHz AMD Opteron 250 (single core) processors while the RX PC
-contains a 3.07 GHz Intel Core i7 CPU 950 (quad core with
-hyperthreading) processor.  The RX PC also contains either an NVIDIA GeForce GTX 285 or GTX 590 GPU in a PCIe slot for the GPU computations. A separate GPU handles display operations.
-
-%SILVIA: I propose to remove the following lines as this is repeated 
-%before the Data Transfer paragraph.
-%The time stamp counter register on the TX PC is used for timing measurements 
-%since it is consistent with but has a higher resolution than the high 
-%precision Linux timers on this system. Long tails in timing 
-%distributions were removed by locking the program to a single CPU core 
-%and moving all system interrupts to the remaining CPUs, and running 
-%at the highest system scheduler priority, on both PCs. 
-%
-
-\subsection{InfiniBand and Tesla cards setup}
-The second setup is a computing cluster composed of 12 identical nodes. 
-Each node contains a Intel Xeon E6520 2.4 GHz CPU and two Tesla M2075 
-GPU cards. The nodes are connected by InfiniBand communication 
-links using Connect-X2 Mellanox or Apenet+ adapters. 
-Apenet+ is an FPGA-based PCIe board supporting peer-to-peer communication with 
-Tesla and Kepler cards~\cite{bib_Apenet}.
-
-\begin{table}[!t]
-  \centering
-  \begin{tabular}{|l|c|c|c|}
-    \hline
-    & GTX 285 & GTX 590 & Tesla M2075 \\
-    \hline
-    \hline
-    %Multiprocessors & & & 14 \\
-    CUDA cores & 240 & 1024 & 448  \\
-    %Peak Single Precision floating point performance (GFlops) & & & 1030 \\
-    Memory interface & GDDR3 & GDDR5 & GDDR5 \\
-    Memory size (GB) & 1 & 3 & 6  \\
-    Memory bandwidth  (GB/s) & 159 & 327 & 150 \\   
-    GPUDirect support & No & No & yes \\
-    \hline
-  \end{tabular}
-  \caption{Capabilities of the GPUs used in this study, according to the manufacturer's specifications.}
-  \label{tab_GPUcards}
-\end{table}
-
-\section{Latency measurements}
-The data flow can be divided into separate steps, each adding its own 
-contribution to the total latency. First, data are transferred from the 
-transmitter to the receiver. Once on the receiver, data
-are copied to the GPU and then are processed by the GPU. 
-%Finally, the output  from the GPU is typically copied back to the receiver PC, and then sent back to the transmitter.
-
-We perform three different sets or measurements:
-\begin{enumerate}
-\item[A)] Latency for data transfer only:\\
-TX $\to$ RX system memory $\to$ TX;
-\item[B)] Latency for data transfer and copy to/from the GPU:\\
-TX $\to$ RX system memory $\to$ GPU memory $\to$ RX system memory $\to$ TX;
-\item[C)] Latency for data transfer, copy to/from the GPU and processing  
-in the GPU, where the data flow is the same as above, except a benchmark algorithm is performed in the GPU.
-\end{enumerate}
-
-The latency of each round-trip is measured on the transmitter using
-the \textsc{rdtsc} time stamp function 
-which was found to be consistent with --- but having a higher resolution than --- the high precision Linux timers on our system. 
-Each test is repeated 10,000
-times and we show the mean and RMS of the repeated measurements. 
-To minimize interrupts we use two strategies first developed during the
-CDF L2 studies. We use \textsc{cpu affinity} to tie our code to a
-single processor, change the scheduler priority of the process to the
-highest value (99) and set the scheduling algorithm to
-\textsc{sched\_fifo} to ensure that our algorithm does not get removed
-from the CPU queue. We also use \textsc{irq affinity} to mask off the
-CPU selected above from all interrupts. Fig.~\ref{fig:affinity} shows that 
-this setup reduces the long tails in the timing distributions.
-\begin{figure}[tbp]
-\centering
-\includegraphics[width=3.5in]{figures/affinity_32_20121017_irqaffinity}
-\caption{Comparison of the timing distributions for default running (solid black)
-and for applying CPU \& IRQ affinities and scheduler priority (dashed
-red). We see a reduction in the long tails, though some remain. }
-\label{fig:affinity}
-\end{figure}
-However, the timing distributions typically exhibit features such as
-multiple humps and long non-Gaussian tails. Preliminarily, we suspect
-these features are due to interactions between the CPU and GPU, or
-possibly power-saving modes on the GPUs, and further studies are
-forthcoming.
-
-\subsection{Data transfer}
-Packets of 32-bit words, from a few bytes to 2 MB, are sent from the receiver
-to the transmitter and back. No GPU is involved in this measurement.
-Due to hardware limitations, the maximum size in input to the S-LINK 
-setup is 4 kB.
-
-The latency as a function of the packet size is shown in 
-Fig.~\ref{fig_DT_SLINK_IB} and in Tab.~\ref{tab_DT}: 
-IB shows a linear increase in latency for 
-packets up to 2 MB, where 1.4 ms are needed for a complete loop.
-A comparison with S-LINK for packets up to 2 kB is presented in the 
-inset of the same figure: IB shows lower latencies and a 
-smaller slope with increasing packet size. About 13 $\mu$s are needed to 
-transfer (round-trip) 4 kB of data.
-
-\begin{table}[!t]
-  \centering
-  \begin{tabular}{|r||r|r|}
-    \hline
-
-    Input  & \multicolumn{2}{|c|}{Data transfer latency ($\mu$s)}  \\
-
-    (Bytes) & \multicolumn{1}{|c|}{S-LINK}   & \multicolumn{1}{|c|}{IB}\\
-
-    \hline
-    \hline
-    8    & 12.1 $\pm$ 0.4 & 3.3    $\pm$ 0.3 \\
-    %\hline
-    32   & 12.6 $\pm$ 0.3 & 3.4    $\pm$ 0.3 \\
-    %\hline
-    128  & 13.2 $\pm$ 0.4 & 6.1    $\pm$ 0.3 \\
-    %\hline
-    512  & 16.6 $\pm$ 0.4 & 7.0    $\pm$ 0.2 \\
-    %\hline
-    1024 & 20.8 $\pm$ 0.2 & 8.3    $\pm$ 0.2 \\
-    %\hline
-    4096 & 40.3 $\pm$ 0.3 & 13.4   $\pm$ 0.3 \\
-    %\hline
-    32k  &                & 34.7   $\pm$ 1.8 \\
-    %\hline
-    65k  &                & 56.0   $\pm$ 1.8 \\
-    %\hline
-    204k &                & 148.3  $\pm$ 1.1\\
-    %\hline
-    512k &                & 350.4  $\pm$ 1.8\\
-    %\hline
-    2M   &                & 1358.1 $\pm$ 1.8\\
-
-    \hline
-  \end{tabular}
-\caption{Data transfer latency. This time measures the round trip 
-from the transmitter to the receiver, without interaction to the GPU. }
-\label{tab_DT}
-\end{table}
-
-\begin{figure}[!t]
-\centering
-\includegraphics[width=3.5in]{figures/DT_SLINK_IB_inset}
-% where an .eps filename suffix will be assumed under latex, 
-% and a .pdf suffix will be assumed for pdflatex; or what has been declared
-% via \DeclareGraphicsExtensions.
-\caption{Data transfer latency as a function of input packets from 32 kB to 
-2 MB and for packets from 1 B to 4 kB (inset, logarithmic scale).}
-\label{fig_DT_SLINK_IB}
-\end{figure}
-
-%\begin{figure}[!h]
-%\centering
-%\includegraphics[width=3.5in]{figures/IB_DT_high}
-%\caption{Data transfer latency for input packets from 32 kB to 2 MB.}
-%\label{fig_DT_IB_HIGH}
-%\end{figure}
-%
-\subsection{Data copy to/from the GPU}
-
-The two experimental setups allow us to test different data transfer 
-strategies to the GPU. The standard data transfer strategy is via the 
-system memory, where the 
-PCIe adapter card and the GPU allocate \emph{separate} buffers on the
-system memory for the copy (as shown in Fig.~\ref{fig_standardDT}).
-This is inefficient, as the data are copied twice in the 
-system memory before being transferred to the GPU/PCIe card. 
-Data may also be transferred using Direct Memory Access (DMA, 
-GPUDirect\cite{bib_GPUDirect}) to the CPU memory:
-the PCIe card and the GPU share the \textit{same} buffer on the CPU memory; as 
-a result the data are copied only once in the CPU memory 
-(Fig.~\ref{fig_GPUDirectV1}). 
-GPUDirect is the standard data transfer mechanism used in the IB-Tesla setup. 
-While the GTX GPUs cannot use GPUDirect functionalities, the firmware for S-LINK 
-PCI cards allows us to specify an address in system memory where the data is deposited, 
-thus achieving a GPUDirect-like setup that avoids an additional copy of the data in system memory.
-In the IB-Tesla setup we can test two additional copy 
-strategies, which are the results of different levels of 
-optimization of the GPUDirect protocol:
-\begin{itemize}
-\item GPU-Aware MPI, where the copy latency is further reduced
-by automatically allocating the buffer on the CPU memory;
-\item peer-to-peer (P2P) strategy, when data are transferred 
-directly to the GPU, without any  
-intermediate copy to the CPU (Fig.~\ref{fig_GPUDirectV2}).
-\end{itemize}
-%
-\begin{figure*}[!t]
-\centering
-\subfigure[]
-{\label{fig_standardDT}
-\includegraphics[width=2in]{figures/noGPUDirect}}
-\hspace{1mm}
-\subfigure[]
-{\label{fig_GPUDirectV1}
-\includegraphics[width=2in]{figures/GPUDirect}}
-\subfigure[]
-{\label{fig_GPUDirectV2}
-\includegraphics[width=2in]{figures/GPUDirect-APE}}
-\caption{Standard data transfer (a), via GPUDirect (b) and via
-  GPUDirect with P2P support (c). In (a), two buffers are required in
-  the main memory. In GPUDirect (b), one of the main memory buffers is
-  eliminated. In GPUDirect with P2P support, data is sent directly
-  from the APEnet+ transceiver to the GPU memory.}
-\end{figure*}
-% No spurious paragraph breaks
-As in the previous measurements, packets of varying size are sent to
-the receiver, copied from the system memory to the GPU and back, and
-finally sent back to the transmitter.  The total round-trip time
-(Tab.~\ref{tab_MC}) comprises the data transfer plus copy latency.
-Measurements done with the P2P protocol are available only for packets
-in the range 32 B -- 4 kB.
-
-For small packet sizes (see Fig.~\ref{fig_DT_MC_SLINK_IB_APE_low}), the 
-S-LINK+GTX and IB+Tesla setups show similar performance, about 
-30 $\mu$s to transfer 100~B of data. For over 100~B of transferred data,  
-the S-LINK latency increases with a much larger slope than IB.
-P2P is the best strategy in terms of latency over the whole considered range 
-(32 B -- 4 kB).
-A comparison between standard GPUDirect and GPU-aware MPI strategy is presented
-in Fig.~\ref{fig_DT_MC_GPUDirect_high}. We find that the GPU-Aware MPI 
-strategy gives consistent reduction only above 200 kB.
-
-\begin{table}[!t]
-  \centering
-  \begin{tabular}{|r||r|r|r|r|}
-    \hline
-    \multicolumn{1}{|c||}{\multirow{2}{*}{Input}} & \multicolumn{4}{|c|}{Data transfer + copy latency  ($\mu$s)}  \\\cline{2-5}
-    \multirow{2}{*}{(Bytes)}
-& \multicolumn{2}{|c|}{GPUDirect}
-%& \multicolumn{1}{|c|}{GPUDirect}
-& \multicolumn{1}{|c|}{GPU-Aware MPI}
-& \multicolumn{1}{|c|}{P2P} \\
-& \multicolumn{1}{|c|}{(S-LINK)}
-& \multicolumn{1}{|c|}{(IB)}
-& \multicolumn{1}{|c|}{(IB)}
-& \multicolumn{1}{|c|}{(Apenet+)} \\
-
-    \hline
-    \hline
-     8    & 29.0 $\pm$ 0.3 & 27.1 $\pm$ 2.0   & 27.3 $\pm$ 0.9 &  \\
-     %\hline
-     32   & 29.3 $\pm$ 0.1 & 27.5 $\pm$ 0.7   & 27.5 $\pm$ 2.4 &  18.1 $\pm$ 1.2 \\
-     %\hline                               
-     128  & 32.6 $\pm$ 1.0 & 29.6 $\pm$ 0.7   & 28.9 $\pm$ 1.9 &  20.6 $\pm$ 3.0\\
-     %\hline                               
-     512  & 38.6 $\pm$ 0.9 & 30.8 $\pm$ 0.7   & 31.5 $\pm$ 0.8 &  20.8 $\pm$ 2.4\\
-     %\hline                               
-     1024 & 46.1 $\pm$ 1.0 & 32.6 $\pm$ 1.7   & 33.5 $\pm$ 2.3 &  21.2 $\pm$ 2.6\\
-     %\hline                               
-     4096 & 86.3 $\pm$ 0.7 & 39.1 $\pm$ 2.2   & 40.5 $\pm$ 1.1  &  25.6 $\pm$ 3.0\\
-     %\hline                               
-     32k  &                & 66.4 $\pm$ 2.3   & 84.9 $\pm$ 2.7 &\\
-     %\hline                               
-     65k  &                & 98.9 $\pm$ 2.1   &  115.4 $\pm$ 2.8 &\\
-    %\hline
-     204k &                & 230.3 $\pm$ 2.4  &  231.0 $\pm$ 4.1 &\\
-     %\hline
-     512k &                & 511.6 $\pm$ 3.3  &  413.0 $\pm$ 3.5 &\\
-     %\hline
-     2M   &                & 1917.8 $\pm$ 6.2 &  1330.5 $\pm$ 4.5 &\\
-
-    \hline
-  \end{tabular}
-\caption{Latency for data transfer plus copy to/from the GPU. The S-LINK measurements here are taken using the GTX 590 GPU, while the IB and Apenet+ measurements use the Tesla M2075 GPU.}
-\label{tab_MC}
-\end{table}
-
-\begin{figure}[!t]
-\centering
-\includegraphics[width=3.5in]{figures/DT_MC_SLINK_IB_APE_low}
-\caption{Data transfer plus copy latency for GPUDirect standard, 
-GPU-Aware MPI and P2P. Input packet size from 1 B (32 B for Apenet+) to 4 kB.}
-\label{fig_DT_MC_SLINK_IB_APE_low}
-\end{figure}
-
-\begin{figure}[!t]
-\centering
-\includegraphics[width=3.5in]{figures/MC_IB_high}
-\caption{Data transfer plus copy latency for GPUDirect standard 
- and GPU-Aware MPI up to 2 MB.}
-\label{fig_DT_MC_GPUDirect_high}
-\end{figure}
-
-\subsection{Data processing on the GPU}
-
-Many tasks performed by trigger systems may benefit from the 
-parallelization available in a GPU (e.g. jet clustering and track finding). 
-For this study, we choose as a benchmark a simplified fast track-fitting
-algorithm which was used in CDF's Silicon Vertex Trigger 
-(SVT)~\cite{bib_SVT1}~\cite{bib_SVT2}.
-This algorithm uses a linearized approximation to track-fitting as 
-implemented in hardware (described in greater detail in~\cite{bib_SVT3}). 
-With the SVT approach, the determination of the track parameters 
-($p_i$) is reduced to a simple scalar product:
-\[
-p_i = \vec{f_i} \cdot \vec{x_i} + q_i,
-\]
-where $\vec{x_i}$ are input silicon hits, and $\vec{f_i}$ and $q_i$ are 
-pre-defined constant sets. For each set of hits, the algorithm
-computes the impact parameter $d_0$, the azimuthal angle $\phi$, 
-the transverse momentum $p_\mathrm{T}$ , and the $\chi^2$ of the
-fitted track by using simple operations such as memory lookup and 
-integer addition and multiplication. In
-our testing of the track-fitting algorithm, each word in input is treated to
-represent a set of silicon hits. While the track-fitting algorithm 
-itself is simple, it must be performed on
-many combinations of hits in a given event, especially in high-occupancy 
-environments. This makes it a good 
-benchmark for testing performance of a GPU, using massive parallelization at low latencies.
-
-In Figs.~\ref{fig_DT_MC_TF_SLINK} and \ref{fig_DT_MC_TF_IB_low}
-and Tabs.~\ref{tab_TF_SLINK} and \ref{tab_TF_IB},
-the total latency for data transfer, 
-copy to the GPU, and processing in the GPU is shown. For comparison, 
- the distributions for data transfer and data transfer plus copy are 
-superimposed.
-
-\begin{table}[!t]
-  \centering
-  \begin{tabular}{|r||r|r|r|}
-    \hline
-
-    \multirow{2}{*}{Input (Bytes)} & \multicolumn{3}{|c|}{ Latency  ($\mu$s)}  \\
-
-                                   & \multicolumn{1}{|c|}{Data transfer}
-                                   & \multicolumn{1}{|c|}{+ copy}
-                                   & \multicolumn{1}{|c|}{+ calculations} \\
-
-    \hline
-    \hline
-    128 & 13.2 $\pm$ 0.4 & 32.6 $\pm$ 1.0 & 39.5 $\pm$ 1.4 \\ 
-    256 & 14.4 $\pm$ 0.4 & 33.9 $\pm$ 0.8 & 41.6 $\pm$ 1.5 \\ 
-    512 & 16.6 $\pm$ 0.4 & 38.6 $\pm$ 0.9 & 46.8 $\pm$ 1.4 \\ 
-    1024 & 20.9 $\pm$ 0.2 & 46.1 $\pm$ 1.0 & 53.9 $\pm$ 1.4 \\ 
-    2048 & 27.7 $\pm$ 0.3 & 59.4 $\pm$ 0.8 & 67.1 $\pm$ 1.2 \\ 
-    4096 & 40.3 $\pm$ 0.2 & 86.3 $\pm$ 0.7 & 93.9 $\pm$ 1.0 \\      
-    \hline
-  \end{tabular}
-\caption{Data transfer + copy + calculation latency (S-LINK + GTX 590).}
-\label{tab_TF_SLINK}
-\end{table}
-
-\begin{table}[!t]
-  \centering
-  \begin{tabular}{|r||r|r|r|}
-    \hline
-
-    \multirow{2}{*}{Input (Bytes)} & \multicolumn{3}{|c|}{ Latency  ($\mu$s)}  \\
-
-                                   & \multicolumn{1}{|c|}{Data transfer}
-                                   & \multicolumn{1}{|c|}{+ copy}
-                                   & \multicolumn{1}{|c|}{+ calculations} \\
-
-    \hline
-    \hline
-     8    & 3.3 $\pm$ 0.3   & 27.1 $\pm$ 2.0 & 33.3 $\pm$ 2.1\\
-     %\hline
-     32   & 3.4 $\pm$ 0.3   & 27.5 $\pm$ 0.7 & 33.2 $\pm$ 2.0\\
-     %\hline
-     128  & 6.1 $\pm$ 0.3  & 29.6 $\pm$ 0.7 & 35.7 $\pm$ 2.0\\
-     %\hline
-     512  & 7.0 $\pm$ 0.2  & 30.8 $\pm$ 0.7 & 36.5 $\pm$ 2.6\\
-     %\hline
-     1024 & 8.3 $\pm$ 0.2  & 32.6 $\pm$ 1.7 & 37.9 $\pm$ 2.7\\
-     %\hline
-     4096 &  13.4 $\pm$ 0.3 &  39.1 $\pm$ 2.2 & 45.8 $\pm$ 2.1\\
-%     %\hline
-%     32k  &  34.7 $\pm$ 1.8 &   66.4 $\pm$ 2.3 & 75.4 $\pm$ 3.9\\
-%     %\hline
-%     262k & 187.6 $\pm$ 1.1 & 280.2 $\pm$ 2.5 & 298.1 $\pm$ 2.7\\
-%     %\hline
-%     512k & 350.4 $\pm$ 1.8 & 511.7 $\pm$ 3.3 &   534.2 $\pm$ 4.4\\
-%     %\hline
-%     2M   &  1358.1 $\pm$ 1.8 & 1917.8 $\pm$ 6.2 & 2001.8 $\pm$ 6.8\\
-%     
-    \hline
-  \end{tabular}
-\caption{Data transfer + copy + calculation latency (IB + Tesla M2075).}
-\label{tab_TF_IB}
-\end{table}
-
-\begin{figure}[!t]
-  \centering
-  \subfigure[]
-{\label{fig_DT_MC_TF_SLINK}
-\includegraphics[width=3.5in]{figures/DT_MC_TF_SLINK}}
-\hspace{1mm}
-\subfigure[]
-{\label{fig_DT_MC_TF_IB_low}
- \includegraphics[width=3.5in]{figures/DT_MC_TF_IB_low}}
-\caption{Total latency for data tranfer, copy to/from the GPU and data 
-processing on the GPU, for S-LINK and GTX 590 card (a) and for IB and Tesla M2075 card (b). Note the $y$ axis goes up to 90 $\mu$s in (a) and 50 $\mu$s in (b).} 
-\end{figure}
-
-%\begin{figure}[!h]
-%  \centering
-%  \includegraphics[width=3.5in]{figures/DT_MC_TF_IB_low}
-%  \caption{Total latency for data tranfer, copy to/from the GPU and data 
-%processing on the GPU, for IB and Tesla card.}
-%  \label{fig_DT_MC_TF_IB_low}
-%\end{figure}
-% 
-Adding the track fitting algorithm on the GPU, the latency increases of
-about 8~$\mu$s for the GTX 590 card and 6 $\mu$s for the Tesla. 
-The two cards have similar performance in terms of calculation latency. 
-The increase in the total latency is fixed as a function of 
-the input data (\textit{i.e.}, the number of calculations since each input 
-word is a track to be fitted), due to the parallel nature of the 
-calculations inside the GPU.
-
-This is visible also in Fig.~\ref{fig_TF_SLINK_IB_fixedSize}, where the total 
-latency is measured for fixed input (2 kB) and output (256 kB) size, and 
-only the number of fits is varied, from 32 to 16,384. The latency 
-distribution is flat up to 1,000 fits and shows a slight slope 
-for larger numbers of fits. The slope is largest for the GTX 285, as 
-expected, due to it having the fewest number of CUDA cores.
-In this plot the latency is dominated by the time needed to transfer 
-the output back to the transmitter.
-Moreover, each point is the mean 
-over 10 different runs, and the error is the relative error on the mean: 
-the smaller this error, the smaller the run-by-run spread --- a highly 
-desirable feature in any real-time event selection system. In both experimental
-setups we find that the run-by-run spread is in the range of 1.5 -- 3 $\mu$s.
-
-\begin{figure}[!t]
-  \centering
-  \includegraphics[width=3.5in]{figures/TF_SLINK_IB_fixedSize}
-  \caption{Total latency (data transfer, copy and track fitting algorithm) for the three different GPU cards used in this study.}
-  \label{fig_TF_SLINK_IB_fixedSize}
-\end{figure}
-
-Figure~\ref{fig_TF_SLINK_IB_fixedSize_CPUvsGPU}
-shows the latency for the same SVT-like track fitting
-algorithm run on the CPU of the experimental setup, compared to the
-performance on the GPU.  We stress that the algorithm is run on a
-single core of the CPU and it is not optimized for the CPU.  The data
-transfer time has been subtracted, so the latency is due only to the
-calculations and to the copy to/from the GPU.  The GPU has much better
-performance for a large number of fits, while it is surpassed by the
-CPU for small number of calculations, as would be expected due to the
-excellent multi-threaded processing of GPUs.
-
-\begin{figure}[!h]
-  \centering
-  \includegraphics[width=3.5in]{figures/TF_SLINK_IB_fixedSize_CPUvsGPU_noDT}
-  \caption{Comparison of the algorithm latency on the different GPUs 
-to the total latency for the same algorithm run on the CPU (IB 
-experimental setup only), as a function of the number of threads.  As would
-be expected, the GPU performs better for a large number of fits.}
-  \label{fig_TF_SLINK_IB_fixedSize_CPUvsGPU}
-\end{figure}
-
-%\begin{figure}[!h]
-%  \centering
-%  \subfigure[]
-%{\label{fig_TF_SLINK_fixedSize_CPUvsGPU}
-%\includegraphics[width=3.5in]{figures/TF_SLINK_fixedSize_CPUvsGPU_noDT}}
-%\hspace{1mm}
-%\subfigure[]
-%{\label{fig_TF_IB_fixedSize_CPUvsGPU}
-% \includegraphics[width=3.5in]{figures/TF_IB_fixedSize_CPUvsGPU_noDT}}
-%\caption{Comparison of the algorithm latency on GPU to the total latency 
-%for the same algorithm run on the CPU of the S-LINK (a) and IB (b) 
-%experimental setups, as a function of the number of threads.  As would
-%be expected, the GPU performs better for a large number of fits. } 
-%\end{figure}
-%
-\subsection{Conclusions}
-
-Following up on initial studies of GPUs to low-latency applications in HEP experiments, 
-we have performed additional benchmark measurements for a variety of GPUs and data transfer 
-strategies. We see improvements in latency for the more modern InfiniBand communication link 
-protocol, compared to the S-LINK protocol. More importantly, we see smaller latencies for the time
-it takes to copy memory into the GPU when using some of NVIDIA's GPUDirect utilities. Using GPU-Aware MPI
-leads to a decrease in that memory transfer time for large data sizes, while using peer-to-peer communication
-strategies shows a strong boost to performance for even small data sizes. As shown in 
-Tabs.~\ref{tab_TF_SLINK} and \ref{tab_TF_IB}, the time to copy memory into and out of the GPU represents a significant overhead 
-for the total latency. These and other strategies for improving the memory transfer speed, along with a smarter implementation of 
-algorithms which reduces the amount of memory transfers, will make GPUs more viable for low-latency applications.
-
-We have also conducted a comparison of different GPUs, showing latency 
-measurements for two commercial GPUs (the GTX 285 and GTX 590 from NVIDIA) 
-and a GPU optimized for scientific and technical computing (the Tesla M2075, 
-also manufactured by NVIDIA). The performance in the InfiniBand+Tesla setup is 
-better, but much of this is from the aforementioned memory transfer improvements.
-When looking only at the additional latency when performing the track-fitting algorithm, 
-the performance of the GTX and Tesla class GPUs are comparable in the region we explored.
-
-Finally, we performed tests while holding the input and output data sizes fixed, only varying the number of 
-calculation threads on the GPU. While there is some increase in latency with an increase in the number of
-calculation threads, that increase is rather small, especially compared to running the same number of 
-calculations on a single CPU thread. If the overhead times can be controlled or even further reduced, GPUs 
-may offer a flexible and cost-effective way to achieving a high level of parallelization in applications that demand low latency.
-
-%%%% IEEE wants placement as !t
-%\begin{figure}[!t]
-%\centering
-%%\includegraphics[width=3.5in]{myFigure}
-%% where an .eps filename suffix will be assumed under latex, 
-%% and a .pdf suffix will be assumed for pdflatex; or what has been declared
-%% via \DeclareGraphicsExtensions.
-%\caption{Daily abstract submission rate of the 2012 NSS-MIC. }
-%\label{fig_sim}
-%\end{figure}
-%
-
-% use section* for acknowledgement
-\section*{Acknowledgment}
+%% gives unnumbered ack in CHEP style
+\ack
 The authors would like to thank the Fermilab staff and the FTK group at the 
@@
 
-\begin{thebibliography}{99}
-\bibitem{bib_nvidia} \url{http://www.nvidia.com}
-\bibitem{bib_cuda}\url{http://www.nvidia.com/object/cuda_home_new.html}
-\bibitem{bib_OpenCL}\url{http://www.khronos.org/opencl/}
-\bibitem{bib_GPU_TIPP2011} W.~Ketchum \textit{et al.}, \emph{Performance Study of GPUs in Real-Time Trigger Applications for HEP Experiments},  Physics Procedia, Volume 37, 2012, 1965-1972.
-\bibitem{bib_filar} W. Iwanski \emph{et al.}, \url{http://hsi.web.cern.ch/HSI/s-link/devices/filar/Welcome.html.}
-\bibitem{bib_solar} W. Iwanski \emph{et al.}, \url{http://hsi.web.cern.ch/HSI/s-link/devices/s32pci64/.}
-\bibitem{bib_slink} E. van der Bij \emph{et al.}, \emph{S-LINK, A data link interface specification for the LHC era}, IEEE Trans.Nucl.Sci. 44 (1997) 398–402,
-\url{http://hsi.web.cern.ch/HSI/s-link/}. \href{http://dx.doi.org/10.1109/23.603679}{doi:10.1109/23.603679}.
-\bibitem{bib_Apenet} R.~Ammendola \textit{et al.}, \textit{APEnet+: a 3D 
-toroidal network enabling Petaflops scale Lattice QCD simulations on 
-commodity clusters},  PoS(Lattice 2010)022.
-\bibitem{bib_GPUDirect} \url{https://developer.nvidia.com/gpudirect}
-\bibitem{bib_SVT1} B.~Ashmanskas \emph{et al.}, \emph{The CDF Silicon
-    Vertex Trigger}, Nucl.~Instrum.~Meth. A \textbf{518} (2004) 532–536. \href{http://dx.doi.org/10.1016/j.nima.2003.11.078}{doi:10.1016/j.nima.2003.11.078}
-\bibitem{bib_SVT2} J.~A.~Adelman \emph{et al.}, \emph{The Silicon
-    Vertex Trigger upgrade at CDF}, Nucl.~Instrum.~Meth. A \textbf{572} (2007)
-  361–364. 
-\href{http://dx.doi.org/10.1016/j.nima.2006.10.383}{doi:10.1016/j.nima.2006.10.383}
-\bibitem{bib_SVT3} S. Amerio \emph{et al.}, \emph{The GigaFitter: Performance at CDF and perspectives for future applications}, J.Phys.Conf.Ser. 219 (2010)
-022001.
-\end{thebibliography}
+%% CHEP2013 recommended style
+\bibliographystyle{iopart-num}
+
+\bibliography{gpu}
 

### 2013-10-03 papers/2013/CHEP/paper.tex
ref and title fixes

@@
       linkcolor=red,       % \ref{...} and \pageref{...}
-      pdftitle={Papers by Peter Wittich},
-      pdfauthor={Peter Wittich},
-      pdfsubject={Just a test},
-      pdfkeywords={test testing testable},
       pagebackref,
@@
 % paper title
-\title{Applications of GPUs to Online Track Reconstruction in HEP Experiments}
+\title{Many-core applications to online track reconstruction in HEP experiments}
 % Author list
@@
   A.~Gianelle$^1$, 
+  W.~Ketchum$^2$,
+  T.~Liu$^3$, 
+  A.~Lonardo$^4$, 
   D.~Lucchesi$^1$,
-  T.~Liu$^2$, 
+  S.~Poprocki$^5$, 
   R.~Rivera$^2$, 
-  W.~Ketchum$^3$,
-  S.~Poprocki$^4$, P.~Wittich$^4$,
-  A.~Lonardo$^5$, and P.~Vicini$^5$
+  P.~Vicini$^4$
+  and 
+  P.~Wittich$^5$,
 }
 \address{$^1$ INFN and University of Padova, Italy}
-\address{$^2$ Fermi National Accelerator Laboratory, Illinois, USA}
-\address{$^3$ Los Alamos National Laboratory, New Mexico, USA}
-\address{$^4$ Cornell University, New York, USA}
-\address{$^5$ INFN Roma, Italy}
+\address{$^2$ Los Alamos National Laboratory, New Mexico, USA}
+\address{$^3$ Fermi National Accelerator Laboratory, Illinois, USA}
+\address{$^4$ INFN Roma, Italy}
+\address{$^5$ Cornell University, New York, USA}
 
@@
 University of Chicago for their support. This work was supported by the
-U.S. Department of Energy and National Science Foundation and the Italian
+U.S. Department of Energy, the U.S. National Science Foundation and the Italian
 Istituto Nazionale di Fisica Nucleare.
 
-%% CHEP2013 recommended style
+%% CHEP2013 recommended style for bibliography
 \bibliographystyle{iopart-num}

### 2013-10-03 papers/2013/CHEP/paper.tex
ref and title fixes

@@
   S.~Poprocki$^5$, 
-  R.~Rivera$^2$, 
+  R.~Rivera$^3$, 
   P.~Vicini$^4$

### 2013-10-03 papers/2013/CHEP/paper.tex
boilerplate text, some placeholder figures

@@
       filecolor=green,     % \href{...} local file
-      linkcolor=red,       % \ref{...} and \pageref{...}
+      linkcolor=blue,       % \ref{...} and \pageref{...}
       pagebackref,
@@
 
+The ultimate goal of this study is to investigate the strengths and weaknesses
+of GPUs when applied in a low latency environment, with particular
+emphasis on the data transfer latency to/from the GPU and the
+algorithm latency for processing on the GPU in a manner similar to a
+typical HEP trigger application.
+
 We showed initial studies on GPU performance in low-latency
@@
 the fast track fitting algorithm of the Silicon Vertex Tracker (SVT)
-system at CDF~\cite{SVT1}.
+system at CDF~\cite{SVT1}. We also compare the performance of GPU
+cards from different manufacturers (AMD vs NVIDIA) and different
+computing environments (\textsc{cuda} and \textsc{opencl}) on the same
+card.
 
-The goal of this study is to investigate the strengths and weaknesses
-of GPUs when applied in a low latency environment, with particular
-emphasis on the data transfer latency to/from the GPU and the
-algorithm latency for processing on the GPU in a manner similar to a
-typical HEP trigger application.
+One of the goals of the studies is to explore using a concrete
+algorithm what gains one can achieve with varying degrees of changes
+to the existing algorithm. Starting with a serial algorithm
+implemented on a CPU, we test an ``embarrassingly parallel'' algorithm
+on the Intel MIC environment. In this case each event is handled
+independently by a core on the accelerator, and the parallelization is
+achieved with only minor changes to the legacy code and is only
+possible in the Intel MIC environment. Next we consider an algorithm
+where we unroll three internal nested loops and run these in parallel
+on a GPU, using the \textsc{cuda} environment. This second approach is
+programmatically more complicated and less trivial to implement. In
+neither case have we re-thought the basic algorithms or the data
+structures used. To achieve optimal performance, these steps would
+have to be taken.  As one might expect, the improvement from the first
+approach is rather modest, and the second approach shows larger
+performance gains.
 
@@
 
+\section{SVT track fitting algorithm}
+
+\section{Results}
+\subsection{Data processing}
+\begin{figure}[tbp]
+  \centering
+  \includegraphics[width=0.9\linewidth]{../../2012/figures/DT_MC_SLINK_IB_APE_low}
+
+  \caption{Algorithm-only comparison for timing as a function of the
+    number of track fits. We compare timing on CPUs (serial), Xeon Phi
+    (embarrassingly parallel), and GPUs (fully parallel.) Conclusion
+    is ...}
+  \label{fig:algo_only_timing}
+\end{figure}
+\begin{figure}[tbp]
+  \centering
+  \includegraphics[width=0.9\linewidth]{../../2012/figures/DT_MC_SLINK_IB_APE_low}
+
+  \caption{Speed-ups for algorithm-only comparison for timing as a
+    function of the number of track fits. We compare timing on CPUs
+    (serial), Xeon Phi (partially parallel), and GPUs (fully
+    parallel.) We conclude ....}
+  \label{fig:algo_only_speedup}
+\end{figure}
+In Fig.~\ref{fig:algo_only_timing} we compare the algorithm time as a
+function of the number of fits performed, for the serial,
+embarrassingly parallel and parallel algorithms. We see that the
+embarrassingly parallel algorithm gives a modest increase with respect
+to the serial (CPU) algorithm. Switching to a fully parallel algorithm
+affords a much more significant speed
+improvement. Fig.~\ref{fig:algo_only_speedup} shows the speed-up for
+each... do we want to show this. Not a lot of new information.
+
+\subsection{Breakdown of computing time}
+\begin{figure}[tbp]
+  \centering
+  \includegraphics[width=0.9\linewidth]{../../2012/figures/DT_MC_SLINK_IB_APE_low}
+  \caption{breakdown of computing time}
+  \label{fig:breakdown}
+\end{figure}
+In Fig.~\ref{fig:breakdown} we show the fractional time spent in
+various parts of the algorithm for the serial algorithm (on a CPU),
+the embarrassingly parallel algorithm (on Intel MIC) and the parallel
+algorithm (on NVidia GPUs), as a function of the number of
+fits. Interesting features are ...
+
 %% gives unnumbered ack in CHEP style

### 2013-10-11 src_mic/mic.h
Parallelize last unpack section

@@ -25,6 +25,23 @@ namespace mic {
 inline void empty() {
 }
 
+
+inline void sort(int* &a, int num) {
+
+  int temp;
+
+  for(int i=0; i<num; i++) {
+    for(int j=i+1; j<num; j++) {
+      if(a[i]>a[j]) {
+        temp=a[i];
+        a[i]=a[j];
+        a[j]=temp;
+      }
+    }
+  }
+}
+
+
 inline void destroy(evt_arrays* evt_dev, fep_arrays* fep_dev, fit_arrays* fit_dev, fout_arrays* fout_dev)
 {
   free(evt_dev);
@@ -282,7 +299,7 @@ inline void gf_unpack(unsigned int const* data_in, int n_words, evt_arrays* evt_
 
 // without thrust...
 
-void k_word_decode(int idx, unsigned int *words, int *ids, int *out1, int *out2, int *out3) {
+void k_word_decode(int idx, unsigned int *words, int *ids, int *out1, int *out2, int *out3, int* &eew, int& ieew) {
 
   /* 
     parallel word_decode kernel.
@@ -297,7 +314,7 @@ void k_word_decode(int idx, unsigned int *words, int *ids, int *out1, int *out2,
   */
 
   int word = words[idx];
-  int ee, ep, lyr;
+  int ee, ep, lyr, it;
 
   lyr = -999; /* Any invalid numbers != 0-7 */
 
@@ -324,6 +341,9 @@ void k_word_decode(int idx, unsigned int *words, int *ids, int *out1, int *out2,
 
   if (ee && ep) { /* End of Event word */
     out1[idx] = word; // ee_word
+    // atomicAdd (on mic) returns the *new* value
+    it = atomicAdd(&ieew, 1);
+    eew[it]=idx;
     lyr = EE_LYR;
   } else if (ee) { /* only EE bit ON is error condition */
     lyr = EE_LYR; /* We have to check */
@@ -347,21 +367,15 @@ void k_word_decode(int idx, unsigned int *words, int *ids, int *out1, int *out2,
   ids[idx] = lyr;
 }
 
-inline void gf_unpack_wot(int n_words, int *ids, int *out1, int *out2, int *out3, struct evt_arrays *evta, int& d_tEvts ) {
+inline void gf_unpack_wot(int start, int stop, int *ids, int *out1, int *out2, int *out3, struct evt_arrays *evta, int evt) {
 
-
-  int tEvts = 0;
-
[…]

### 2013-10-11 src_mic/svt_mic.cc

@@ -5,6 +5,7 @@
 #include <math.h>
 #include "mic.h"
 
+
 // calculate mean and stdev on an array of count floats
 void get_mean(float* times_array, int count, float *mean, float *stdev) {
 
@@ -27,7 +28,7 @@ void get_mean(float* times_array, int count, float *mean, float *stdev) {
 // returns the number of output words
 int svt_mic(extra_data edata, unsigned int *data_in, int n_words, unsigned int *data_rec, int* tev, float timer[3])
 {
-  int totEvts = 0;
+  int totEvts = NEVTS;
   int ie = 0;
   int ow = 0;
   evt_arrays* evt_dev = 0;
@@ -39,7 +40,6 @@ int svt_mic(extra_data edata, unsigned int *data_in, int n_words, unsigned int *
 
 #pragma offload target(mic) \
   in(data_in:length(n_words)) in(edata) \
-  inout(totEvts)  \
   nocopy(evt_dev) nocopy(fep_dev) nocopy(fit_dev) nocopy(fout_dev) \
   out(data_rec:length(n_words)) out(ow) out(timer:length(3))
 {
@@ -55,13 +55,23 @@ int svt_mic(extra_data edata, unsigned int *data_in, int n_words, unsigned int *
   int *out2 = (int *)malloc(sizeW);
   int *out3 = (int *)malloc(sizeW);
 
+  int iee= 0;
+  int *ee = (int *)malloc((NEVTS+1) * sizeof(int));
+  ee[0]=-1;
+
   #pragma omp parallel for
   for (int idx=0; idx < n_words; idx++) {
-    mic::k_word_decode(idx, data_in, ids, out1, out2, out3);
+    mic::k_word_decode(idx, data_in, ids, out1, out2, out3, ee, iee);
   }
   mic::init_evt(evt_dev, fep_dev);
-  mic::gf_unpack_wot(n_words, ids, out1, out2, out3, evt_dev, totEvts);
+  mic::sort(ee, NEVTS+1);
 
+  #pragma omp parallel for
+  for (int idxe=0; idxe < NEVTS; idxe++) {
+    mic::gf_unpack_wot(ee[idxe]+1, ee[idxe+1], ids, out1, out2, out3, evt_dev, idxe);
+  }
+
+  free(ee);
   free(ids);
   free(out1);
   free(out2);

### 2013-10-11 src_mic/svtsim_functions.h

@@ -30,7 +30,7 @@ typedef struct tf_arrays      *tf_arrays_t;
 
 
 //GF
-#define NEVTS 3000
+#define NEVTS 100
 #define NROADS 32
 #define EVT 0 //all words from the input file are considered belonging to the same event
 

### 2013-10-11 src_opencl/svt_gpu_opencl.cpp
changes and updates for working with the AMD card ... things compile, and fitting code is progressing again

@@ -130,12 +130,18 @@ int main(int argc, char* argv[]) {
   gf_init(&tf);
   svtsim_fconread(tf);
 
-  struct evt_arrays *evt; int totEvts;
+  struct evt_arrays *evt = new evt_arrays;
+  int totEvts;
   struct fep_arrays *fep_dev = new fep_arrays;
   struct extra_data *edata_dev = new extra_data;
   struct fit_arrays *fit_dev = new fit_arrays;
 
-  gf_init_evt(&evt);
+  bool is_null=false;
+  if(evt==NULL) is_null=true;;
+
+  printf("Is evt_arrays ptr null? bool=%d",is_null);
+
+  //gf_init_evt(&evt);
 
   gettimeofday(&tBegin, NULL);
 
@@ -196,9 +202,9 @@ int main(int argc, char* argv[]) {
     dev_i = 0;
     plat_i = 1;	
     
-    //this is for CPU...
-    //dev_i = 0;
-    //plat_i = 0;	
+    //this is for AMD card...
+    dev_i = 0;
+    plat_i = 0;	
 
     //============================================================================
     //============================================================================
@@ -256,7 +262,8 @@ int main(int argc, char* argv[]) {
     std::string buildOptions;
     { // create preprocessor defines for the kernel
       char buf[256]; 
-      sprintf(buf,"-cl-mad-enable -D DATAWORD=%s ", dataType.getType().c_str());
+      //sprintf(buf,"-cl-mad-enable -D DATAWORD=%s ", dataType.getType().c_str());
+      sprintf(buf,"-cl-mad-enable -I./");
       buildOptions = buf;
       std::cout << __FANCY__ << "buildOptions = " << buildOptions << std::endl;
     }
@@ -287,9 +294,39 @@ int main(int argc, char* argv[]) {
     std::cout << __FANCY__ << "Entry point set to " << kernelFunc_fep_set << std::endl;
 
     //====================================
+    //build program source
+    std::string kernel_fit_file = "gf_fit.cl";
+    std::ifstream fit_file((kernel_fit_file).c_str());
[…]

### 2013-10-11 src_opencl/svt_utils_opencl.h

@@ -65,9 +65,9 @@ struct extra_data {
 
 struct fit_arrays {
 
-  unsigned long fit_fit[NEVTS][6][MAXROAD][MAXCOMB][MAXCOMB5H];
-  int fit_err[NEVTS][MAXROAD][MAXCOMB][MAXCOMB5H];
   int fit_err_sum[NEVTS];
+  int fit_fit[NEVTS][6][MAXROAD][MAXCOMB][MAXCOMB5H];
+  int fit_err[NEVTS][MAXROAD][MAXCOMB][MAXCOMB5H];
 
   //int dummy;
 

### 2013-10-14 papers/2013/CHEP/paper.tex
moar dummy plots
Inserted figures.

@@
 \usepackage{graphicx}
+\usepackage{subfigure}
 \usepackage{iopams}
@@
   R.~Rivera$^3$, 
+  L.~Tosoratto$^4$,
   P.~Vicini$^4$
@@
 \begin{abstract}
-  %%% NSS PAPER abstract
-  One of the most important issues that particle physics experiments
-  at hadron colliders have to solve is real-time selection of the most
-  interesting events. Typical collision frequencies do not allow all
-  events to be written to tape for offline analysis, and in most
-  cases, only a small fraction can be saved. The most commonly used
-  strategy is based on two or three selection levels, with the low
-  level ones usually exploiting dedicated hardware to decide within a
-  few to ten microseconds if the event should be kept or not. This
-  strict time requirement has made the usage of commercial devices
-  inadequate, but recent improvements to Graphics Processing Units
-  (GPUs) have substantially changed the conditions. Thanks to their
-  highly parallel, multi-threaded, multicore architecture with
-  remarkable computational power and high memory bandwidth, these
-  commercial devices may be used in scientific applications, among
-  which the event selection system (trigger) in particular may
-  benefit, even at low levels. This paper describes the results of an
-  R\&D project to study the performance of GPU technology for low
-  latency applications, such as HEP fast tracking trigger algorithms.
-  On two different setups, we measure the latency to transfer data
-  to/from the GPU, exploring the timing of different I/O technologies
-  on different GPU models. We then describe the implementation and the
-  performance of a track fitting algorithm which mimics the CDF
-  Silicon Vertex Tracker.  These studies provide performance
-  benchmarks to investigate the potential and limitations of GPUs for
-  future real-time applications in HEP experiments.
+One of the most important issues facing particle physics experiments at hadron 
+colliders is real-time selection of interesting events for offline storage. 
+Collision frequencies do not allow all events to be written to tape for 
+offline analysis, and in most cases, only a small fraction can be saved. 
+Typical trigger systems use commercial computers in the final stage of 
+processing. Much of the effort is focused on understanding the latency for 
+trigger systems. In this paper we describe updates to a previous study of 
+latencies in GPU for potential trigger applications, where we measured the 
+latency to transfer data to/from the GPU, exploring the timing of different 
+I/O technologies on different GPU models.  Those studies, where a simplified 
+track fitting algorithm was parallelized and run on a GPU, show that latencies 
+of few tens of microseconds can be achieved to transfer and process packets of 
+4 kB of data, combining the modern Infiniband data transfer technology with 
+direct access to GPU memory allowed by NVIDIA GPUDirect utilities.  We now 
+have expanded our latency studies to include other multi-core systems 
+(Intel Xeon Phi in addition to NVIDIA GPUs). We also discuss the implementation 
+of a scaled-up version of the algorithm 
+used at CDF for online track reconstruction - the SVT algorithm - as a 
+realistic test-case for low-latency trigger systems using new computing 
+architectures for LHC experiment.
+
 \end{abstract}
@@
 \textsc{cuda} and \textsc{opencl} make these devices more accessible
-to the general user.  However, GPUs are not designed for low-latency
-response, which is a critical requirement in a trigger system.
-
+to the general user.  
 The ultimate goal of this study is to investigate the strengths and weaknesses
@@
 card. 
-
 One of the goals of the studies is to explore using a concrete
@@
 
+\section{SVT track fitting algorithm}
+The Silicon Vertex Trigger (SVT)~\cite{bib_SVT1}~\cite{bib_SVT2} was a 
+track reconstruction processor used at 
+trigger level in the CDF experiment at Tevatron accelerator. It reconstructs 
+tracks in about 20 $\mu s$ in time in two steps: for each event 
+first low resolution tracks (\textit{roads}) are found among the hits left 
+in the tracking detector; then track fitting is performed on 
+all possible combinations of hits inside a road. 
+This algorithm uses a linearized approximation to track-fitting as 
+implemented in hardware (described in greater detail in~\cite{bib_SVT3}). 
+With the SVT approach, the determination of the track parameters 
+($p_i$) is reduced to a simple scalar product:
+\[
+p_i = \vec{f_i} \cdot \vec{x_i} + q_i,
+\]
+where $\vec{x_i}$ are input silicon hits, and $\vec{f_i}$ and $q_i$ are 
+pre-defined constant sets. For each set of hits, the algorithm
+computes the impact parameter $d_0$, the azimuthal angle $\phi$, 
+the transverse momentum $p_\mathrm{T}$ , and the $\chi^2$ of the
+fitted track by using simple operations such as memory lookup and 
+integer addition and multiplication.
+
 \section{Experimental setup and data flow}
-The studies are performed on two different experimental setups,
-equipped with different data transfer links and GPU cards.  In both
-setups two PCs are used, one acting as a transmitter (TX) and the
+
+The many-core devices used in this study are listed in Table~\ref{t_hwspecs}.
+
+\begin{table}[!t]
+  \centering
+  \begin{tabular}{|l|c|c|c|c|}
+    \hline
+    Model & Tesla M2050 & Tesla K20m & GeForce GTX  Titan & MIC 5110P \\
+    \hline
+    \hline
+    Performance (SP, GFlops) & 1030 & 3520 & 4500 & 2022 \\
+    Memory bandwidth  (GB/s) & 148 & 208 & 288  & 320\\   
+    Memory size (GB) & 3 & 5 & 6 & 8 \\
+    Number of cores & 448 & 2496 & 2688 & 240 \\
+    Clock speed (GHz) & 1.15 & 0.706 & 0.837 & 1.053 \\
+    \hline
+  \end{tabular}
+  \caption{Capabilities of the many-core devices used in this study, according to the manufacturer's specifications.}
+  \label{tab_hwspecs}
+\end{table}
+
+To measure the data transfer latency a setup made of two PCs is used, 
+one acting as a transmitter (TX) and the
 other as a receiver (RX).  Data are transferred from TX to RX,
-processed on the GPU, and sent back to the receiver (see
-Fig.~\ref{fig_data_flow}).  The latency for a complete loop is
-measured on the transmitter using the time stamp counter register.
+processed on the GPU and sent back to the receiver (see
+Fig.~\ref{fig_data_flow}).  
+The GPU on the RX is a Tesla M2075, similar to the M2050 used for the data processing tests. The latency for a complete loop is
+measured on the transmitter using standard C libraries.
+
+\begin{figure}[!h]
+\centering
+\includegraphics[width=3.5in]{figures/SetUp-general}
+\caption{Data flow. Data is sent from the transmitter PC to the
+  receiver PC, where it is processed by the GPU before being returned
+  to the transmitter PC. The transmitter plays the role of the
+  detector as the source of the data and as an upstream trigger
+  processor as the data's ultimate sink. The receiver PC plays the
+  role of a component in the trigger system. }
+\label{fig_data_flow}
+\end{figure}
 
@@
 
-The input data is a set of 32-bit words, each representing a set of
-hits from the detector. When the track fitting algorithm is run, the
-output data is composed of 32-bit words as well, representing the
-results of the fit; when the input data is not processed by the
-fitting algorithm, the output data is a ....
+\section{Results}
 
-\section{SVT track fitting algorithm}
+\subsection{Data processing}
 
-\section{Results}
-\subsection{Data processing}
+\begin{figure*}[!t]
+\centering
+\subfigure[]
+{\label{fig:algo_only_timing}
+\includegraphics[width=3in]{figures/TimeComp_MIC.pdf}}
+\hspace{1mm}
+\subfigure[]
+{\label{fig_breakdown_MIC}
+\includegraphics[width=3in]{figures/TimeCompZoom_MIC.pdf}}
+\caption{breakdown}
+\end{figure*}
+
+% \begin{figure}[tbp]
+%   \centering
+%   \includegraphics[width=0.5\linewidth]{figures/TimeComp_MIC.pdf}
+%    \caption{Algorithm-only comparison for timing as a function of the
+%     number of track fits. We compare timing on CPUs (serial), Xeon Phi
+%     (embarrassingly parallel), and GPUs (fully parallel.) }
+%   \label{fig:algo_only_timing}
+% \end{figure}
+%
+% \begin{figure}[tbp]
+%   \centering
+%   \includegraphics[width=0.5\linewidth]{figures/TimeCompZoom_MIC.png} 
+%   \caption{ZOOM}
+%   \label{fig:algo_only_timing_zoom}
+% \end{figure}
+%
  \begin{figure}[tbp]
    \centering
-  \includegraphics[width=0.9\linewidth]{../../2012/figures/DT_MC_SLINK_IB_APE_low}
-
+   \includegraphics[width=0.5\linewidth]{figures/Speedup_MIC.pdf}
    \caption{Algorithm-only comparison for timing as a function of the
      number of track fits. We compare timing on CPUs (serial), Xeon Phi
-    (embarrassingly parallel), and GPUs (fully parallel.) Conclusion
-    is ...}
-  \label{fig:algo_only_timing}
-\end{figure}
-\begin{figure}[tbp]
-  \centering
-  \includegraphics[width=0.9\linewidth]{../../2012/figures/DT_MC_SLINK_IB_APE_low}
-
-  \caption{Speed-ups for algorithm-only comparison for timing as a
-    function of the number of track fits. We compare timing on CPUs
-    (serial), Xeon Phi (partially parallel), and GPUs (fully
-    parallel.) We conclude ....}
+     (embarrassingly parallel), and GPUs (fully parallel.) }
    \label{fig:algo_only_speedup}
  \end{figure}
+
 In Fig.~\ref{fig:algo_only_timing} we compare the algorithm time as a
@@
 \subsection{Breakdown of computing time}
-\begin{figure}[tbp]
+
+\begin{figure*}[!t]
 \centering
-  \includegraphics[width=0.9\linewidth]{../../2012/figures/DT_MC_SLINK_IB_APE_low}
-  \caption{breakdown of computing time}
-  \label{fig:breakdown}
-\end{figure}
+\subfigure[]
+{\label{fig_breakdown_CPU}
+\includegraphics[width=1.8in]{figures/Cpui7_perc.pdf}}
+\hspace{1mm}
+\subfigure[]
+{\label{fig_breakdown_MIC}
+\includegraphics[width=1.8in]{figures/Mic5_perc_nk.pdf}}
+\subfigure[]
+{\label{fig_breakdown_GPU}
+\includegraphics[width=1.8in]{figures/titan_perc.pdf}}
+\caption{breakdown}
+\end{figure*}
+
 In Fig.~\ref{fig:breakdown} we show the fractional time spent in
@@
 fits. Interesting features are ...
+
+\subsection{Data transfer}
+
+\begin{figure*}[!t]
+\centering
+\subfigure[]
+{\label{fig_standardDT}
+\includegraphics[width=2in]{figures/noGPUDirect}}
+\hspace{1mm}
+\subfigure[]
+{\label{fig_GPUDirectV1}
+\includegraphics[width=2in]{figures/GPUDirect}}
+\subfigure[]
+{\label{fig_GPUDirectV2}
+\includegraphics[width=2in]{figures/GPUDirect-APE}}
+\caption{Standard data transfer (a), via GPUDirect (b) and via
+  GPUDirect with P2P support (c). In (a), two buffers are required in
+  the main memory. In GPUDirect (b), one of the main memory buffers is
+  eliminated. In GPUDirect with P2P support, data is sent directly
+  from the APEnet+ transceiver to the GPU memory.}
+\end{figure*}
+
+\begin{figure}[tbp]
+  \centering
+  \includegraphics[width=0.7\linewidth]{figures/datatransfer.pdf}
+  \caption{data transfer latencies}
+  \label{fig:xferlatency}
+\end{figure}
+
+\begin{figure}[tbp]
+  \centering
+  \includegraphics[width=0.7\linewidth]{figures/cudaware.pdf}
+  \caption{data transfer latencies}
+  \label{fig:transferOnly}
+\end{figure}
+
+In Fig.~\ref{fig:xferlatency} we show the latency for GPU direct v1,
+CUDA-aware MPI and P2P, as a function of the packet size. Does not
+include the algorithm time.
 
@@
 
+\begin{thebibliography}{99}
+\bibitem{bib_nvidia} \url{http://www.nvidia.com}
+\bibitem{bib_cuda}\url{http://www.nvidia.com/object/cuda_home_new.html}
+\bibitem{bib_OpenCL}\url{http://www.khronos.org/opencl/}
+%\bibitem{bib_GPU_TIPP2011} W.~Ketchum \textit{et al.}, \emph{Performance Study of GPUs in Real-Time Trigger Applications for HEP Experiments},  Physics Procedia, Volume 37, 2012, 1965-1972.
+%\bibitem{bib_Apenet} R.~Ammendola \textit{et al.},% \emph{APEnet+: a 3D toroidal network enabling Petaflops scale Lattice QCD simulations on commodity clusters}  PoS(Lattice 2010)022.
+%\bibitem{bib_GPUDirect} \url{https://developer.nvidia.com/gpudirect}
+%\bibitem{bib_SVT1} B.~Ashmanskas \emph{et al.}, \emph{The CDF Silicon    Vertex Trigger}, Nucl.~Instrum.~Meth.
+A \textbf{518} (2004) 532–536.
+\href{http://dx.doi.org/10.1016/j.nima.2003.11.078}{doi:10.1016/j.nima.2003.11.078}
+%\bibitem{bib_SVT2} J.~A.~Adelman \emph{et al.}, \emph{The Silicon    Vertex Trigger upgrade at CDF}, Nucl.~Instrum.~Meth. A \textbf{572} (2007)  361–364. 
+%\href{http://dx.doi.org/10.1016/j.nima.2006.10.383}{doi:10.1016/j.nima.2006.10.383}
+%\bibitem{bib_SVT3} S. Amerio \emph{et al.}, \emph{The GigaFitter: Performance at CDF and perspectives for future applications}, J.Phys.Conf.Ser. 219 (2010) 022001.
+
+\end{thebibliography}
+
 \end{document}

### 2013-10-15 papers/2013/CHEP/paper.tex
Added captions and all text (draft version) except for Conclusions

@@
 \usepackage{subfigure}
+\usepackage{url}
 \usepackage{iopams}
@@
 We showed initial studies on GPU performance in low-latency
-environments ($\sim$100\,$\mu$s) in a previous paper~\cite{TIPP2011,
-  NSS2012}.  In this paper we extend those studies measuring the
+environments ($\sim$100\,$\mu$s) in a previous paper~\cite{bib_tipp2011,bib_nss2012}.  In this paper we extend those studies measuring the
 timing performance of different GPU cards in different data I/O
@@
 the fast track fitting algorithm of the Silicon Vertex Tracker (SVT)
-system at CDF~\cite{SVT1}. We also compare the performance of GPU
-cards from different manufacturers (AMD vs NVIDIA) and different
-computing environments (\textsc{cuda} and \textsc{opencl}) on the same
-card. 
+system at CDF~\cite{bib_SVT1}. 
+%We also compare the performance of GPU
+%cards from different manufacturers (AMD vs NVIDIA) and different
+%computing environments (\textsc{cuda} and \textsc{opencl}) on the same
+%card. 
 One of the goals of the studies is to explore using a concrete
@@
 to the existing algorithm. Starting with a serial algorithm
-implemented on a CPU, we test an ``embarrassingly parallel'' algorithm
+implemented on a CPU, we test an \textit{embarrassingly parallel} algorithm
 on the Intel MIC environment. In this case each event is handled
@@
 where we unroll three internal nested loops and run these in parallel
-on a GPU, using the \textsc{cuda} environment. This second approach is
-programmatically more complicated and less trivial to implement. In
+on a GPU, using the \textsc{cuda}~\ref{bib_cuda} environment. This second 
+approach is programmatically more complicated and less trivial to implement. In
 neither case have we re-thought the basic algorithms or the data
@@
 \section{SVT track fitting algorithm}
-The Silicon Vertex Trigger (SVT)~\cite{bib_SVT1}~\cite{bib_SVT2} was a 
+The Silicon Vertex Trigger (SVT)~\cite{bib_SVT1}~\cite{bib_SVT2} is a 
 track reconstruction processor used at 
@@
 first low resolution tracks (\textit{roads}) are found among the hits left 
-in the tracking detector; then track fitting is performed on 
-all possible combinations of hits inside a road. 
+in the tracking detector by charged particles; then track fitting is 
+performed on all possible combinations of hits inside a road. 
 This algorithm uses a linearized approximation to track-fitting as 
@@
 
+We ported the track fitting part to parallel devices, as it is well suited to 
+be parallelized.
+
+\subsection{Code implementation}
+The starting point of of our studies is the SVT track fitting simulation code, 
+written in C language. SVT track fitting is divided into three main functions: 
+first the unpacking of input data and filling of all the necessary arrays, 
+followed by the computation of all possible combinations of hits in each road 
+and finally the fit on each combination. Three main loops are 
+present - on events, roads and hit combinations. 
+To be run on GPUs, the code has been ported to \textsc{cuda}: each step - 
+unpack, combine and track fit - is performed by a specific kernel;
+the three nested loops are unrolled so that each GPU thread processes a 
+single combination of hits. 
+To run on MIC, we adopted the so-called \textit{embarassingly parallel} 
+approach and used pragma OpenMP for statements to unroll the external loop 
+on the events, so that each core processes a single event: the porting 
+requires much less effort compared to CUDA, but the level of parallelism is 
+limited.
+
 \section{Experimental setup and data flow}
-
-The many-core devices used in this study are listed in Table~\ref{t_hwspecs}.
+The many-core devices used in this study are listed in Table~\ref{tab_hwspecs}.
 
@@
 
-To measure the data transfer latency a setup made of two PCs is used, 
-one acting as a transmitter (TX) and the
-other as a receiver (RX).  Data are transferred from TX to RX,
-processed on the GPU and sent back to the receiver (see
+To measure the data transfer latency we use a computing cluster composed of 
+12 identical nodes.  Each node contains a Intel Xeon E6520 2.4 GHz 
+CPU and two Tesla M2075 GPU cards. The nodes are connected by InfiniBand 
+communication links using Connect-X2 Mellanox or Apenet+ adapters. 
+Apenet+ is an FPGA-based PCIe board supporting peer-to-peer communication with 
+Tesla and Kepler cards~\cite{bib_Apenet}.
+Two nodes of this cluster are used to measure data transfer latency,
+one acting as a transmitter and the
+other as a receiver.  Data are transferred from the transmitter to the 
+receiver, processed on the GPU and sent back to the receiver (see
 Fig.~\ref{fig_data_flow}).  
-The GPU on the RX is a Tesla M2075, similar to the M2050 used for the data processing tests. The latency for a complete loop is
-measured on the transmitter using standard C libraries.
+The latency for a complete loop is measured on the transmitter using 
+standard C libraries.
 
@@
 \section{Results}
+We use as input data events with a fixed number of roads and combinations: 
+each event has 2048 combinations to be fitted. To explore different data 
+taking conditions the number of events ranges from one to 3000, i.e. from 2048 
+to about six millions of combinations to fit.
 
 \subsection{Data processing}
+Each data sample is processed 100 times by the track fitting algorithm. 
+The average latency as a function of the number of fits is presented in 
+Fig.~\ref{fig:algo_only_timing} for the serial,
+embarrassingly parallel and parallel algorithms. We see that the
+embarrassingly parallel algorithm gives a modest increase with respect
+to the serial (CPU) algorithm. Switching to a fully parallel algorithm
+affords a much more significant speed improvement. 
+Accelerators performance drop with decreasing number of fits, as can be seen 
+in Fig.~\ref{fig:algo_only_timing_zoom}.
+Fig.~\ref{fig:algo_only_speedup} shows the speed-up with respect to the serial 
+algorithm run on a standard CPU (Intel Xeon E5630): the maximum gain is 
+obtained processing at least 500 events. This means that to fully exploit 
+parallel architectures millions of fits have to be performed in parallel.
+
+%\begin{figure*}[!t]
+%\centering
+%\subfigure[]
+%{\label{fig:algo_only_timing}
+%\includegraphics[width=3in]{figures/TimeComp_MIC.pdf}}
+%\hspace{1mm}
+%\subfigure[]
+%{\label{fig_algo_only_timing_zoom}
+%\includegraphics[width=3in]{figures/TimeCompZoom_MIC.pdf}}
+%\caption{breakdown}
+%\end{figure*}
+%
+
+ \begin{figure}[!tbp]
+    \centering
+    \includegraphics[width=0.9\linewidth]{figures/TimeComp_MIC.pdf}
+     \caption{Algorithm-only comparison for timing as a function of the
+      number of track fits. We compare timing on CPUs (serial), Xeon Phi
+      (embarrassingly parallel), and GPUs (fully parallel). }
+    \label{fig:algo_only_timing}
+  \end{figure}
+
+  \begin{figure}[!tbp]
+    \centering
+    \includegraphics[width=0.9\linewidth]{figures/TimeCompZoom_MIC.png} 
+    \caption{Algorithm-only comparison for timing as a function of the
+      number of track fits: zoom in the low number of fits region.}
+    \label{fig:algo_only_timing_zoom}
+  \end{figure}
+
+ \begin{figure}[!tbp]
+   \centering
+   \includegraphics[width=0.9\linewidth]{figures/Speedup_MIC.pdf}
+   \caption{Speed-up with respect to a standard CPU (Intel Xeon E5630) .}
+   \label{fig:algo_only_speedup}
+ \end{figure}
+
+\subsubsection{Breakdown of computing time}
+In Fig.~\ref{fig:breakdown} we show the fractional time spent in
+various parts of the algorithm for the embarrassingly parallel algorithm 
+(on Intel MIC) and the parallel algorithm (on NVidia Titan GPU), as a 
+function of the number of fits. On both accelerator cards the fractional times
+are constant over 500 input events, where computing resources are saturated. 
+Differently from MIC, the fit stage takes most of the time on the GPU: this 
+could be explained with the intense memory access intrinsic in 
+this part of the algorithm.
 
@@
 \centering
+%\subfigure[]
+%{\label{fig:breakdown_CPU}
+%\includegraphics[width=0.4\linewidth]{figures/Cpui7_perc.pdf}}
+%\hspace{1mm}
 \subfigure[]
-{\label{fig:algo_only_timing}
-\includegraphics[width=3in]{figures/TimeComp_MIC.pdf}}
-\hspace{1mm}
+{\label{fig:breakdown_MIC}
+\includegraphics[width=0.45\linewidth]{figures/Mic5_perc_nk.pdf}}
 \subfigure[]
-{\label{fig_breakdown_MIC}
-\includegraphics[width=3in]{figures/TimeCompZoom_MIC.pdf}}
-\caption{breakdown}
+{\label{fig:breakdown_GPU}
+\includegraphics[width=0.45\linewidth]{figures/titan_perc.pdf}}
+\caption{Breakdown of computing time.}
 \end{figure*}
 
-% \begin{figure}[tbp]
-%   \centering
-%   \includegraphics[width=0.5\linewidth]{figures/TimeComp_MIC.pdf}
-%    \caption{Algorithm-only comparison for timing as a function of the
-%     number of track fits. We compare timing on CPUs (serial), Xeon Phi
-%     (embarrassingly parallel), and GPUs (fully parallel.) }
-%   \label{fig:algo_only_timing}
-% \end{figure}
-%
-% \begin{figure}[tbp]
-%   \centering
-%   \includegraphics[width=0.5\linewidth]{figures/TimeCompZoom_MIC.png} 
-%   \caption{ZOOM}
-%   \label{fig:algo_only_timing_zoom}
-% \end{figure}
-%
- \begin{figure}[tbp]
-   \centering
-   \includegraphics[width=0.5\linewidth]{figures/Speedup_MIC.pdf}
-   \caption{Algorithm-only comparison for timing as a function of the
-     number of track fits. We compare timing on CPUs (serial), Xeon Phi
-     (embarrassingly parallel), and GPUs (fully parallel.) }
-   \label{fig:algo_only_speedup}
- \end{figure}
-
-In Fig.~\ref{fig:algo_only_timing} we compare the algorithm time as a
-function of the number of fits performed, for the serial,
-embarrassingly parallel and parallel algorithms. We see that the
-embarrassingly parallel algorithm gives a modest increase with respect
-to the serial (CPU) algorithm. Switching to a fully parallel algorithm
-affords a much more significant speed
-improvement. Fig.~\ref{fig:algo_only_speedup} shows the speed-up for
-each... do we want to show this. Not a lot of new information.
-
-\subsection{Breakdown of computing time}
-
-\begin{figure*}[!t]
-\centering
-\subfigure[]
-{\label{fig_breakdown_CPU}
-\includegraphics[width=1.8in]{figures/Cpui7_perc.pdf}}
-\hspace{1mm}
-\subfigure[]
-{\label{fig_breakdown_MIC}
-\includegraphics[width=1.8in]{figures/Mic5_perc_nk.pdf}}
-\subfigure[]
-{\label{fig_breakdown_GPU}
-\includegraphics[width=1.8in]{figures/titan_perc.pdf}}
-\caption{breakdown}
-\end{figure*}
-
-In Fig.~\ref{fig:breakdown} we show the fractional time spent in
-various parts of the algorithm for the serial algorithm (on a CPU),
-the embarrassingly parallel algorithm (on Intel MIC) and the parallel
-algorithm (on NVidia GPUs), as a function of the number of
-fits. Interesting features are ...
-
 \subsection{Data transfer}
+The experimental setup described in Fig.~\ref{fig_data_flow} allows us to 
+test different data transfer  strategies to the GPU. The standard data transfer 
+strategy is via the system memory, where the 
+PCIe adapter card and the GPU allocate \emph{separate} buffers on the
+system memory for the copy (as shown in Fig.~\ref{fig_standardDT}).
+This is inefficient, as the data are copied twice in the 
+system memory before being transferred to the GPU/PCIe card. 
+Data may also be transferred using Direct Memory Access (DMA, 
+GPUDirect\cite{bib_GPUDirect}) to the CPU memory:
+the PCIe card and the GPU share the \textit{same} buffer on the CPU memory; as 
+a result the data are copied only once in the CPU memory 
+(Fig.~\ref{fig_GPUDirectV1}). 
+With our experimental setup two additional copy strategies can be tested
+which are the results of different levels of 
+optimization of the GPUDirect protocol:
+\begin{itemize}
+\item CUDA-Aware MPI, where the copy latency is further reduced
+by automatically allocating the buffer on the CPU memory;
+\item peer-to-peer (P2P) strategy, when data are transferred 
+directly to the GPU, without any  
+intermediate copy to the CPU (Fig.~\ref{fig_GPUDirectV2}).
+\end{itemize}
 
@@
 {\label{fig_standardDT}
-\includegraphics[width=2in]{figures/noGPUDirect}}
+\includegraphics[width=0.3\linewidth]{figures/noGPUDirect}}
 \hspace{1mm}
@@
 {\label{fig_GPUDirectV1}
-\includegraphics[width=2in]{figures/GPUDirect}}
+\includegraphics[width=0.3\linewidth]{figures/GPUDirect}}
 \subfigure[]
 {\label{fig_GPUDirectV2}
-\includegraphics[width=2in]{figures/GPUDirect-APE}}
+\includegraphics[width=0.3\linewidth]{figures/GPUDirect-APE}}
 \caption{Standard data transfer (a), via GPUDirect (b) and via
@@
   \centering
-  \includegraphics[width=0.7\linewidth]{figures/datatransfer.pdf}
-  \caption{data transfer latencies}
+  \includegraphics[width=0.9\linewidth]{figures/datatransfer.pdf}
+  \caption{Total latency (data transfer, copy to/from the GPU and data processing on the GPU) as a function of data buffer sizes, for three different levels of optimization of GPUDirect: v1.0, CUDA-aware MPI and peer-to-peer.
+The smallest transferred data packed is 600 kB.}
   \label{fig:xferlatency}
@@
 
+In Fig.~\ref{fig:xferlatency} we show the total latency (data transfer, copy to/from the GPU and data processing on the GPU) as a function of data packet size when data are transferred using GPUDirect v1.0,
+CUDA-aware MPI and P2P. For the packet sizes over 600 kB CUDA-aware MPI gives the best performance.
+
 \begin{figure}[tbp]
   \centering
-  \includegraphics[width=0.7\linewidth]{figures/cudaware.pdf}
+  \includegraphics[width=0.9\linewidth]{figures/cudaware.pdf}
   \caption{data transfer latencies}
@@
 
-In Fig.~\ref{fig:xferlatency} we show the latency for GPU direct v1,
-CUDA-aware MPI and P2P, as a function of the packet size. Does not
-include the algorithm time.
+The data transfer latency accounts for a significant part if the total latency, 
+as can be seen in Fig.~\ref{fig:transferOnly}: about 20-25\% of total latency  is due to moving the data to and from the GPU.
+
+\section{Conclusions}
 
@@
 
+\section*{References}
+
 %% CHEP2013 recommended style for bibliography
-\bibliographystyle{iopart-num}
+%\bibliographystyle{iopart-num}
 
-\bibliography{gpu}
+%\bibliography{gpu}
 
@@
 \bibitem{bib_nvidia} \url{http://www.nvidia.com}
-\bibitem{bib_cuda}\url{http://www.nvidia.com/object/cuda_home_new.html}
-\bibitem{bib_OpenCL}\url{http://www.khronos.org/opencl/}
-%\bibitem{bib_GPU_TIPP2011} W.~Ketchum \textit{et al.}, \emph{Performance Study of GPUs in Real-Time Trigger Applications for HEP Experiments},  Physics Procedia, Volume 37, 2012, 1965-1972.
-%\bibitem{bib_Apenet} R.~Ammendola \textit{et al.},% \emph{APEnet+: a 3D toroidal network enabling Petaflops scale Lattice QCD simulations on commodity clusters}  PoS(Lattice 2010)022.
-%\bibitem{bib_GPUDirect} \url{https://developer.nvidia.com/gpudirect}
-%\bibitem{bib_SVT1} B.~Ashmanskas \emph{et al.}, \emph{The CDF Silicon    Vertex Trigger}, Nucl.~Instrum.~Meth.
+
+\bibitem{bib_tipp2011} W.~Ketchum \textit{et al.}, \emph{Performance Study of GPUs in Real-Time Trigger Applications for HEP Experiments},  Physics Procedia, Volume 37, 2012, 1965-1972.
+
+\bibitem{bib_nss2012} P.~Wittich \textit{et al.}, \emph{Applications of GPUs to online track reconstruction in HEP experiments}, published in Nuclear Science Symposium and Medical Imaging Conference (NSS/MIC), 2012 IEEE, \href{http://dx.doi.org/10.1109/NSSMIC.2012.6551422}{10.1109/NSSMIC.2012.6551422}
+
+\bibitem{bib_SVT1} B.~Ashmanskas \emph{et al.}, \emph{The CDF Silicon    Vertex Trigger}, Nucl.~Instrum.~Meth.
 A \textbf{518} (2004) 532–536.
 \href{http://dx.doi.org/10.1016/j.nima.2003.11.078}{doi:10.1016/j.nima.2003.11.078}
-%\bibitem{bib_SVT2} J.~A.~Adelman \emph{et al.}, \emph{The Silicon    Vertex Trigger upgrade at CDF}, Nucl.~Instrum.~Meth. A \textbf{572} (2007)  361–364. 
-%\href{http://dx.doi.org/10.1016/j.nima.2006.10.383}{doi:10.1016/j.nima.2006.10.383}
-%\bibitem{bib_SVT3} S. Amerio \emph{et al.}, \emph{The GigaFitter: Performance at CDF and perspectives for future applications}, J.Phys.Conf.Ser. 219 (2010) 022001.
+
+\bibitem{bib_cuda} \url{http://www.nvidia.com/object/cuda_home_new.html}
+
+\bibitem{bib_OpenCL} \url{http://www.khronos.org/opencl/}
+
+%\bibitem{bib_GPU_TIPP2011} Physics Procedia, Volume 37, 2012, 1965-1972.
+\bibitem{bib_Apenet} R.~Ammendola \textit{et al.}, \emph{APEnet+: a 3D toroidal network enabling Petaflops scale Lattice QCD simulations on commodity clusters}  PoS(Lattice 2010)022.
+
+\bibitem{bib_GPUDirect} \url{https://developer.nvidia.com/gpudirect}
+
+\bibitem{bib_SVT2} J.~A.~Adelman \emph{et al.}, \emph{The Silicon    Vertex Trigger upgrade at CDF}, Nucl.~Instrum.~Meth. A \textbf{572} (2007)  361–364.
+
+\bibitem{bib_SVT3} S. Amerio \emph{et al.}, \emph{The GigaFitter: Performance at CDF and perspectives for future applications}, J.Phys.Conf.Ser. 219 (2010) 022001.
 

### 2013-10-16 papers/2013/CHEP/paper.tex
fix refs

@@
 \begin{abstract}
-One of the most important issues facing particle physics experiments at hadron 
-colliders is real-time selection of interesting events for offline storage. 
-Collision frequencies do not allow all events to be written to tape for 
-offline analysis, and in most cases, only a small fraction can be saved. 
-Typical trigger systems use commercial computers in the final stage of 
-processing. Much of the effort is focused on understanding the latency for 
-trigger systems. In this paper we describe updates to a previous study of 
-latencies in GPU for potential trigger applications, where we measured the 
-latency to transfer data to/from the GPU, exploring the timing of different 
-I/O technologies on different GPU models.  Those studies, where a simplified 
-track fitting algorithm was parallelized and run on a GPU, show that latencies 
-of few tens of microseconds can be achieved to transfer and process packets of 
-4 kB of data, combining the modern Infiniband data transfer technology with 
-direct access to GPU memory allowed by NVIDIA GPUDirect utilities.  We now 
-have expanded our latency studies to include other multi-core systems 
-(Intel Xeon Phi in addition to NVIDIA GPUs). We also discuss the implementation 
-of a scaled-up version of the algorithm 
-used at CDF for online track reconstruction - the SVT algorithm - as a 
-realistic test-case for low-latency trigger systems using new computing 
-architectures for LHC experiment.
+  One of the most important issues facing particle physics experiments
+  at hadron colliders is real-time selection of interesting events for
+  offline storage.  Collision frequencies do not allow all events to
+  be written to tape for offline analysis, and in most cases, only a
+  small fraction can be saved.  Typical trigger systems use commercial
+  computers in the final stage of processing. Much of the effort is
+  focused on understanding the latency for trigger systems. In this
+  paper we describe updates to a previous study of latencies in GPU
+  for potential trigger applications, where we measured the latency to
+  transfer data to/from the GPU, exploring the timing of different I/O
+  technologies on different GPU models.  Those studies, where a
+  simplified track fitting algorithm was parallelized and run on a
+  GPU, show that latencies of few tens of microseconds can be achieved
+  to transfer and process packets of 4 kB of data, combining the
+  modern Infiniband data transfer technology with direct access to GPU
+  memory allowed by NVIDIA GPUDirect utilities.  We now have expanded
+  our latency studies to include other multi-core systems (Intel Xeon
+  Phi in addition to NVIDIA GPUs). We also discuss the implementation
+  of a scaled-up version of the algorithm used at CDF for online track
+  reconstruction - the SVT algorithm - as a realistic test-case for
+  low-latency trigger systems using new computing architectures for
+  LHC experiment.
 
@@
 We showed initial studies on GPU performance in low-latency
-environments ($\sim$100\,$\mu$s) in a previous paper~\cite{bib_tipp2011,bib_nss2012}.  In this paper we extend those studies measuring the
+environments ($\sim$100\,$\mu$s) in a previous paper~\cite{TIPP2011,NSS2012}.  In this paper we extend those studies measuring the
 timing performance of different GPU cards in different data I/O
@@
 the fast track fitting algorithm of the Silicon Vertex Tracker (SVT)
-system at CDF~\cite{bib_SVT1}. 
+system at CDF~\cite{SVT1}. 
 %We also compare the performance of GPU
@@
 \section{SVT track fitting algorithm}
-The Silicon Vertex Trigger (SVT)~\cite{bib_SVT1}~\cite{bib_SVT2} is a 
+The Silicon Vertex Trigger (SVT)~\cite{SVT1,SVT2} is a 
 track reconstruction processor used at 
@@
 This algorithm uses a linearized approximation to track-fitting as 
-implemented in hardware (described in greater detail in~\cite{bib_SVT3}). 
+implemented in hardware (described in greater detail in~\cite{SVT3}). 
 With the SVT approach, the determination of the track parameters 
@@
 single combination of hits. 
-To run on MIC, we adopted the so-called \textit{embarassingly parallel} 
+To run on MIC, we adopted the so-called \textit{embarrassingly parallel} 
 approach and used pragma OpenMP for statements to unroll the external loop 
@@
 Apenet+ is an FPGA-based PCIe board supporting peer-to-peer communication with 
-Tesla and Kepler cards~\cite{bib_Apenet}.
+Tesla and Kepler cards~\cite{apenet2010}.
 Two nodes of this cluster are used to measure data transfer latency,
@@
 %% CHEP2013 recommended style for bibliography
-%\bibliographystyle{iopart-num}
+\bibliographystyle{iopart-num}
 
-%\bibliography{gpu}
+\bibliography{gpu}
 
-\begin{thebibliography}{99}
-\bibitem{bib_nvidia} \url{http://www.nvidia.com}
+% \bibitem{bib_cuda} \url{http://www.nvidia.com/object/cuda_home_new.html}
 
-\bibitem{bib_tipp2011} W.~Ketchum \textit{et al.}, \emph{Performance Study of GPUs in Real-Time Trigger Applications for HEP Experiments},  Physics Procedia, Volume 37, 2012, 1965-1972.
+% \bibitem{bib_OpenCL} \url{http://www.khronos.org/opencl/}
 
-\bibitem{bib_nss2012} P.~Wittich \textit{et al.}, \emph{Applications of GPUs to online track reconstruction in HEP experiments}, published in Nuclear Science Symposium and Medical Imaging Conference (NSS/MIC), 2012 IEEE, \href{http://dx.doi.org/10.1109/NSSMIC.2012.6551422}{10.1109/NSSMIC.2012.6551422}
+% %\bibitem{bib_GPU_TIPP2011} Physics Procedia, Volume 37, 2012, 1965-1972.
+% \bibitem{bib_Apenet} R.~Ammendola \textit{et al.}, \emph{APEnet+: a 3D toroidal network enabling Petaflops scale Lattice QCD simulations on commodity clusters}  PoS(Lattice 2010)022.
 
-\bibitem{bib_SVT1} B.~Ashmanskas \emph{et al.}, \emph{The CDF Silicon    Vertex Trigger}, Nucl.~Instrum.~Meth.
-A \textbf{518} (2004) 532–536.
-\href{http://dx.doi.org/10.1016/j.nima.2003.11.078}{doi:10.1016/j.nima.2003.11.078}
-
-\bibitem{bib_cuda} \url{http://www.nvidia.com/object/cuda_home_new.html}
-
-\bibitem{bib_OpenCL} \url{http://www.khronos.org/opencl/}
-
-%\bibitem{bib_GPU_TIPP2011} Physics Procedia, Volume 37, 2012, 1965-1972.
-\bibitem{bib_Apenet} R.~Ammendola \textit{et al.}, \emph{APEnet+: a 3D toroidal network enabling Petaflops scale Lattice QCD simulations on commodity clusters}  PoS(Lattice 2010)022.
-
-\bibitem{bib_GPUDirect} \url{https://developer.nvidia.com/gpudirect}
-
-\bibitem{bib_SVT2} J.~A.~Adelman \emph{et al.}, \emph{The Silicon    Vertex Trigger upgrade at CDF}, Nucl.~Instrum.~Meth. A \textbf{572} (2007)  361–364.
-
-\bibitem{bib_SVT3} S. Amerio \emph{et al.}, \emph{The GigaFitter: Performance at CDF and perspectives for future applications}, J.Phys.Conf.Ser. 219 (2010) 022001.
-
-\end{thebibliography}
+% \bibitem{bib_GPUDirect} \url{https://developer.nvidia.com/gpudirect}
 

### 2013-10-16 src_opencl/svt_gpu_opencl.cpp
changes to run through whole scheme...everything compiles and runs without crashing, but output in fout_arrays is junk right now

@@ -135,6 +135,7 @@ int main(int argc, char* argv[]) {
   struct fep_arrays *fep_dev = new fep_arrays;
   struct extra_data *edata_dev = new extra_data;
   struct fit_arrays *fit_dev = new fit_arrays;
+  struct fout_arrays *fout_dev = new fout_arrays;
 
   bool is_null=false;
   if(evt==NULL) is_null=true;;
@@ -329,14 +330,28 @@ int main(int argc, char* argv[]) {
     cl::Kernel kernel_kFit(fit_program, kernelFunc_kFit.c_str(), &err);
     CL_HELPERFUNCS::checkErr(err, "Kernel::Kernel()");
     std::cout << __FANCY__ << "Entry point set to " << kernelFunc_kFit << std::endl;
-    /*
+
     //====================================
     //set kernel entry point
     std::string kernelFunc_fit_format = "gf_fit_format_GPU";
-    cl::Kernel kernel_fep_set(program, kernelFunc_fep_set.c_str(), &err);
+    cl::Kernel kernel_fit_format(fit_program, kernelFunc_fit_format.c_str(), &err);
     CL_HELPERFUNCS::checkErr(err, "Kernel::Kernel()");
-    std::cout << __FANCY__ << "Entry point set to " << kernelFunc_fep_set << std::endl;
-    */
+    std::cout << __FANCY__ << "Entry point set to " << kernelFunc_fit_format << std::endl;
+
+    //====================================
+    //set kernel entry point
+    std::string kernelFunc_comparator = "gf_comparator_GPU";
+    cl::Kernel kernel_comparator(fit_program, kernelFunc_comparator.c_str(), &err);
+    CL_HELPERFUNCS::checkErr(err, "Kernel::Kernel()");
+    std::cout << __FANCY__ << "Entry point set to " << kernelFunc_comparator << std::endl;
+
+    //====================================
+    //set kernel entry point
+    std::string kernelFunc_compute_eeword = "gf_compute_eeword_GPU";
+    cl::Kernel kernel_compute_eeword(fit_program, kernelFunc_compute_eeword.c_str(), &err);
[…]

### 2013-10-16 papers/2013/CHEP/paper.tex
start changing abstract to focus on what's new

@@
 \begin{abstract}
-  One of the most important issues facing particle physics experiments
-  at hadron colliders is real-time selection of interesting events for
-  offline storage.  Collision frequencies do not allow all events to
-  be written to tape for offline analysis, and in most cases, only a
-  small fraction can be saved.  Typical trigger systems use commercial
-  computers in the final stage of processing. Much of the effort is
-  focused on understanding the latency for trigger systems. In this
-  paper we describe updates to a previous study of latencies in GPU
-  for potential trigger applications, where we measured the latency to
-  transfer data to/from the GPU, exploring the timing of different I/O
-  technologies on different GPU models.  Those studies, where a
-  simplified track fitting algorithm was parallelized and run on a
-  GPU, show that latencies of few tens of microseconds can be achieved
-  to transfer and process packets of 4 kB of data, combining the
-  modern Infiniband data transfer technology with direct access to GPU
-  memory allowed by NVIDIA GPUDirect utilities.  We now have expanded
-  our latency studies to include other multi-core systems (Intel Xeon
-  Phi in addition to NVIDIA GPUs). We also discuss the implementation
-  of a scaled-up version of the algorithm used at CDF for online track
-  reconstruction - the SVT algorithm - as a realistic test-case for
-  low-latency trigger systems using new computing architectures for
-  LHC experiment.
+  In this paper we describe updates to a previous study of latencies
+  in GPU for potential trigger applications, where we measured the
+  latency to transfer data to/from the GPU, exploring the timing of
+  different I/O technologies on different GPU models.  Those studies,
+  where a simplified track fitting algorithm was parallelized and run
+  on a GPU, show that latencies of few tens of microseconds can be
+  achieved to transfer and process packets of 4 kB of data, combining
+  the modern Infiniband data transfer technology with direct access to
+  GPU memory allowed by NVIDIA GPUDirect utilities.  We now have
+  expanded our latency studies to include other multi-core systems
+  (Intel Xeon Phi in addition to NVIDIA GPUs). We also discuss the
+  implementation of a scaled-up version of the algorithm used at CDF
+  for online track reconstruction - the SVT algorithm - as a realistic
+  test-case for low-latency trigger systems using new computing
+  architectures for LHC experiment.
 

### 2013-10-16 src_opencl/svt_gpu_opencl.cpp
adding in initialization of arrays on gpu, which makes fout_ntrks sensible but still not the same as on the cpu...

@@ -252,6 +252,34 @@ int main(int argc, char* argv[]) {
 
     //====================================
     //build program source
+    std::string kernel_file_init = "gf_init.cl";
+    std::ifstream file_init((kernel_file_init).c_str());
+    CL_HELPERFUNCS::checkErr(file_init.is_open() ? CL_SUCCESS:-1, (kernel_file_init).c_str());
+    
+    
+    //TODO-RAR add build options
+    std::string prog_init(std::istreambuf_iterator<char>(file_init),
+		     (std::istreambuf_iterator<char>()));
+    std::string buildOptions_init;
+    { // create preprocessor defines for the kernel
+      char buf[256]; 
+      //sprintf(buf,"-cl-mad-enable -D DATAWORD=%s ", dataType.getType().c_str());
+      sprintf(buf,"-cl-mad-enable -I./");
+      buildOptions_init = buf;
+      std::cout << __FANCY__ << "buildOptions = " << buildOptions_init << std::endl;
+    }
+    
+    cl::Program::Sources source_init(1,std::make_pair(prog_init.c_str(), prog_init.length()));
+    
+    cl::Program program_init(context, source_init);
+    
+    err = program_init.build(*(deviceList[plat_i]),buildOptions_init.c_str());
+	      
+    CL_HELPERFUNCS::displayBuildLog(program_init, device);
+    CL_HELPERFUNCS::checkErr(err, "Build::Build()");
+
+    //====================================
+    //build program source
     std::string kernel_file = "gf_fep.cl";
     std::ifstream file((kernel_file).c_str());
     CL_HELPERFUNCS::checkErr(file.is_open() ? CL_SUCCESS:-1, (kernel_file).c_str());
@@ -280,6 +308,13 @@ int main(int argc, char* argv[]) {
     
     
     
+    //====================================
+    //set kernel entry point
+    std::string kernelFunc_init = "init_arrays_GPU";
+    cl::Kernel kernel_init(program_init, kernelFunc_init.c_str(), &err);
[…]

### 2013-10-16 src_opencl/svt_gpu_opencl.cpp
fix to the initialization call, still some differences between this and cpu running

@@ -464,23 +464,36 @@ int main(int argc, char* argv[]) {
 			    &err);
     CL_HELPERFUNCS::checkErr(err, "Buffer::Buffer() OUT");
 
-    err = kernel_init.setArg(0, evt_CL);
-    err = kernel_init.setArg(1, fout_dev_CL);
+    err = kernel_init.setArg(1, evt_CL);
+    err = kernel_init.setArg(0, fout_dev_CL);
     CL_HELPERFUNCS::checkErr(err, "Kernel::setArg()");
 
     err = queue.enqueueNDRangeKernel(
 				     kernel_init,
 				     cl::NullRange,
-				     cl::NDRange(NEVTS*MAXROAD),
+				     cl::NDRange(NEVTS*(NSVX_PLANE+1),MAXROAD),
 				     cl::NDRange(NSVX_PLANE+1,1),
 				     NULL,
 				     &event);
     CL_HELPERFUNCS::checkErr(err, "ComamndQueue::enqueueNDRangeKernel()");
 
     event.wait();
+    /*
+    err = queue.enqueueReadBuffer(
+				  fout_dev_CL,
+				  CL_TRUE,
+				  0,
+				  sizeof(fout_arrays),
+				  fout_dev);
+    CL_HELPERFUNCS::checkErr(err, "CommandQueue::enqueueReadBuffer()");
 
+    err = queue.finish();
+    CL_HELPERFUNCS::checkErr(err, "ComamndQueue::clFinish2()");
 
-
+    for(int ie=0; ie<NEVTS; ie++){
+      printf("\nEvent %d, evt_nroads = %d, fep_nroads=%d, fit_err_sum=%d, fout_ntrks=%d",ie,evt->evt_nroads[ie],fep_dev->fep_nroads[ie],fit_dev->fit_err_sum[ie],fout_dev->fout_ntrks[ie]);
+    }
+    */
     gettimeofday(&ptBegin, NULL);
     gf_fep_unpack_evt(evt, k, data_send);
     printf("Total events %d\n\n",evt->totEvts);
@@ -654,7 +667,7 @@ int main(int argc, char* argv[]) {
           ((ptEnd.tv_usec + 1000000 * ptEnd.tv_sec) - (ptBegin.tv_usec + 1000000 * ptBegin.tv_sec))/1000.0);
 
     for(int ie=0; ie<NEVTS; ie++){
[…]

### 2013-10-17 src_opencl/svt_gpu_opencl.cpp
fix some logical bugs

@@ -22,6 +22,31 @@ void svt_GPU(tf_arrays_t tf, unsigned int *data_in, int n_words) {
 
 }
 
+
+void set_outcable_fout(fout_arrays* fout_dev, int totEvts, unsigned int *&data_rec, int &ow) {
+
+  svtsim_cable_t *out;
+  out = svtsim_cable_new();
+
+  svtsim_cable_copywords(out, 0, 0);
+
+  for (int ie=0; ie < totEvts; ie++) {
+    for (int nt=0; nt < fout_dev->fout_ntrks[ie]; nt++) {
+      svtsim_cable_addwords(out, fout_dev->fout_gfword[ie][nt], NTFWORDS);
+    }
+    svtsim_cable_addword(out, fout_dev->fout_ee_word[ie]);
+  }
+
+  ow = out->ndata;
+
+  for (int i=0; i < ow ; i++) {
+    data_rec[i] = out->data[i];
+  }
+
+//  svtsim_free(out);
+}
+
+
 void set_outcable(tf_arrays_t tf) {
 
   svtsim_cable_copywords(tf->out, 0, 0);
@@ -128,7 +153,10 @@ int main(int argc, char* argv[]) {
   tf_arrays_t tf;
 
   gf_init(&tf);
-  svtsim_fconread(tf);
+
+  unsigned int *data_rec = (unsigned int*)malloc(2500000*sizeof(unsigned));
+  int ow;
+
 
   struct evt_arrays *evt = new evt_arrays;
   int totEvts;
@@ -137,8 +165,12 @@ int main(int argc, char* argv[]) {
   struct fit_arrays *fit_dev = new fit_arrays;
   struct fout_arrays *fout_dev = new fout_arrays;
 
+  svtsim_fconread(tf, edata_dev);
+
+  free(tf);
+
   bool is_null=false;
-  if(evt==NULL) is_null=true;;
+  if(evt==NULL) is_null=true;
 
   printf("Is evt_arrays ptr null? bool=%d",is_null);
 
@@ -148,7 +180,7 @@ int main(int argc, char* argv[]) {
 
   if ( strcmp(where,"cpu") == 0 ) { // CPU
     printf("Start work on CPU..... \n");
-    
+   /* 
     gettimeofday(&ptBegin, NULL);
     gf_fep_unpack(tf, k, data_send);
     gettimeofday(&ptEnd, NULL);
[…]

### 2013-10-17 src_opencl/svt_gpu_opencl.h

@@ -5,7 +5,9 @@ extern "C" {
 
   int gf_init(tf_arrays_t* ptr_tf);
   int gf_init_evt(evt_arrays_t* ptr_evt);
-  int svtsim_fconread(tf_arrays_t tf);
+  //int svtsim_fconread(tf_arrays_t tf);
+  int svtsim_fconread(tf_arrays_t tf, struct extra_data* edata);
+
   int gf_fep_unpack(tf_arrays_t tf, int n_words_in, void* data);
   int gf_fep_unpack_evt(evt_arrays_t tf, int n_words_in, void* data);
   int gf_fep_comb(tf_arrays_t tf);

### 2013-10-17 src_opencl/svtsim_functions.c

@@ -352,22 +352,8 @@ svtsim_cable_t * svtsim_cable_new(void) {
 }
  
 
-
- 
- /*
-  * Map layerMask, lcMask into fit block (0..31) for fully general case
-  * in which all fit constants are available.  Note that blocks 25..30
-  * are set up for some exotic cases that may never really occur.  Note
-  * also that before 2002-12-15 or so, only fit block 0 was actually
-  * used, and that as of 2002-12-15, no fit constants yet exist (or are
-  * planned) that make use of the LC bits.
- 
-  * Note even further that the upgraded TF has NO room for fit constants
-  * that use the LC bits, but we do use them 
-  * to select the constants when we have 5/5 roads
-  */
  int
- svtsim_whichFit_full(tf_arrays_t tf, int zin, int layerMask, int lcMask)
+ svtsim_whichFit_full(int zin, int layerMask, int lcMask)
  {
    /*
     * Key (*=LC, 5=XFT):
@@ -488,13 +474,11 @@ svtsim_cable_t * svtsim_cable_new(void) {
      
    }
  }
- 
- /*
-  * Map layerMask, lcMask into fit block (0..31), allowing for
-  * degeneracy in TF mkaddr map
-  */
- int
- svtsim_whichFit(tf_arrays_t tf, int zin, int layerMask, int lcMask)
+
+
+
+  int
+ svtsim_whichFit(struct extra_data* edata, int zin, int layerMask, int lcMask)
  {
 #ifdef DEBUG_SVT 
    printf("in svtsim_whichFit: zin = %d, layerMask = %x, lcMask = %x\n", zin, layerMask, lcMask);
@@ -502,8 +486,8 @@ svtsim_cable_t * svtsim_cable_new(void) {
 
    int which0 = 0, which = 0;
    if (zin<0 || zin>=SVTSIM_NBAR) zin = 0;
-   which0 = svtsim_whichFit_full(tf, zin, layerMask, lcMask);
-   which = tf->whichFit[zin][which0];
+   which0 = svtsim_whichFit_full(zin, layerMask, lcMask);
+   which = edata->whichFit[zin][which0];
 
 #ifdef DEBUG_SVT 
    printf("in svtsim_whichFit: which0 = %d, which = %x\n", which0, which);
[…]

### 2013-10-17 src_opencl/svt_gpu_opencl.cpp
incorporating alessios updates

@@ -484,7 +484,7 @@ int main(int argc, char* argv[]) {
 			  NULL,
 			  &err);
     CL_HELPERFUNCS::checkErr(err, "Buffer::Buffer() OUT");
-    
+        
     cl::Buffer fout_dev_CL(
 			    context,
 			    CL_HELPERFUNCS::isDeviceTypeGPU(&deviceList,plat_i,dev_i) ?
@@ -509,6 +509,17 @@ int main(int argc, char* argv[]) {
     CL_HELPERFUNCS::checkErr(err, "ComamndQueue::enqueueNDRangeKernel(init)");
 
     event.wait();
+
+    err = queue.enqueueWriteBuffer(
+				   edata_dev_CL,
+				   CL_TRUE,
+				   0,
+				   sizeof(extra_data),
+				   edata_dev);
+    CL_HELPERFUNCS::checkErr(err, "ComamndQueue::enqueueWriteBuffer()");
+
+    event.wait();
+
     /*
     err = queue.enqueueReadBuffer(
 				  fout_dev_CL,
@@ -568,7 +579,8 @@ int main(int argc, char* argv[]) {
 				   sizeof(evt_arrays),
 				   evt);
     CL_HELPERFUNCS::checkErr(err, "CommandQueue::enqueueWriteBuffer()");
- 
+
+    /* 
     err = queue.enqueueWriteBuffer(
 				  edata_dev_CL,
 				  CL_TRUE,
@@ -576,7 +588,7 @@ int main(int argc, char* argv[]) {
 				  sizeof(extra_data),
 				  edata_dev);
     CL_HELPERFUNCS::checkErr(err, "CommandQueue::enqueueWriteBuffer()");
-
+    */
    
     err = queue.enqueueNDRangeKernel(
 				     kernel_fep_comb,

### 2013-10-17 src_opencl/svt_gpu_opencl.h

@@ -6,8 +6,7 @@ extern "C" {
   int gf_init(tf_arrays_t* ptr_tf);
   int gf_init_evt(evt_arrays_t* ptr_evt);
   //int svtsim_fconread(tf_arrays_t tf);
-  int svtsim_fconread(tf_arrays_t tf, struct extra_data* edata);
-
+  int svtsim_fconread(tf_arrays_t tf, struct extra_data* edata_dev);
   int gf_fep_unpack(tf_arrays_t tf, int n_words_in, void* data);
   int gf_fep_unpack_evt(evt_arrays_t tf, int n_words_in, void* data);
   int gf_fep_comb(tf_arrays_t tf);

### 2013-10-17 src_opencl/svt_utils_opencl.h

@@ -87,3 +87,40 @@ struct fout_arrays {
 
 };
 
+struct gf_data_arrays {
+
+ //output "cable"
+  svtsim_cable_t *out;
+  
+  //gf_memory
+  //int wedge[NEVTS]; //handled in extra_data
+  unsigned int *mem_mkaddr[NEVTS];
+  short mem_coeff[NFITTER][SVTNHITS][MAXCOE_VSIZE]; 
+  int mem_nintcp;
+  short (*intcp)[NFITTER];
+  int mem_fitsft[NFITTER][NSHIFTS]; 
+  /* int chi2[NCHI][MAXCHI2A]; */
+  int minhits; /* The minimum number of hits that we require 
+		  (including XFT hit)*/
+  int chi2cut;
+  int svt_emsk;
+  int gf_emsk;
+  int cdf_emsk;
+  int eoe_emsk; /* MASK for the errors */
+
+  //int whichFit[SVTSIM_NBAR][FITBLOCK];/* handles TF mkaddr degeneracy */ //handled in extra_data
+  int ifitpar[NFITPAR][DIMSPA+1][SVTSIM_NBAR][FITBLOCK];       /* TF coefficients, P0s */
+  unsigned long lfitpar[NFITPAR][DIMSPA+1][SVTSIM_NBAR][FITBLOCK];/* full-precision coeffs, P0s */ //handled in extra_data
+  //long long int lfitparfcon[NFITPAR][DIMSPA+1][SVTSIM_NBAR][FITBLOCK]; /* full-precision coeffs, P0s as read from fcon files SA*/
+  float gcon[NFITPAR][DIMSPA+1][SVTSIM_NBAR][FITBLOCK];     /* [fdc012][I0123PC][z][whichfit] */
+  int xftphi_shiftbits[NFITPAR];/* TF bit shifts */
+  int xftcrv_shiftbits[NFITPAR];
+  int result_shiftbits[NFITPAR];
+  int lMap[SVTSIM_NBAR][SVTSIM_NPL];/* map physical layer => cable layer */
+  int oX[SVTSIM_NBAR], oY[SVTSIM_NBAR];/* origin used for fcon */
+  float phiUnit, dvxUnit, crvUnit;/* units for fit parameters */
+  float k0Unit, k1Unit, k2Unit;/* units for constraints */
+  int dphiNumer, dphiDenom;/* dphi(wedge) = 2pi*N/D */
+  int mkaddrBogusValue;
+
+};

### 2013-10-17 src_opencl/svtsim_functions.c

@@ -527,7 +527,7 @@ svtsim_cable_t * svtsim_cable_new(void) {
    return 0;
  }
  
- int svtsim_fconread(tf_arrays_t tf, struct extra_data* edata)
+int svtsim_fconread(tf_arrays_t tf, struct extra_data *edata_dev)
  {
    int i = 0, j = 0, k = 0, which = 0;
    char fconFile[SVTSIM_NBAR][100];
@@ -542,7 +542,7 @@ svtsim_cable_t * svtsim_cable_new(void) {
    tf->dphiNumer = WEDGE;
    
    for (i = 0; i<NEVTS; i++) {
-     edata->wedge[i]=WEDGE;
+     edata_dev->wedge[i]=WEDGE;
    }
 
    for (i = 0; i<SVTSIM_NBAR; i++) {
@@ -579,7 +579,7 @@ svtsim_cable_t * svtsim_cable_new(void) {
       /*
        * Initialize all whichFit entries to bogus values
        */
-      for (i = 0; i<SVTSIM_NEL(edata->whichFit[k]); i++) edata->whichFit[k][i] = -1;
+      for (i = 0; i<SVTSIM_NEL(edata_dev->whichFit[k]); i++) edata_dev->whichFit[k][i] = -1;
  
       /*
        * Open .fcon
@@ -642,7 +642,7 @@ svtsim_cable_t * svtsim_cable_new(void) {
         }
         if (bogusLayers) continue;
         which = svtsim_whichFit_full(k, layerMask, 0);
-        edata->whichFit[k][which] = which;
+        edata_dev->whichFit[k][which] = which;
         if(0) printf("w=%d z=%d: found detLayers=%s => "
  		    "amLayers=%s mask=%x which=%d\n", 
  		    WEDGE, k, str, layers, layerMask, which);
@@ -692,15 +692,15 @@ svtsim_cable_t * svtsim_cable_new(void) {
   	    tf->lfitpar[l][j][k][which] = 
   	      (tf->lfitpar[l][j][k][which] >> h[j]);
   	    /* fill lfitparfcon array with constants as read from fcon file SA*/
-  	    edata->lfitparfcon[l][j][k][which] = fc[j];
+  	    edata_dev->lfitparfcon[l][j][k][which] = fc[j];
   #ifdef DEBUG_READFCON
[…]

### 2013-10-18 src_opencl/svt_gpu_opencl.cpp
updates to clean things up a bit

@@ -180,7 +180,7 @@ int main(int argc, char* argv[]) {
 
   if ( strcmp(where,"cpu") == 0 ) { // CPU
     printf("Start work on CPU..... \n");
-   /* 
+
     gettimeofday(&ptBegin, NULL);
     gf_fep_unpack(tf, k, data_send);
     gettimeofday(&ptEnd, NULL);
@@ -211,9 +211,12 @@ int main(int argc, char* argv[]) {
     gettimeofday(&ptEnd, NULL);
     printf("Time to CPU fit: %.3f ms\n",
           ((ptEnd.tv_usec + 1000000 * ptEnd.tv_sec) - (ptBegin.tv_usec + 1000000 * ptBegin.tv_sec))/1000.0);
-
+    /*
+    for(int ie=0; ie<NEVTS; ie++)
+      printf("Event %d, evt_nroads = %d, fep_nroads=%d, fit_err_sum=%d, fout_ntrks=%d, fout_parity=%d\n",ie,tf->evt_nroads[ie],tf->fep_nroads[ie],tf->fit_err_sum[ie],tf->fout_ntrks[ie],tf->fout_parity[ie]);
+    */
     printf(".... fits %d events! \n", tf->totEvts);
-*/
+
   } else { // GPU
 
     HELPERFUNCS::TemplateDataType<DATAWORD> dataType;
@@ -438,8 +441,8 @@ int main(int argc, char* argv[]) {
     std::cout << "device type " << CL_HELPERFUNCS::isDeviceTypeGPU(&deviceList,plat_i,dev_i) << std::endl;
 
 
-    printf("Size of arrays is evt_arrays=%d, fep_arrays=%d, fit_arrays=%d, extra_data=%d\n",
-	   sizeof(evt_arrays),sizeof(fep_arrays),sizeof(fit_arrays), sizeof(extra_data));
+    //printf("Size of arrays is evt_arrays=%d, fep_arrays=%d, fit_arrays=%d, extra_data=%d\n",
+    //	   sizeof(evt_arrays),sizeof(fep_arrays),sizeof(fit_arrays), sizeof(extra_data));
 
     
     cl::Buffer evt_CL(
@@ -520,29 +523,6 @@ int main(int argc, char* argv[]) {
 
     event.wait();
 
-    /*
-    err = queue.enqueueReadBuffer(
-				  fout_dev_CL,
-				  CL_TRUE,
-				  0,
[…]

### 2013-10-18 src_opencl/svtsim_defines.h

@@ -23,7 +23,7 @@
 
 
 //GF
-#define NEVTS 100
+#define NEVTS 10
 #define NROADS 64
 #define EVT 0 //all words from the input file are considered belonging to the same event
 

### 2013-10-18 src_opencl/svtsim_functions.c

@@ -588,9 +588,9 @@ int svtsim_fconread(tf_arrays_t tf, struct extra_data *edata_dev)
       FILE *fp = 0;
       fp = fopen(fcf[k],"r");
       //fp = blobopen(fcf[k], "r");
-#ifdef DEBUG_READFCON
+      //#ifdef DEBUG_READFCON
       printf("fconread: opening %s for input, barrel %d\n", fcf[k],k);
-#endif
+      //#endif
       if (!fp) {
         fprintf(stderr, "svtsim: error opening %s for input\n", fcf[k]);
         return -1;

### 2013-10-20 papers/2013/CHEP/paper.tex
add conclusion, more captions, some text edits and new bib stuff

@@
 \usepackage{url}
+\usepackage{xspace}
 \usepackage{iopams}
@@
       bookmarksopen=true]{hyperref}
+
+\providecommand{\fixme}[1]{\xspace{\sffamily{[\bfseries{}FIXME:} #1]}}
 
@@
   expanded our latency studies to include other multi-core systems
-  (Intel Xeon Phi in addition to NVIDIA GPUs). We also discuss the
+  (Intel Xeon Phi/MIC in addition to NVIDIA GPUs). We also discuss the
   implementation of a scaled-up version of the algorithm used at CDF
@@
 high-level programming architectures based on C/C++ such as
-\textsc{cuda} and \textsc{opencl} make these devices more accessible
-to the general user.  
-The ultimate goal of this study is to investigate the strengths and weaknesses
-of GPUs when applied in a low latency environment, with particular
+\textsc{cuda}~\cite{bib_cuda} and \textsc{opencl}~\cite{bib_opencl}
+make these devices more accessible to the general user.  The ultimate
+goal of this study is to investigate the strengths and weaknesses of
+GPUs when applied in a low latency environment, with particular
 emphasis on the data transfer latency to/from the GPU and the
@@
 We showed initial studies on GPU performance in low-latency
-environments ($\sim$100\,$\mu$s) in a previous paper~\cite{TIPP2011,NSS2012}.  In this paper we extend those studies measuring the
-timing performance of different GPU cards in different data I/O
-environments. The algorithm run on the GPU is a complete version of
-the fast track fitting algorithm of the Silicon Vertex Tracker (SVT)
-system at CDF~\cite{SVT1}. 
+environments ($\sim 100~\mu$s) in a previous
+paper~\cite{TIPP2011,NSS2012}.  In this paper we extend those studies
+measuring the timing performance of different GPU cards in different
+data I/O environments. The algorithm run on the GPU is a complete
+version of the fast track fitting algorithm of the Silicon Vertex
+Tracker (SVT) system at CDF~\cite{SVT1}.
 %We also compare the performance of GPU
@@
 to the existing algorithm. Starting with a serial algorithm
-implemented on a CPU, we test an \textit{embarrassingly parallel} algorithm
-on the Intel MIC environment. In this case each event is handled
-independently by a core on the accelerator, and the parallelization is
-achieved with only minor changes to the legacy code and is only
-possible in the Intel MIC environment. Next we consider an algorithm
-where we unroll three internal nested loops and run these in parallel
-on a GPU, using the \textsc{cuda}~\ref{bib_cuda} environment. This second 
-approach is programmatically more complicated and less trivial to implement. In
-neither case have we re-thought the basic algorithms or the data
-structures used. To achieve optimal performance, these steps would
-have to be taken.  As one might expect, the improvement from the first
-approach is rather modest, and the second approach shows larger
+implemented on a CPU, we test an \textit{embarrassingly parallel}
+algorithm on the Intel MIC environment. In this case each event is
+handled independently by a core on the accelerator, and the
+parallelization is achieved with only minor changes to the legacy code
+and is only possible in the Intel MIC environment. Next we consider an
+algorithm where we unroll three internal nested loops and run these in
+parallel on a GPU, using the \textsc{cuda} environment. This second
+approach is programmatically more complicated and less trivial to
+implement. In neither case have we re-thought the basic algorithms or
+the data structures used. To achieve optimal performance, these steps
+would have to be taken.  As one might expect, the improvement from the
+first approach is rather modest, and the second approach shows larger
 performance gains.
@@
 We ported the track fitting part to parallel devices, as it is well suited to 
-be parallelized.
+parallelization.
 
 \subsection{Code implementation}
-The starting point of of our studies is the SVT track fitting simulation code, 
-written in C language. SVT track fitting is divided into three main functions: 
-first the unpacking of input data and filling of all the necessary arrays, 
-followed by the computation of all possible combinations of hits in each road 
-and finally the fit on each combination. Three main loops are 
-present - on events, roads and hit combinations. 
-To be run on GPUs, the code has been ported to \textsc{cuda}: each step - 
-unpack, combine and track fit - is performed by a specific kernel;
-the three nested loops are unrolled so that each GPU thread processes a 
-single combination of hits. 
-To run on MIC, we adopted the so-called \textit{embarrassingly parallel} 
-approach and used pragma OpenMP for statements to unroll the external loop 
-on the events, so that each core processes a single event: the porting 
-requires much less effort compared to CUDA, but the level of parallelism is 
-limited.
+The starting point of of our studies is the SVT track fitting
+simulation code, written in the \textsc{c} language. SVT track fitting
+is divided into three main functions: the unpacking of input data and
+filling of all the necessary arrays, followed by the computation of
+all possible combinations of hits in each road and finally the fit of
+each combination. Three main loops are present - on events, roads and
+hit combinations.  To be run on GPUs, the code has been ported to
+\textsc{cuda}: each step - unpack, combine and track fit - is
+performed by a specific kernel; the three nested loops are unrolled so
+that each GPU thread processes a single combination of hits.  To run
+on MIC, we adopted the so-called \textit{embarrassingly parallel}
+approach and used \textsc{pragma OpenMP} for statements to unroll the external
+loop on the events, so that each core processes a single event: the
+porting requires much less effort compared to \textsc{cuda}, but the level of
+parallelism is limited.
 
@@
 receiver, processed on the GPU and sent back to the receiver (see
-Fig.~\ref{fig_data_flow}).  
+Fig.~\ref{fig:data_flow}).  
 The latency for a complete loop is measured on the transmitter using 
 standard C libraries.
-
-\begin{figure}[!h]
+\begin{figure}[tbph]
 \centering
@@
   role of a component in the trigger system. }
-\label{fig_data_flow}
+\label{fig:data_flow}
 \end{figure}
-
 In this setup, the transmitter can represent the detector, as
@@
     \centering
-    \includegraphics[width=0.9\linewidth]{figures/TimeComp_MIC.pdf}
+    \includegraphics[width=0.9\linewidth]{figures/TimeComp_MIC}
     \caption{Algorithm-only comparison for timing as a function of the
-      number of track fits. We compare timing on CPUs (serial), Xeon Phi
-      (embarrassingly parallel), and GPUs (fully parallel). }
+      number of track fits. We compare timing on CPUs (serial), Intel
+      MIC (embarrassingly parallel), and GPUs (fully parallel), in
+      blue, greed, and red, respectively. The GPUs exhibit the best
+      performance due to the full parallelization. }
     \label{fig:algo_only_timing}
@@
     \caption{Algorithm-only comparison for timing as a function of the
-      number of track fits: zoom in the low number of fits region.}
+      number of track fits: zoom in the low number of fits region. At
+      low number of fits, the CPU performs better, due to start-up
+      costs associated with data transfers.}
     \label{fig:algo_only_timing_zoom}
@@
    \centering
-   \includegraphics[width=0.9\linewidth]{figures/Speedup_MIC.pdf}
-   \caption{Speed-up with respect to a standard CPU (Intel Xeon E5630) .}
+   \includegraphics[width=0.9\linewidth]{figures/Speedup_MIC}
+   \caption{Speed-up with respect to a standard CPU (Intel Xeon
+     E5630). The speed-ups plateau after about one million fits for
+     the GPUs (in red) while the Intel MIC still shows a small slope
+     beyond that point. \fixme{is this true}}
    \label{fig:algo_only_speedup}
@@
 this part of the algorithm.  
-
-\begin{figure*}[!t]
+\begin{figure*}[tbp]
 \centering
-%\subfigure[]
-%{\label{fig:breakdown_CPU}
-%\includegraphics[width=0.4\linewidth]{figures/Cpui7_perc.pdf}}
-%\hspace{1mm}
 \subfigure[]
 {\label{fig:breakdown_MIC}
-\includegraphics[width=0.45\linewidth]{figures/Mic5_perc_nk.pdf}}
+\includegraphics[width=0.45\linewidth]{figures/Mic5_perc_nk}}
 \subfigure[]
 {\label{fig:breakdown_GPU}
-\includegraphics[width=0.45\linewidth]{figures/titan_perc.pdf}}
-\caption{Breakdown of computing time.}
+\includegraphics[width=0.45\linewidth]{figures/titan_perc}}
+\caption{Breakdown of computing time for MIC (a) and the GTX Titan GPU
+  (b). \fixme{we need a legend for these plots. Also can we make them
+    the same size and aspect ratio.}}
+\label{fig:breakdown}
 \end{figure*}
@@
 \subsection{Data transfer}
-The experimental setup described in Fig.~\ref{fig_data_flow} allows us to 
+The experimental setup described in Fig.~\ref{fig:data_flow} allows us to 
 test different data transfer  strategies to the GPU. The standard data transfer 
@@
 PCIe adapter card and the GPU allocate \emph{separate} buffers on the
-system memory for the copy (as shown in Fig.~\ref{fig_standardDT}).
+system memory for the copy (as shown in Fig.~\ref{fig:standardDT}).
 This is inefficient, as the data are copied twice in the 
@@
 Data may also be transferred using Direct Memory Access (DMA, 
-GPUDirect\cite{bib_GPUDirect}) to the CPU memory:
+GPUDirect~\cite{bib_GPUDirect}) to the CPU memory:
 the PCIe card and the GPU share the \textit{same} buffer on the CPU memory; as 
 a result the data are copied only once in the CPU memory 
-(Fig.~\ref{fig_GPUDirectV1}). 
+(Fig.~\ref{fig:GPUDirectV1}). 
 With our experimental setup two additional copy strategies can be tested
@@
 \begin{itemize}
-\item CUDA-Aware MPI, where the copy latency is further reduced
+\item \textsc{cuda}-Aware MPI, where the copy latency is further reduced
 by automatically allocating the buffer on the CPU memory;
@@
 directly to the GPU, without any  
-intermediate copy to the CPU (Fig.~\ref{fig_GPUDirectV2}).
+intermediate copy to the CPU (Fig.~\ref{fig:GPUDirectV2}).
 \end{itemize}
@@
 \subfigure[]
-{\label{fig_standardDT}
+{\label{fig:standardDT}
 \includegraphics[width=0.3\linewidth]{figures/noGPUDirect}}
@@
 \subfigure[]
-{\label{fig_GPUDirectV1}
+{\label{fig:GPUDirectV1}
 \includegraphics[width=0.3\linewidth]{figures/GPUDirect}}
 \subfigure[]
-{\label{fig_GPUDirectV2}
+{\label{fig:GPUDirectV2}
 \includegraphics[width=0.3\linewidth]{figures/GPUDirect-APE}}
@@
   \centering
-  \includegraphics[width=0.9\linewidth]{figures/datatransfer.pdf}
-  \caption{Total latency (data transfer, copy to/from the GPU and data processing on the GPU) as a function of data buffer sizes, for three different levels of optimization of GPUDirect: v1.0, CUDA-aware MPI and peer-to-peer.
-The smallest transferred data packed is 600 kB.}
+  \includegraphics[width=0.9\linewidth]{figures/datatransfer}
+  \caption{Total latency (data transfer, copy to/from the GPU and data
+    processing on the GPU) as a function of data buffer sizes, for
+    three different levels of optimization of GPUDirect: v1.0,
+    \textsc{cuda}-aware MPI and peer-to-peer. The smallest transferred
+    data packed is 600 kB. \fixme{Below it says 600 kb is the place where
+    P2P gets worse?}}
   \label{fig:xferlatency}
@@
 
-In Fig.~\ref{fig:xferlatency} we show the total latency (data transfer, copy to/from the GPU and data processing on the GPU) as a function of data packet size when data are transferred using GPUDirect v1.0,
-CUDA-aware MPI and P2P. For the packet sizes over 600 kB CUDA-aware MPI gives the best performance.
+In Fig.~\ref{fig:xferlatency} we show the total latency (data
+transfer, copy to/from the GPU and data processing on the GPU) as a
+function of data packet size when data are transferred using GPUDirect
+v1.0, \textsc{cuda}-aware MPI and P2P. For the packet sizes over 600
+kB \textsc{cuda}-aware MPI gives the best performance. For smaller
+packet sizes, P2P gives better performance \fixme{do we have this plot?}.
 
@@
   \centering
-  \includegraphics[width=0.9\linewidth]{figures/cudaware.pdf}
-  \caption{data transfer latencies}
+  \includegraphics[width=0.9\linewidth]{figures/cudaware}
+  \caption{Time per fit, in msec. The two curves show total timing
+    with and without calculations performed on the GPU, thereby
+    showing the considerable time spent in data transfer. About
+    20-25\% of the time is spent in data transfer.}
   \label{fig:transferOnly}
@@
 
-The data transfer latency accounts for a significant part if the total latency, 
-as can be seen in Fig.~\ref{fig:transferOnly}: about 20-25\% of total latency  is due to moving the data to and from the GPU.
+The data transfer latency accounts for a significant part of the total
+latency, as can be seen in Fig.~\ref{fig:transferOnly}: about 20-25\%
+of total latency is due to moving the data to and from the GPU.
 
 \section{Conclusions}
+We have examined a staged approach to using accelerator cards in a
+hadron collider trigger application. We have demonstrated that in this
+application, significant gains can be achieved with the
+`embarrassingly parallel' approach on an Intel MIC
+architecture, with the smallest amount of required changes to an
+existing serial code base. However, better performance is achieved
+with GPUs and a more complete event-level parallelization using
+\textsc{cuda}.
 
@@
 
-% \bibitem{bib_cuda} \url{http://www.nvidia.com/object/cuda_home_new.html}
-
-% \bibitem{bib_OpenCL} \url{http://www.khronos.org/opencl/}
-
-% %\bibitem{bib_GPU_TIPP2011} Physics Procedia, Volume 37, 2012, 1965-1972.
-% \bibitem{bib_Apenet} R.~Ammendola \textit{et al.}, \emph{APEnet+: a 3D toroidal network enabling Petaflops scale Lattice QCD simulations on commodity clusters}  PoS(Lattice 2010)022.
-
-% \bibitem{bib_GPUDirect} \url{https://developer.nvidia.com/gpudirect}
-
 \end{document}

### 2013-10-20 papers/2013/CHEP/paper.tex
opencl bib

@@
 }
-\address{$^1$ INFN and University of Padova, Italy}
-\address{$^2$ Los Alamos National Laboratory, New Mexico, USA}
-\address{$^3$ Fermi National Accelerator Laboratory, Illinois, USA}
-\address{$^4$ INFN Roma, Italy}
-\address{$^5$ Cornell University, New York, USA}
+\address{$^1$ INFN and University of Padova, Padova, Italy}
+\address{$^2$ Los Alamos National Laboratory, Los Alamos, New Mexico, USA}
+\address{$^3$ Fermi National Accelerator Laboratory, Batavia, Illinois, USA}
+\address{$^4$ INFN Roma, Rome, Italy}
+\address{$^5$ Cornell University, Ithaca, New York, USA}
 
@@
 various parts of the algorithm for the embarrassingly parallel algorithm 
-(on Intel MIC) and the parallel algorithm (on NVidia Titan GPU), as a 
+(on Intel MIC) and the parallel algorithm (on NVIDIA Titan GPU), as a 
 function of the number of fits. On both accelerator cards the fractional times

### 2013-10-21 papers/2013/CHEP/paper.tex
english pass

@@
-\documentclass[letterpaper]{jpconf}
+\documentclass[a4]{jpconf}
 \usepackage{graphicx}
@@
   test-case for low-latency trigger systems using new computing
-  architectures for LHC experiment.
-
+  architectures for LHC experiment and examine the
+  complexity/performance trade-off in porting existing serial
+  algorithms to many-core devices.
 \end{abstract}
@@
 algorithm latency for processing on the GPU in a manner similar to a
-typical HEP trigger application.
+typical HEP trigger application, and to understand the cost/complexity
+ratio of porting legacy serial code to many-core devices..
 
@@
 
-We ported the track fitting part to parallel devices, as it is well suited to 
-parallelization.
+We ported the track fitting as it is well suited to parallelization -
+each track can be handled independently.
 
@@
 \section{Experimental setup and data flow}
-The many-core devices used in this study are listed in Table~\ref{tab_hwspecs}.
+The many-core devices used in this study are listed in
+Table~\ref{tab_hwspecs}. The GPUs include a less expensive gaming
+class GPU (GTX) and ones optimized for scientific computing
+(Tesla). The MIC corresponds to a Xeon Phi introduced in November 2012.
 
@@
   \end{tabular}
-  \caption{Capabilities of the many-core devices used in this study, according to the manufacturer's specifications.}
+  \caption{Capabilities of the many-core devices used in this study,
+    according to the manufacturer's specifications. The first three
+    are GPUs and the final is an Intel Xeon Phi.}
   \label{tab_hwspecs}
@@
 are constant over 500 input events, where computing resources are saturated. 
-Differently from MIC, the fit stage takes most of the time on the GPU: this 
-could be explained with the intense memory access intrinsic in 
+Unlike the MIC, the fit stage takes most of the time on the GPU: this 
+could be caused by the intense memory access frequency intrinsic to
 this part of the algorithm.  
@@
 \caption{Breakdown of computing time for MIC (a) and the GTX Titan GPU
-  (b). \fixme{we need a legend for these plots. Also can we make them
-    the same size and aspect ratio.}}
+  (b). Red corresponds to unpacking, green corresponds to generating
+  hit combinations, blue is the linearized track fit, and magenta
+  corresponds to offloading (MIC only)\fixme{we need a legend for
+    these plots, and it doesn't work in B\&W. Also can we make them
+    the same size and aspect ratio.} For MIC, combinations and fitting
+  take the same amount of time for large number of events. For GPU,
+  fitting dominates. }
 \label{fig:breakdown}

### 2013-10-21 src_opencl/svt_gpu_opencl.cpp
Now option -s cpu works again

@@ -10,6 +10,7 @@
 #include <string>
 #include <fstream>
 #include <iterator>
+#include <stdio.h>
 
 typedef unsigned int	   DATAWORD; //e.g. UINT32 or float //NOTE: this must absolutely match the typedef in kernel!!!!
 
@@ -165,20 +166,25 @@ int main(int argc, char* argv[]) {
   struct fit_arrays *fit_dev = new fit_arrays;
   struct fout_arrays *fout_dev = new fout_arrays;
 
+
   svtsim_fconread(tf, edata_dev);
 
-  free(tf);
+  //free(tf);
 
   bool is_null=false;
+
   if(evt==NULL) is_null=true;
 
-  printf("Is evt_arrays ptr null? bool=%d",is_null);
+  printf("Is evt_arrays ptr null? bool=%d \n",is_null);
 
   //gf_init_evt(&evt);
 
   gettimeofday(&tBegin, NULL);
 
   if ( strcmp(where,"cpu") == 0 ) { // CPU
+    memcpy(tf->wedge, edata_dev->wedge, sizeof(edata_dev->wedge));
+    memcpy(tf->whichFit, edata_dev->whichFit, sizeof(edata_dev->whichFit));
+    memcpy(tf->lfitparfcon, edata_dev->lfitparfcon, sizeof(edata_dev->lfitparfcon));
     printf("Start work on CPU..... \n");
 
     gettimeofday(&ptBegin, NULL);
@@ -186,17 +192,17 @@ int main(int argc, char* argv[]) {
     gettimeofday(&ptEnd, NULL);
     printf("Time to CPU unpack: %.3f ms\n",
           ((ptEnd.tv_usec + 1000000 * ptEnd.tv_sec) - (ptBegin.tv_usec + 1000000 * ptBegin.tv_sec))/1000.0);
-
+/*
     for(int ie=0; ie<NEVTS; ie++){
       printf("\nEvent %d, nroads = %d",ie,tf->evt_nroads[ie]);
     }
-
+*/
     gettimeofday(&ptBegin, NULL);
     gf_fep_comb(tf);
     gettimeofday(&ptEnd, NULL);
     printf("Time to CPU comb: %.3f ms\n",
           ((ptEnd.tv_usec + 1000000 * ptEnd.tv_sec) - (ptBegin.tv_usec + 1000000 * ptBegin.tv_sec))/1000.0);
-
+/*
[…]

### 2013-10-21 src_opencl/svtsim_functions.c

@@ -476,6 +476,26 @@ svtsim_cable_t * svtsim_cable_new(void) {
  }
 
 
+  int
+ svtsim_whichFit_tf(tf_arrays_t tf, int zin, int layerMask, int lcMask)
+ {
+#ifdef DEBUG_SVT 
+   printf("in svtsim_whichFit: zin = %d, layerMask = %x, lcMask = %x\n", zin, layerMask, lcMask);
+#endif
+
+   int which0 = 0, which = 0;
+   if (zin<0 || zin>=SVTSIM_NBAR) zin = 0;
+   which0 = svtsim_whichFit_full(zin, layerMask, lcMask);
+   which = tf->whichFit[zin][which0];
+
+#ifdef DEBUG_SVT 
+   printf("in svtsim_whichFit: which0 = %d, which = %x\n", which0, which);
+#endif
+
+   return which;
+ }
+
+
 
   int
  svtsim_whichFit(struct extra_data* edata, int zin, int layerMask, int lcMask)
@@ -1855,7 +1875,7 @@ int gf_fep_unpack_evt(evt_arrays_t tf, int n_words_in, void* data_in) {
  
      iz = i&7, lcl = i>>3 & 0x1f, hit = i>>8 & 0x3f;
 
-     which = svtsim_whichFit(tf, iz, hit, lcl);
+     which = svtsim_whichFit_tf(tf, iz, hit, lcl);
      coeff = iz + which*6;  /* poor choice for illegal iz=6,7, but compatible */
      intcp = which;
 

### 2013-10-21 src_opencl/svt_gpu_opencl.cpp
updates with some timing calculation things added, probably breaking CPU mode in src_opencl, but thats ok since we can just use version in src directory

@@ -2,6 +2,7 @@
 #include <sys/time.h>
 #include "semaphore.c"
 #include "svt_gpu_opencl.h"
+#include <math.h>
 
 #define __NO_STD_VECTOR // Use cl::vector instead of STL version
 #include <CL/cl.hpp>
@@ -11,6 +12,34 @@
 #include <fstream>
 #include <iterator>
 #include <stdio.h>
+
+const int gf_maskdata[] = {
+  0x00000000,
+  0x00000001, 0x00000003, 0x00000007, 0x0000000f,
+  0x0000001f, 0x0000003f, 0x0000007f, 0x000000ff,
+  0x000001ff, 0x000003ff, 0x000007ff, 0x00000fff,
+  0x00001fff, 0x00003fff, 0x00007fff, 0x0000ffff,
+  0x0001ffff, 0x0003ffff, 0x0007ffff, 0x000fffff,
+  0x001fffff, 0x003fffff, 0x007fffff, 0x00ffffff,
+  0x01ffffff, 0x03ffffff, 0x07ffffff, 0x0fffffff,
+  0x1fffffff, 0x3fffffff, 0x7fffffff, 0xffffffff
+};
+
+const unsigned long gf_maskdata3[] = {
+  0x000000000000UL,
+  0x000000000001UL, 0x000000000003UL, 0x000000000007UL, 0x00000000000fUL,
+  0x00000000001fUL, 0x00000000003fUL, 0x00000000007fUL, 0x0000000000ffUL,
+  0x0000000001ffUL, 0x0000000003ffUL, 0x0000000007ffUL, 0x000000000fffUL,
+  0x000000001fffUL, 0x000000003fffUL, 0x000000007fffUL, 0x00000000ffffUL,
+  0x00000001ffffUL, 0x00000003ffffUL, 0x00000007ffffUL, 0x0000000fffffUL,
+  0x0000001fffffUL, 0x0000003fffffUL, 0x0000007fffffUL, 0x000000ffffffUL,
+  0x000001ffffffUL, 0x000003ffffffUL, 0x000007ffffffUL, 0x00000fffffffUL,
+  0x00001fffffffUL, 0x00003fffffffUL, 0x00007fffffffUL, 0x0000ffffffffUL,
+  0x0001ffffffffUL, 0x0003ffffffffUL, 0x0007ffffffffUL, 0x000fffffffffUL,
[…]

### 2013-10-21 papers/2013/CHEP/paper.tex
Modified abstract and intro.

@@
 \begin{abstract}
-  In this paper we describe updates to a previous study of latencies
-  in GPU for potential trigger applications, where we measured the
-  latency to transfer data to/from the GPU, exploring the timing of
-  different I/O technologies on different GPU models.  Those studies,
-  where a simplified track fitting algorithm was parallelized and run
-  on a GPU, show that latencies of few tens of microseconds can be
-  achieved to transfer and process packets of 4 kB of data, combining
-  the modern Infiniband data transfer technology with direct access to
-  GPU memory allowed by NVIDIA GPUDirect utilities.  We now have
-  expanded our latency studies to include other multi-core systems
-  (Intel Xeon Phi/MIC in addition to NVIDIA GPUs). We also discuss the
-  implementation of a scaled-up version of the algorithm used at CDF
-  for online track reconstruction - the SVT algorithm - as a realistic
-  test-case for low-latency trigger systems using new computing
-  architectures for LHC experiment and examine the
+  Interest in parallel architectures applied to real time selections is 
+  growing in High Energy Physics (HEP) experiments. In this paper we describe
+  performance measurements of Graphic Processing Units (GPUs) and Intel Many
+  Integrated Core architecture (MIC) when applied to a typical HEP online task: 
+  the selection of events based on the trajectories of charged particles.
+  We use as benchmark a scaled-up version of the algorithm used at CDF 
+  experiment at Tevatron for online track reconstruction - the SVT algorithm - 
+as a realistic  test-case for low-latency trigger systems using new computing
+  architectures for LHC experiment. We examine the
   complexity/performance trade-off in porting existing serial
-  algorithms to many-core devices.
+  algorithms to many-core devices. Measurements of both data processing 
+  and data transfer latency are shown, considering different I/O strategies 
+  to/from the parallel devices. 
 \end{abstract}
@@
 consumer market, are less expensive than dedicated hardware, and are
-easier to support.  Among the commercial devices, Graphic Processing
-Units (GPUs) are of particular interest for online selections given
-their impressive computing power (the latest NVIDIA~\cite{bib_nvidia}
-GPU architecture, Kepler, exceeds Teraflop computing power); moreover,
+easier to support.  Among the commercial devices, many-core architectures 
+such as Graphic Processing Units (GPUs)~\cite{bib_gpu} and Intel Many
+  Integrated Core (MIC)~\cite{bib_intelMIC} are of particular interest for 
+online 
+selections given their great computing power (the latest 
+\textsc{nvidia}~\cite{bib_nvidia} GPU architecture, Kepler, exceeds Teraflop 
+computing power); moreover,
 high-level programming architectures based on C/C++ such as
-\textsc{cuda}~\cite{bib_cuda} and \textsc{opencl}~\cite{bib_opencl}
-make these devices more accessible to the general user.  The ultimate
+\textsc{cuda}~\cite{bib_cuda}  and 
+\textsc{opencl}~\cite{bib_opencl} make these devices more accessible to 
+the general user.  The ultimate
 goal of this study is to investigate the strengths and weaknesses of
-GPUs when applied in a low latency environment, with particular
-emphasis on the data transfer latency to/from the GPU and the
-algorithm latency for processing on the GPU in a manner similar to a
+many-core devices when applied in a low latency environment, with particular
+emphasis on the data transfer latency to/from the device and the
+algorithm latency for processing on the device in a manner similar to a
 typical HEP trigger application, and to understand the cost/complexity
-ratio of porting legacy serial code to many-core devices..
+ratio of porting legacy serial code to many-core devices.
 
 We showed initial studies on GPU performance in low-latency
-environments ($\sim 100~\mu$s) in a previous
-paper~\cite{TIPP2011,NSS2012}.  In this paper we extend those studies
-measuring the timing performance of different GPU cards in different
-data I/O environments. The algorithm run on the GPU is a complete
+environments ($\sim 100~\mu$s) in  previous
+papers~\cite{TIPP2011,NSS2012}.  In this paper we extend those studies
+to include other many-core systems (Intel MIC in addition to GPUs).
+The algorithm run on the parallel architecture is a complete
 version of the fast track fitting algorithm of the Silicon Vertex
@@
 %card. 
-One of the goals of the studies is to explore using a concrete
-algorithm what gains one can achieve with varying degrees of changes
-to the existing algorithm. Starting with a serial algorithm
+Starting with a serial algorithm
 implemented on a CPU, we test an \textit{embarrassingly parallel}
@@
 first approach is rather modest, and the second approach shows larger
-performance gains.
+performance gains. For GPUs, we also test different strategies to transfer
+ data to/from the device.
 
@@
     according to the manufacturer's specifications. The first three
-    are GPUs and the final is an Intel Xeon Phi.}
+    are \textsc{nvidia} GPUs and the final is an Intel Xeon Phi.}
   \label{tab_hwspecs}
@@
    \caption{Speed-up with respect to a standard CPU (Intel Xeon
-     E5630). The speed-ups plateau after about one million fits for
-     the GPUs (in red) while the Intel MIC still shows a small slope
-     beyond that point. \fixme{is this true}}
+     E5630). The speed-ups plateau after about two million fits.
+     %for
+     %the GPUs (in red) while the Intel MIC still shows a small slope
+     %beyond that point. 
+     \fixme{Silvia: removed the comment about the different GPU/MI beahviour. These are the speedup values for the MIC: 0.07, 0.49, 1.79, 2.25, 2.32, 3.35, 4.38,4.74, 5.36, 5.16, 5.34 }}
    \label{fig:algo_only_speedup}
@@
   hit combinations, blue is the linearized track fit, and magenta
-  corresponds to offloading (MIC only)\fixme{we need a legend for
-    these plots, and it doesn't work in B\&W. Also can we make them
-    the same size and aspect ratio.} For MIC, combinations and fitting
+  corresponds to offloading (MIC only). For MIC, combinations and fitting
   take the same amount of time for large number of events. For GPU,
@@
 v1.0, \textsc{cuda}-aware MPI and P2P. For the packet sizes over 600
-kB \textsc{cuda}-aware MPI gives the best performance. For smaller
-packet sizes, P2P gives better performance \fixme{do we have this plot?}.
+kB \textsc{cuda}-aware MPI gives the best performance (see also~\cite{NSS2012} and~\cite{bib_mvapich})  \fixme{do we have this plot?
+Silvia: no, and will not have in time for the proceedings...
+but we can put a reference, see slides after page 47.}.
 

### 2013-10-22 papers/2013/CHEP/paper.tex
better conclusions, bib cleanups

@@
       low number of fits, the CPU performs better, due to start-up
-      costs associated with data transfers.}
+      costs associated with data transfers to the accelerator card.}
     \label{fig:algo_only_timing_zoom}
@@
    \caption{Speed-up with respect to a standard CPU (Intel Xeon
-     E5630). The speed-ups plateau after about two million fits.
-     %for
-     %the GPUs (in red) while the Intel MIC still shows a small slope
-     %beyond that point. 
-     \fixme{Silvia: removed the comment about the different GPU/MI beahviour. These are the speedup values for the MIC: 0.07, 0.49, 1.79, 2.25, 2.32, 3.35, 4.38,4.74, 5.36, 5.16, 5.34 }}
+     E5630). The speed-ups plateau after about two million fits.}
    \label{fig:algo_only_speedup}
@@
 \caption{Breakdown of computing time for MIC (a) and the GTX Titan GPU
-  (b). Red corresponds to unpacking, green corresponds to generating
-  hit combinations, blue is the linearized track fit, and magenta
-  corresponds to offloading (MIC only). For MIC, combinations and fitting
-  take the same amount of time for large number of events. For GPU,
-  fitting dominates. }
+  (b). White corresponds to unpacking, green hash corresponds to
+  generating hit combinations, solid blue is the linearized track fit,
+  and magenta cross-hatch corresponds to offloading (MIC only). For
+  MIC, combinations and fitting take the same amount of time for large
+  number of events. For GPU, fitting dominates. }
 \label{fig:breakdown}
@@
     three different levels of optimization of GPUDirect: v1.0,
-    \textsc{cuda}-aware MPI and peer-to-peer. The smallest transferred
-    data packed is 600 kB. \fixme{Below it says 600 kb is the place where
-    P2P gets worse?}}
+    \textsc{cuda}-aware MPI and P2P. The smallest transferred data
+    packed is 600 kB. \fixme{Below it says 600 kb is the place where
+      P2P gets worse?} \textsc{cuda}-aware MPI shows the best
+    performance for larger packet size.}
   \label{fig:xferlatency}
@@
 v1.0, \textsc{cuda}-aware MPI and P2P. For the packet sizes over 600
-kB \textsc{cuda}-aware MPI gives the best performance (see also~\cite{NSS2012} and~\cite{bib_mvapich})  \fixme{do we have this plot?
-Silvia: no, and will not have in time for the proceedings...
-but we can put a reference, see slides after page 47.}.
+kB \textsc{cuda}-aware MPI gives the best performance (see
+also~\cite{NSS2012} and~\cite{bib_mvapich})  \fixme{do we have this
+  plot? Silvia: no, and will not have in time for the
+  proceedings... but we can put a reference, see slides after page
+  47. Peter: what page 47? I'm confused.}.
 
@@
 \section{Conclusions}
-We have examined a staged approach to using accelerator cards in a
-hadron collider trigger application. We have demonstrated that in this
-application, significant gains can be achieved with the
-`embarrassingly parallel' approach on an Intel MIC
+We have implemented a full version of the CDF SVT tracking algorithm
+on GPUs and Intel MIC. We examined a staged approach to using
+accelerator cards in a hadron collider trigger application. We have
+demonstrated that in this application, significant gains can be
+achieved with the `embarrassingly parallel' approach on an Intel MIC
 architecture, with the smallest amount of required changes to an
@@
 with GPUs and a more complete event-level parallelization using
-\textsc{cuda}.
+\textsc{cuda}.  We have updated latency studies and shown that for
+larger packet sizes (greater than 600~kB), \textsc{cuda}-aware MPI
+outperforms P2P, contrary to naive expectation.  Even at large packet
+size, the data transfer takes an appreciable fraction of the total
+algorithm time (about 20-25\%).
 

### 2013-10-23 src_opencl/clHelperFuncs.cpp
Fix the problem with output. Now I have a strange problem: CL_OUT_OF_RESOURCES.

@@ -21,6 +21,7 @@ void CL_HELPERFUNCS::getPlatformsAndDevices(cl::vector< cl::Platform > *platform
 	cout << __FANCY__ << "Num of Platforms: " << (*platformList).size() << endl;
 
 	string platformVendor;
+	string platformVersion;
 	cl_device_type deviceType;
 	GPU_device_index = -1; GPU_platform_index = -1;
 
@@ -30,6 +31,8 @@ void CL_HELPERFUNCS::getPlatformsAndDevices(cl::vector< cl::Platform > *platform
 		(*platformList)[i].getInfo((cl_platform_info)CL_PLATFORM_VENDOR, &platformVendor);
 		cout << __FANCY__ << "Platform " << i << " is by: " << platformVendor << "\n";
 
+		(*platformList)[i].getInfo((cl_platform_info)CL_PLATFORM_VERSION, &platformVersion);
+		cout << __FANCY__ << "Platform " << i << " version is: " << platformVersion << "\n";
 
 		(*deviceList).push_back(new cl::vector< cl::Device >());
 

### 2013-10-23 src_opencl/svt_gpu_opencl.cpp

@@ -511,7 +511,7 @@ int main(int argc, char* argv[]) {
 				   0,
 				   sizeof(extra_data),
 				   edata_dev);
-    CL_HELPERFUNCS::checkErr(err, "ComamndQueue::enqueueWriteBuffer()");
+    CL_HELPERFUNCS::checkErr(err, "ComamndQueue::enqueueWriteBuffer(edata)");
 
     event.wait();
 
@@ -859,7 +859,7 @@ int main(int argc, char* argv[]) {
 				   0,
 				   sizeof(evt_arrays),
 				   evt);
-    CL_HELPERFUNCS::checkErr(err, "CommandQueue::enqueueWriteBuffer()");
+    CL_HELPERFUNCS::checkErr(err, "CommandQueue::enqueueWriteBuffer(evt)");
    
     err = queue.finish();
 
@@ -968,7 +968,7 @@ int main(int argc, char* argv[]) {
     err = queue.enqueueNDRangeKernel(
 				     kernel_compute_eeword,
 				     cl::NullRange,
-				     cl::NDRange(NEVTS+156),
+				     cl::NDRange(256*((NEVTS%256)+1)),
 				     cl::NDRange(256),
 				     NULL,
 				     &event);
@@ -1012,7 +1012,7 @@ int main(int argc, char* argv[]) {
 				  0,
 				  sizeof(fout_arrays),
 				  fout_dev);
-    CL_HELPERFUNCS::checkErr(err, "CommandQueue::enqueueReadBuffer()");
+    CL_HELPERFUNCS::checkErr(err, "CommandQueue::enqueueReadBuffer(fout)");
 
     err = queue.finish();
     gettimeofday(&ptEnd, NULL);

### 2013-10-23 src_opencl/svt_gpu_opencl.cpp
Fix a bug in range kernel call

@@ -211,7 +211,7 @@ int main(int argc, char* argv[]) {
   gettimeofday(&tBegin, NULL);
 
   const bool PRINT_TIME=false;
-  const int N_LOOPS=1000;
+  const int N_LOOPS=1;
   const int N_CHECKS=5;
   float times[N_CHECKS][N_LOOPS];
   int n_iters=0;
@@ -968,8 +968,8 @@ int main(int argc, char* argv[]) {
     err = queue.enqueueNDRangeKernel(
 				     kernel_compute_eeword,
 				     cl::NullRange,
-				     cl::NDRange(256*((NEVTS%256)+1)),
-				     cl::NDRange(256),
+				     cl::NDRange(NEVTS),
+				     cl::NDRange(1),
 				     NULL,
 				     &event);
     CL_HELPERFUNCS::checkErr(err, "ComamndQueue::enqueueNDRangeKernel(compute_eeword)");

### 2013-10-24 papers/2013/CHEP/paper.tex
fix mvapich fix
editing pass

@@
 \section{Introduction}
-%%% Verbatim from NSS paper
 Real-time event reconstruction plays a fundamental role in High Energy
@@
 level based on custom hardware, followed by one or two levels usually
-based on farms of general purpose processors.  The possibility of also
-using commercial devices at a low trigger level is very appealing:
-they are subject to continuous performance improvements driven by the
-consumer market, are less expensive than dedicated hardware, and are
-easier to support.  Among the commercial devices, many-core architectures 
-such as Graphic Processing Units (GPUs)~\cite{bib_gpu} and Intel Many
-  Integrated Core (MIC)~\cite{bib_intelMIC} are of particular interest for 
-online 
-selections given their great computing power (the latest 
-\textsc{nvidia}~\cite{bib_nvidia} GPU architecture, Kepler, exceeds Teraflop 
-computing power); moreover,
-high-level programming architectures based on C/C++ such as
-\textsc{cuda}~\cite{bib_cuda}  and 
-\textsc{opencl}~\cite{bib_opencl} make these devices more accessible to 
-the general user.  The ultimate
-goal of this study is to investigate the strengths and weaknesses of
-many-core devices when applied in a low latency environment, with particular
-emphasis on the data transfer latency to/from the device and the
-algorithm latency for processing on the device in a manner similar to a
-typical HEP trigger application, and to understand the cost/complexity
-ratio of porting legacy serial code to many-core devices.
+based on farms of general purpose processors.  At all levels, latency
+is a concern: for a fixed processing time, the faster a decision is
+rendered about accepting or rejecting an event improves the purity of
+the collected data sample. The possibility of using commercial devices
+at a low trigger level is very appealing: they are subject to
+continuous performance improvements driven by the consumer market, are
+less expensive than dedicated hardware, and are easier to support.
+Among the commercial devices, many-core architectures such as Graphic
+Processing Units (GPUs)~\cite{bib_gpu} and Intel Many Integrated Core
+(MIC)~\cite{bib_intelMIC} are of particular interest for online
+selections given their great computing power: the latest
+\textsc{nvidia}~\cite{bib_nvidia} GPU architecture, Kepler, exceeds
+Teraflop computing power. Moreover, high-level programming
+architectures based on \textsc{c/c++} such as
+\textsc{cuda}~\cite{bib_cuda} and \textsc{opencl}~\cite{bib_opencl}
+make programming these devices more accessible to the general
+physicist user.  The goal of this study is to investigate the
+strengths and weaknesses of many-core devices when applied in a low
+latency environment, with particular emphasis on the data transfer
+latency to/from the device and the algorithm latency for processing on
+the device in a manner similar to a typical HEP trigger application,
+and to understand the cost/complexity ratio of porting legacy serial
+code to many-core devices.
 
 We showed initial studies on GPU performance in low-latency
-environments ($\sim 100~\mu$s) in  previous
+environments ($\approx 100~\mu$s) in previous
 papers~\cite{TIPP2011,NSS2012}.  In this paper we extend those studies
-to include other many-core systems (Intel MIC in addition to GPUs).
-The algorithm run on the parallel architecture is a complete
-version of the fast track fitting algorithm of the Silicon Vertex
+to include other many-core architectures (Intel MIC in addition to
+GPUs).  The algorithm run on the parallel architecture is a complete
+version of the fast track-fitting algorithm of the Silicon Vertex
 Tracker (SVT) system at CDF~\cite{SVT1}.
@@
 performance gains. For GPUs, we also test different strategies to transfer
- data to/from the device.
+data to and from the device.
 
 \section{SVT track fitting algorithm}
-The Silicon Vertex Trigger (SVT)~\cite{SVT1,SVT2} is a 
-track reconstruction processor used at 
-trigger level in the CDF experiment at Tevatron accelerator. It reconstructs 
-tracks in about 20 $\mu s$ in time in two steps: for each event 
-first low resolution tracks (\textit{roads}) are found among the hits left 
-in the tracking detector by charged particles; then track fitting is 
-performed on all possible combinations of hits inside a road. 
-This algorithm uses a linearized approximation to track-fitting as 
-implemented in hardware (described in greater detail in~\cite{SVT3}). 
-With the SVT approach, the determination of the track parameters 
-($p_i$) is reduced to a simple scalar product:
+The Silicon Vertex Trigger (SVT)~\cite{SVT1,SVT2} is a track
+reconstruction processor used in the CDF experiment at Tevatron
+accelerator. It reconstructs tracks in about 20 $\mu s$ in two steps:
+first, low resolution tracks (\textit{roads}) are found in each event
+among the energy deposits left in the tracking detector by charged
+particles; second, track fitting is performed on all possible
+combinations of hits inside a road.  This algorithm uses a linearized
+approximation to track-fitting as implemented in hardware (described
+in greater detail in~\cite{SVT3}).  With the linearized track fit of
+the SVT approach, the determination of the track parameters ($p_i$) is
+reduced to a simple scalar product:
 \[
@@
 simulation code, written in the \textsc{c} language. SVT track fitting
-is divided into three main functions: the unpacking of input data and
-filling of all the necessary arrays, followed by the computation of
-all possible combinations of hits in each road and finally the fit of
-each combination. Three main loops are present - on events, roads and
-hit combinations.  To be run on GPUs, the code has been ported to
-\textsc{cuda}: each step - unpack, combine and track fit - is
-performed by a specific kernel; the three nested loops are unrolled so
-that each GPU thread processes a single combination of hits.  To run
-on MIC, we adopted the so-called \textit{embarrassingly parallel}
-approach and used \textsc{pragma OpenMP} for statements to unroll the external
-loop on the events, so that each core processes a single event: the
-porting requires much less effort compared to \textsc{cuda}, but the level of
-parallelism is limited.
+is divided into three main functions: first, the unpacking of input
+data and filling of all the necessary data structures; second, the
+computation of all possible combinations of hits in each road and
+third, the linearized track fit of each combination of hits. Three
+main loops are present - on events, roads and hit combinations.  To be
+run on GPUs, the code has been ported to \textsc{cuda}: each step -
+unpack, combine and track fit - is performed by a specific kernel; the
+three nested loops are unrolled so that each GPU thread processes a
+single combination of hits.  To run on MIC, we adopted the so-called
+\textit{embarrassingly parallel} approach and used \textsc{pragma
+  OpenMP} for statements to unroll the external loop on the events, so
+that each core processes a single event: the porting requires much
+less effort compared to \textsc{cuda}, but the level of parallelism is
+limited.
 
@@
 The latency for a complete loop is measured on the transmitter using 
-standard C libraries.
-\begin{figure}[tbph]
+standard \textsc{c} libraries.
+\begin{figure}[btp]
 \centering
@@
 \section{Results}
-We use as input data events with a fixed number of roads and combinations: 
-each event has 2048 combinations to be fitted. To explore different data 
-taking conditions the number of events ranges from one to 3000, i.e. from 2048 
-to about six millions of combinations to fit.
+The input data consists of events with a fixed number of roads and
+combinations: each event has 2048 combinations to be fitted. To
+explore different data-taking conditions, the number of events ranges
+from one to 3000, \emph{i.e.,} from 2048 to about six millions of combinations
+to fit.
 
@@
 affords a much more significant speed improvement. 
+\begin{figure}[tbp]
+  \centering
+  \includegraphics[width=0.9\linewidth]{figures/TimeComp_MIC}
+  \caption{Algorithm-only comparison for timing as a function of the
+    number of track fits. We compare timing on CPUs (serial), Intel
+    MIC (embarrassingly parallel), and GPUs (fully parallel), in
+    blue, greed, and red, respectively. The GPUs exhibit the best
+    performance due to the full parallelization. }
+  \label{fig:algo_only_timing}
+\end{figure}
+\begin{figure}[tbp]
+  \centering
+  \includegraphics[width=0.9\linewidth]{figures/TimeCompZoom_MIC.png} 
+  \caption{Algorithm-only comparison for timing as a function of the
+    number of track fits: zoom in the low number of fits region. At
+    low number of fits, the CPU performs better, due to start-up
+    costs associated with data transfers to the accelerator card.}
+  \label{fig:algo_only_timing_zoom}
+\end{figure}
+\begin{figure}[tbp]
+  \centering
+  \includegraphics[width=0.9\linewidth]{figures/Speedup_MIC}
+  \caption{Speed-up with respect to a standard CPU (Intel Xeon
+    E5630). The speed-ups plateau after about two million fits.}
+  \label{fig:algo_only_speedup}
+\end{figure}
 Accelerators performance drop with decreasing number of fits, as can be seen 
 in Fig.~\ref{fig:algo_only_timing_zoom}.
-Fig.~\ref{fig:algo_only_speedup} shows the speed-up with respect to the serial 
+Figure~\ref{fig:algo_only_speedup} shows the speed-up with respect to the serial 
 algorithm run on a standard CPU (Intel Xeon E5630): the maximum gain is 
@@
 
- \begin{figure}[!tbp]
-    \centering
-    \includegraphics[width=0.9\linewidth]{figures/TimeComp_MIC}
-    \caption{Algorithm-only comparison for timing as a function of the
-      number of track fits. We compare timing on CPUs (serial), Intel
-      MIC (embarrassingly parallel), and GPUs (fully parallel), in
-      blue, greed, and red, respectively. The GPUs exhibit the best
-      performance due to the full parallelization. }
-    \label{fig:algo_only_timing}
-  \end{figure}
-
-  \begin{figure}[!tbp]
-    \centering
-    \includegraphics[width=0.9\linewidth]{figures/TimeCompZoom_MIC.png} 
-    \caption{Algorithm-only comparison for timing as a function of the
-      number of track fits: zoom in the low number of fits region. At
-      low number of fits, the CPU performs better, due to start-up
-      costs associated with data transfers to the accelerator card.}
-    \label{fig:algo_only_timing_zoom}
-  \end{figure}
-
- \begin{figure}[!tbp]
-   \centering
-   \includegraphics[width=0.9\linewidth]{figures/Speedup_MIC}
-   \caption{Speed-up with respect to a standard CPU (Intel Xeon
-     E5630). The speed-ups plateau after about two million fits.}
-   \label{fig:algo_only_speedup}
- \end{figure}
-
 \subsubsection{Breakdown of computing time}
@@
 various parts of the algorithm for the embarrassingly parallel algorithm 
-(on Intel MIC) and the parallel algorithm (on NVIDIA Titan GPU), as a 
+(on Intel MIC) and the parallel algorithm (on \textsc{nvidia} Titan GPU), as a 
 function of the number of fits. On both accelerator cards the fractional times
-are constant over 500 input events, where computing resources are saturated. 
+are constant for more than 500 input events, where computing resources are saturated. 
 Unlike the MIC, the fit stage takes most of the time on the GPU: this 
@@
     \textsc{cuda}-aware MPI and P2P. The smallest transferred data
-    packed is 600 kB. \fixme{Below it says 600 kb is the place where
-      P2P gets worse?} \textsc{cuda}-aware MPI shows the best
+    packed is 600 kB.  \textsc{cuda}-aware MPI shows the best
     performance for larger packet size.}
@@
 kB \textsc{cuda}-aware MPI gives the best performance (see
-also~\cite{NSS2012} and~\cite{bib_mvapich})  \fixme{do we have this
-  plot? Silvia: no, and will not have in time for the
-  proceedings... but we can put a reference, see slides after page
-  47. Peter: what page 47? I'm confused.}.
+also~\cite{NSS2012} and~\cite{bib_mvapich})  % \fixme{do we have this
+  % plot? Silvia: no, and will not have in time for the
+  % proceedings... but we can put a reference, see slides after page
+  % 47. Peter: I don't clearly see what that tells me....}.
 

### 2013-10-24 src_opencl/svt_gpu_opencl.cpp
this code works for non-100 events on the AMD card at FNAL!

@@ -71,6 +71,7 @@ void set_outcable_fout(fout_arrays* fout_dev, int totEvts, unsigned int *&data_r
 
   for (int i=0; i < ow ; i++) {
     data_rec[i] = out->data[i];
+    //printf("output word %d is 0x%x\n",i,out->data[i]);
   }
 
 //  svtsim_free(out);
@@ -294,7 +295,7 @@ int main(int argc, char* argv[]) {
     //plat_i = 1;	
     
     //this is for AMD card...
-    dev_i = 0;
+    dev_i = 1;
     plat_i = 0;	
 
 
@@ -319,8 +320,8 @@ int main(int argc, char* argv[]) {
     CL_HELPERFUNCS::checkErr(err, "Context::Context()");
 
     //run specified kernel on device based on plat_i and dev_i
-    //std::cout << __FANCY__ << "selected [Platform, Device] is [" << plat_i << "," << dev_i << "]" << 
-    //  " type=" << (CL_HELPERFUNCS::isDeviceTypeGPU(&deviceList,plat_i,dev_i)?"GPU":"CPU") << std::endl;
+    std::cout << __FANCY__ << "selected [Platform, Device] is [" << plat_i << "," << dev_i << "]" << 
+      " type=" << (CL_HELPERFUNCS::isDeviceTypeGPU(&deviceList,plat_i,dev_i)?"GPU":"CPU") << std::endl;
     
     cl::Device device = (*deviceList[plat_i])[dev_i]; //get current device to simplify calls
     
@@ -353,7 +354,7 @@ int main(int argc, char* argv[]) {
     
     err = program_init.build(*(deviceList[plat_i]),buildOptions_init.c_str());
 	      
-    //CL_HELPERFUNCS::displayBuildLog(program_init, device);
+    CL_HELPERFUNCS::displayBuildLog(program_init, device);
     CL_HELPERFUNCS::checkErr(err, "Build::Build()");
 
     //====================================
@@ -381,7 +382,7 @@ int main(int argc, char* argv[]) {
     
     err = program.build(*(deviceList[plat_i]),buildOptions.c_str());
 	      
[…]

### 2013-10-24 src_opencl/svtsim_defines.h

@@ -23,7 +23,7 @@
 
 
 //GF
-#define NEVTS 10
+#define NEVTS 200
 #define NROADS 64
 #define EVT 0 //all words from the input file are considered belonging to the same event
 

### 2013-10-24 src_opencl/svtsim_functions.c

@@ -563,6 +563,7 @@ int svtsim_fconread(tf_arrays_t tf, struct extra_data *edata_dev)
    
    for (i = 0; i<NEVTS; i++) {
      edata_dev->wedge[i]=WEDGE;
+     tf->wedge[i]=WEDGE;
    }
 
    for (i = 0; i<SVTSIM_NBAR; i++) {
@@ -600,6 +601,7 @@ int svtsim_fconread(tf_arrays_t tf, struct extra_data *edata_dev)
        * Initialize all whichFit entries to bogus values
        */
       for (i = 0; i<SVTSIM_NEL(edata_dev->whichFit[k]); i++) edata_dev->whichFit[k][i] = -1;
+      for (i = 0; i<SVTSIM_NEL(tf->whichFit[k]); i++) tf->whichFit[k][i] = -1;
  
       /*
        * Open .fcon
@@ -663,6 +665,7 @@ int svtsim_fconread(tf_arrays_t tf, struct extra_data *edata_dev)
         if (bogusLayers) continue;
         which = svtsim_whichFit_full(k, layerMask, 0);
         edata_dev->whichFit[k][which] = which;
+        tf->whichFit[k][which] = which;
         if(0) printf("w=%d z=%d: found detLayers=%s => "
  		    "amLayers=%s mask=%x which=%d\n", 
  		    WEDGE, k, str, layers, layerMask, which);
@@ -713,6 +716,7 @@ int svtsim_fconread(tf_arrays_t tf, struct extra_data *edata_dev)
   	      (tf->lfitpar[l][j][k][which] >> h[j]);
   	    /* fill lfitparfcon array with constants as read from fcon file SA*/
   	    edata_dev->lfitparfcon[l][j][k][which] = fc[j];
+  	    tf->lfitparfcon[l][j][k][which] = fc[j];
   #ifdef DEBUG_READFCON
   	    printf("lfitparfcon[%d][%d][%d][%d] = %.6x\n", l, j, k, which, edata_dev->lfitparfcon[l][j][k][which]);
   #endif
@@ -721,6 +725,7 @@ int svtsim_fconread(tf_arrays_t tf, struct extra_data *edata_dev)
   	tf->lfitpar[l][DIMSPA][k][which] = fc[DIMSPA];
   	/* fill lfitparfcon array with constants as read from fcon file SA*/
[…]

### 2013-10-29 papers/2013/CHEP/paper.tex
Added comments from Ape group

@@
 three nested loops are unrolled so that each GPU thread processes a
-single combination of hits.  To run on MIC, we adopted the so-called
+single combination of hits.  To run on MIC, where cores are more powerful 
+but fewer in number, we adopted the so-called
 \textit{embarrassingly parallel} approach and used \textsc{pragma
-  OpenMP} for statements to unroll the external loop on the events, so
+  OpenMP} for statements to unroll only the external loop on the events, so
 that each core processes a single event: the porting requires much
@@
 CPU and two Tesla M2075 GPU cards. The nodes are connected by InfiniBand 
-communication links using Connect-X2 Mellanox or Apenet+ adapters. 
-Apenet+ is an FPGA-based PCIe board supporting peer-to-peer communication with 
+communication links using Connect-X2 Mellanox or APEnet+ adapters. 
+APEnet+ is an FPGA-based PCIe board supporting peer-to-peer communication with 
 Tesla and Kepler cards~\cite{apenet2010}.
@@
   \centering
-  \includegraphics[width=0.9\linewidth]{figures/datatransfer}
+  \includegraphics[width=0.85\linewidth]{figures/datatransfer}
   \caption{Total latency (data transfer, copy to/from the GPU and data
@@
 function of data packet size when data are transferred using GPUDirect
-v1.0, \textsc{cuda}-aware MPI and P2P. For the packet sizes over 600
-kB \textsc{cuda}-aware MPI gives the best performance (see
-also~\cite{NSS2012} and~\cite{bib_mvapich})  % \fixme{do we have this
+v1.0, \textsc{cuda}-aware MPI and P2P. For the packet sizes considered in this test 
+(over 600 kB) \textsc{cuda}-aware MPI gives the best performance. This is expected as 
+P2P is optimized for small packet sizes
+ (see also~\cite{NSS2012} and~\cite{bib_mvapich}). As a matter of fact, for larger packet size, the channel 
+throughput becomes dominant: the shortest transfer 
+time of CUDA aware-MPI system is easily explained comparing the link bandwidth of Mellanox board (40 Gb/s) 
+with the smaller throughput of a APEnet+ single link (30 Gb/s).  
+The data transfer latency accounts for a significant part of the total
+latency, as can be seen in Fig.~\ref{fig:transferOnly}: about 20-25\%
+of total latency is due to moving the data to and from the GPU.
+
+% \fixme{do we have this
   % plot? Silvia: no, and will not have in time for the
@@
   \centering
-  \includegraphics[width=0.9\linewidth]{figures/cudaware}
+  \includegraphics[width=0.85\linewidth]{figures/cudaware}
   \caption{Time per fit, in msec. The two curves show total timing
@@
 \end{figure}
-
-The data transfer latency accounts for a significant part of the total
-latency, as can be seen in Fig.~\ref{fig:transferOnly}: about 20-25\%
-of total latency is due to moving the data to and from the GPU.
 
@@
 with GPUs and a more complete event-level parallelization using
-\textsc{cuda}.  We have updated latency studies and shown that for
-larger packet sizes (greater than 600~kB), \textsc{cuda}-aware MPI
-outperforms P2P, contrary to naive expectation.  Even at large packet
+\textsc{cuda}.  We have updated latency studies and shown that for larger packet sizes 
+(greater than 600 kB), cuda- aware MPI outperforms P2P. This result fully agrees with expectations 
+since the P2P mechanism is optimized (and effective) to reduce the transfer latency for 
+small size packets (less than 128KB).
+%We have updated latency studies and shown that for
+%larger packet sizes (greater than 600~kB), \textsc{cuda}-aware MPI
+%outperforms P2P, contrary to naive expectation.  
+Even at large packet
 size, the data transfer takes an appreciable fraction of the total
@@
 \ack
-The authors would like to thank the Fermilab staff and the FTK group at the 
-University of Chicago for their support. This work was supported by the
+The authors would like to thank the Fermilab staff,  the FTK group at the 
+University of Chicago and the INFN-APE group in Rome for their support. This work was supported by the
 U.S. Department of Energy, the U.S. National Science Foundation and the Italian
-Istituto Nazionale di Fisica Nucleare.
+Istituto Nazionale di Fisica Nucleare. This work was partially supported by the 
+EU Framework Programme 7 project EURETILE under grant number 247846.
 

### 2013-10-29 papers/2013/CHEP/paper.tex
Added comments from Ape group
Fixed a typo

@@
 v1.0, \textsc{cuda}-aware MPI and P2P. For the packet sizes considered in this test 
-(over 600 kB) \textsc{cuda}-aware MPI gives the best performance. This is expected as 
+ \textsc{cuda}-aware MPI gives the best performance. This is expected as 
 P2P is optimized for small packet sizes
@@
 \textsc{cuda}.  We have updated latency studies and shown that for larger packet sizes 
-(greater than 600 kB), cuda- aware MPI outperforms P2P. This result fully agrees with expectations 
+(greater than 600 kB), \textsc{cuda}-aware MPI outperforms P2P. This result fully agrees with expectations 
 since the P2P mechanism is optimized (and effective) to reduce the transfer latency for 
-small size packets (less than 128KB).
+small size packets.
 %We have updated latency studies and shown that for

### 2013-11-12 papers/2013/NSS/NSS.tex
starting the work on the NSS paper

--- 2013-10-29 papers/2013/CHEP/paper.tex
@@
-\documentclass[a4]{jpconf}
+\documentclass[journal]{IEEEtran}
 \usepackage{graphicx}
 \usepackage{subfigure}
-\usepackage{url}
-\usepackage{xspace}
-\usepackage{iopams}
-\usepackage[pdftex,
-      colorlinks=true,
-      urlcolor=blue,       % \href{...}{...} external (URL)
-      filecolor=green,     % \href{...} local file
-      linkcolor=blue,       % \ref{...} and \pageref{...}
-      pagebackref,
-      pdfpagemode=UseNone,
-      bookmarksopen=true]{hyperref}
 
-\providecommand{\fixme}[1]{\xspace{\sffamily{[\bfseries{}FIXME:} #1]}}
+\usepackage{hyperref}
+\usepackage{multirow}
+\usepackage[all]{hypcap} % fix links to figs/tabs
 
+\renewcommand{\arraystretch}{1.1} % make tables more readable
+
+% disable the running header on all pages following the title page
+\pagestyle{empty}
+%
 \begin{document}
 % paper title
-\title{Many-core applications to online track reconstruction in HEP experiments}
+\title{Title of NSS Talk}
 % Author list
-\author{S.~Amerio$^1$, 
-  D.~Bastieri$^1$, 
-  M.~Corvo$^1$, 
-  A.~Gianelle$^1$, 
-  W.~Ketchum$^2$,
-  T.~Liu$^3$, 
-  A.~Lonardo$^4$, 
-  D.~Lucchesi$^1$,
-  S.~Poprocki$^5$, 
-  R.~Rivera$^3$, 
-  L.~Tosoratto$^4$,
-  P.~Vicini$^4$
+\author{  A.~Gianelle, 
+  S.~Amerio, 
+  D.~Bastieri, 
+  M.~Corvo, 
+  W.~Ketchum,
+  T.~Liu, 
+  A.~Lonardo, 
+  D.~Lucchesi,
+  S.~Poprocki, 
+  R.~Rivera, 
+  L.~Tosoratto,
+  P.~Vicini
   and 
-  P.~Wittich$^5$,
+  P.~Wittich
 }
-\address{$^1$ INFN and University of Padova, Padova, Italy}
-\address{$^2$ Los Alamos National Laboratory, Los Alamos, New Mexico, USA}
-\address{$^3$ Fermi National Accelerator Laboratory, Batavia, Illinois, USA}
-\address{$^4$ INFN Roma, Rome, Italy}
-\address{$^5$ Cornell University, Ithaca, New York, USA}
-
-\ead{silvia.amerio@pd.infn.it}
-
+\thanks{Manuscript received November 15, 2013.
+This work was supported by the National Science Foundation, the
+Department of Energy Office of Science and INFN.}% <-this % stops a space
+\thanks{A.~Gianelle, S.~Amerio, D.~Bastieri, M.~Corvo, and D.~Lucchesi are with INFN Padova.}% 
+\thanks{W.~Ketchum is with Los Alamos National Laboratory.}%
+\thanks{T.~Liu and R.~Rivera are with Fermi National Accelerator Laboratory.}%
+\thanks{A.~Lonardo, L.~Tosoratto and P.~Vicini are with INFN Roma}% 
+\thanks{S.~Poprocki and P.~Wittich are with Cornell University.}%
+% make the title area
+\maketitle
+% after \maketitle, add the following command to remove the header on the title page
+\thispagestyle{empty}
+% abstract usually in passive voice
 \begin{abstract}
@@
   and data transfer latency are shown, considering different I/O strategies 
-  to/from the parallel devices. 
-\end{abstract}
+  to/from the parallel devices. \end{abstract}
 
@@
 Tracker (SVT) system at CDF~\cite{SVT1}.
-%We also compare the performance of GPU
-%cards from different manufacturers (AMD vs NVIDIA) and different
-%computing environments (\textsc{cuda} and \textsc{opencl}) on the same
-%card. 
+We also investigate the performance of the algorithm on a GPU manufactured by AMD using a different
+computing environment: \textsc{opencl}.
 Starting with a serial algorithm
@@
 third, the linearized track fit of each combination of hits. Three
-main loops are present - on events, roads and hit combinations.  To be
-run on GPUs, the code has been ported to \textsc{cuda}: each step -
-unpack, combine and track fit - is performed by a specific kernel; the
+main loops are present - on events, roads and hit combinations.
+
+To be run on NVidia GPUs, the code has been ported to \textsc{cuda}: each step --
+unpack, combine and track fit -- is performed by a specific kernel; the
 three nested loops are unrolled so that each GPU thread processes a
-single combination of hits.  To run on MIC, where cores are more powerful 
+single combination of hits.  The \textsc{cuda} implementation makes use of 
+\textsc{thrust}~\cite{bib_thrust}, a C++ template library for \textsc{cuda}, in the unpacking step.
+
+To implement the algorithm to run on an AMD GPU, we have ported the \textit{combine} and \textit{track fit}
+\textsc{cuda} kernels to \textsc{opencl}, which requires minimal changes. Because the \textsc{thrust} template libraries
+can only be used with \textsc{cuda}, we resort to unpacking serially on the CPU.
+
+To run on MIC, where cores are more powerful 
 but fewer in number, we adopted the so-called
@@
 Table~\ref{tab_hwspecs}. The GPUs include a less expensive gaming
-class GPU (GTX) and ones optimized for scientific computing
+class GPUs (the NVidia GTX and AMD Radeon cards) and ones optimized for scientific computing
 (Tesla). The MIC corresponds to a Xeon Phi introduced in November 2012.
@@
   \centering
-  \begin{tabular}{|l|c|c|c|c|}
+  \begin{tabular}{|l|c|c|c|c|c|}
     \hline
-    Model & Tesla M2050 & Tesla K20m & GeForce GTX  Titan & MIC 5110P \\
+    Model & Tesla M2050 & Tesla K20m & GeForce GTX  Titan & MIC 5110P & Radeon HD 7970 \\
     \hline
     \hline
-    Performance (SP, GFlops) & 1030 & 3520 & 4500 & 2022 \\
-    Memory bandwidth  (GB/s) & 148 & 208 & 288  & 320\\   
-    Memory size (GB) & 3 & 5 & 6 & 8 \\
-    Number of cores & 448 & 2496 & 2688 & 240 \\
-    Clock speed (GHz) & 1.15 & 0.706 & 0.837 & 1.053 \\
+    Performance (SP, GFlops) & 1030 & 3520 & 4500 & 2022 & 3790\\
+    Memory bandwidth  (GB/s) & 148 & 208 & 288  & 320 & 264\\   
+    Memory size (GB) & 3 & 5 & 6 & 8 & 3\\
+    Number of cores & 448 & 2496 & 2688 & 240 & 2048\\
+    Clock speed (GHz) & 1.15 & 0.706 & 0.837 & 1.053 & 1.375\\
     \hline
@@
     according to the manufacturer's specifications. The first three
-    are \textsc{nvidia} GPUs and the final is an Intel Xeon Phi.}
+    are \textsc{nvidia} GPUs, the MIC 5110P is an Intel Xeon Phi, and the final one is an AMD GPU.}
   \label{tab_hwspecs}
@@
 standard \textsc{c} libraries.
-\begin{figure}[btp]
+\begin{figure}[!t]
 \centering
@@
 
+ADDITIONAL PARAGRAPH ON THE FERMILAB SETUP.
+
 \section{Results}
@@
 affords a much more significant speed improvement. 
-\begin{figure}[tbp]
+\begin{figure}[!t]
   \centering
@@
 \end{figure}
-\begin{figure}[tbp]
+\begin{figure}[!t]
   \centering
@@
 this part of the algorithm.  
-\begin{figure*}[tbp]
+\begin{figure*}[!t]
 \centering
@@
 \end{figure*}
+
+\subsection{Data processing in \textsc{opencl}}
+RESULTS FROM THE FERMILAB SETUP.
 
@@
 
-\begin{figure}[tbp]
+\begin{figure}[!t]
   \centering
@@
 
-%% gives unnumbered ack in CHEP style
-\ack
+%%%% IEEE wants placement as !t
+%\begin{figure}[!t]
+%\centering
+%%\includegraphics[width=3.5in]{myFigure}
+%% where an .eps filename suffix will be assumed under latex, 
+%% and a .pdf suffix will be assumed for pdflatex; or what has been declared
+%% via \DeclareGraphicsExtensions.
+%\caption{Daily abstract submission rate of the 2012 NSS-MIC. }
+%\label{fig_sim}
+%\end{figure}
+%
+
+% use section* for acknowledgement
+\section*{Acknowledgment}
 The authors would like to thank the Fermilab staff,  the FTK group at the 
@@
 
-\section*{References}
-
-%% CHEP2013 recommended style for bibliography
 \bibliographystyle{iopart-num}
-
 \bibliography{gpu}

### 2013-11-13 papers/2013/NSS/NSS.tex
some fixes

@@
 \usepackage{subfigure}
+% \usepackage[cmex10]{amsmath}
+% \usepackage{cite}
 
 \usepackage{hyperref}
-\usepackage{multirow}
-\usepackage[all]{hypcap} % fix links to figs/tabs
+% \usepackage{multirow}
+% \usepackage[all]{hypcap} % fix links to figs/tabs
 
@@
 % paper title
-\title{Title of NSS Talk}
+\title{Applications of Many-Core Technologies to On-line Event Reconstruction in High Energy Physics
+Experiments}
 % Author list
@@
 \section{Introduction}
-Real-time event reconstruction plays a fundamental role in High Energy
+\IEEEPARstart{R}{eal}-time event reconstruction plays a fundamental role in High Energy
 Physics (HEP) experiments at hadron colliders.  Reducing the rate of
@@
 
-\begin{table}[!t]
+\begin{table*}[tbp]
+  \caption{Capabilities of the many-core devices used in this study,
+    according to the manufacturer's specifications. The first three
+    are \textsc{nvidia} GPUs, the MIC 5110P is an Intel Xeon Phi, and the final one is an AMD GPU.}
+  \label{tab_hwspecs}
   \centering
@@
   \end{tabular}
-  \caption{Capabilities of the many-core devices used in this study,
-    according to the manufacturer's specifications. The first three
-    are \textsc{nvidia} GPUs, the MIC 5110P is an Intel Xeon Phi, and the final one is an AMD GPU.}
-  \label{tab_hwspecs}
-\end{table}
+\end{table*}
 
@@
 standard \textsc{c} libraries.
-\begin{figure}[!t]
+\begin{figure}[tbp]
 \centering
@@
 
-\bibliographystyle{iopart-num}
+\bibliographystyle{IEEEtran}
+
 \bibliography{gpu}

### 2013-11-13 papers/2013/NSS/NSS.tex
re-enable packages I commented out while debugging

@@
 \usepackage{subfigure}
-% \usepackage[cmex10]{amsmath}
-% \usepackage{cite}
+\usepackage[cmex10]{amsmath}
+\usepackage{cite}
 
 \usepackage{hyperref}
-% \usepackage{multirow}
-% \usepackage[all]{hypcap} % fix links to figs/tabs
+\usepackage{multirow}
+\usepackage[all]{hypcap} % fix links to figs/tabs
 
@@
 % Author list
-\author{  A.~Gianelle, 
+\author{A.~Gianelle, 
   S.~Amerio, 

### 2013-11-13 papers/2013/NSS/NSS.tex
update IEEEtrans, fix problem with weird first page

@@
   P.~Wittich
-}
 \thanks{Manuscript received November 15, 2013.
-This work was supported by the National Science Foundation, the
-Department of Energy Office of Science and INFN.}% <-this % stops a space
+This work was supported by the  US National Science Foundation, the US 
+Department of Energy Office of Science and the Italian
+Istituto Nazionale di Fisica Nucleare. This work was partially supported by the 
+EU Framework Programme 7 project EURETILE under grant number 247846.}% <-this % stops a space
 \thanks{A.~Gianelle, S.~Amerio, D.~Bastieri, M.~Corvo, and D.~Lucchesi are with INFN Padova.}% 
@@
 \thanks{S.~Poprocki and P.~Wittich are with Cornell University.}%
+}
 % make the title area

### 2013-11-13 papers/2013/NSS/NSS.tex
make bibtex behave better
paragraph on OpenCL added

@@
 \thanks{T.~Liu and R.~Rivera are with Fermi National Accelerator Laboratory.}%
-\thanks{A.~Lonardo, L.~Tosoratto and P.~Vicini are with INFN Roma}% 
-\thanks{S.~Poprocki and P.~Wittich are with Cornell University.}%
+\thanks{A.~Lonardo, L.~Tosoratto and P.~Vicini are with INFN Roma.}% 
+\thanks{S.~Poprocki and P.~Wittich are with Cornell University.}
 }
@@
 \subsection{Data processing in \textsc{opencl}}
-RESULTS FROM THE FERMILAB SETUP.
+We also measure the track fitting algorithm latency in an implementation  
+using \textsc{opencl}. Due to limitations on the available space for storing data on the GPU, we only test the 
+algorithm on up to 1000 events (\textit{i.e.} about 2 million combinations to fit). In Fig.~\ref{fig:opencl_timing} 
+we show the results of algorithm latency measurements for three different modes: running the serial algorithm  
+on the CPU (single-core), running the \textsc{opencl} algorithm on the CPU (8-core), and running the \textsc{opencl} algorithm
+on the AMD Radeon GPU. The \textsc{opencl} implementation of the algorithm on the CPU provides a significant speedup---about 
+a factor of 5---over running the algorithm serially on the CPU. Running the same \textsc{opencl} algorithm on the GPU provides an 
+even greater speedup due to the increased number of cores and parallel threads that can be run, though there is additional latency to copy data 
+into and out of the GPU that makes running on the GPU take longer for small numbers of combinations to fit.
+\begin{figure}[!t]
+  \centering
+  \includegraphics[width=0.9\linewidth]{figures/Timing_OpenCL}
+  \caption{The timing of the \textsc{opencl} algorithm as a function of the
+    number of track fits. We compare timing running the algorithm serially on a CPU, 
+    using \textsc{opencl} on the CPU, and on an AMD GPU.}
+  \label{fig:opencl_timing}
+\end{figure}
 

### 2013-11-13 papers/2013/NSS/NSS.tex
remove dummy fig line so it can compile
modified intro and added openCL plot

@@
 EU Framework Programme 7 project EURETILE under grant number 247846.}% <-this % stops a space
-\thanks{A.~Gianelle, S.~Amerio, D.~Bastieri, M.~Corvo, and D.~Lucchesi are with INFN Padova.}% 
+\thanks{D.~Bastieri and D.~Lucchesi are with University of Padova and INFN Padova.}% 
+\thanks{A.~Gianelle is with INFN Padova.}% 
+\thanks{M.~Corvo is with University of Ferrara.}% 
+\thanks{S.~Amerio is with University of Padova}% 
 \thanks{W.~Ketchum is with Los Alamos National Laboratory.}%
@@
   growing in High Energy Physics (HEP) experiments. In this paper we describe
-  performance measurements of Graphic Processing Units (GPUs) and Intel Many
-  Integrated Core architecture (MIC) when applied to a typical HEP online task: 
+  performance measurements of many-core devices 
+ %Graphic Processing Units (GPUs) and Intel Many
+  %Integrated Core architecture (MIC) 
+  when applied to a typical HEP online task: 
   the selection of events based on the trajectories of charged particles.
@@
   complexity/performance trade-off in porting existing serial
-  algorithms to many-core devices. Measurements of both data processing 
-  and data transfer latency are shown, considering different I/O strategies 
+  algorithms to many-core devices. We measure performance of
+different architectures (Intel Xeon Phi and AMD GPU's, in addition 
+to NVidia GPU's) and different software
+environments (OpenCL, in addition to NVidia CUDA).
+Measurements of both data processing  and data transfer latency are shown, 
+considering different I/O strategies 
   to/from the parallel devices. \end{abstract}
@@
   \centering
-  \includegraphics[width=0.9\linewidth]{figures/Timing_OpenCL}
+  \includegraphics[width=0.9\linewidth]{figures/TotalTimeOpencl_1000}
   \caption{The timing of the \textsc{opencl} algorithm as a function of the

### 2013-11-13 papers/2013/NSS/NSS.tex
minor adds for the setup and conclusions

@@
 
-ADDITIONAL PARAGRAPH ON THE FERMILAB SETUP.
+We have an additional setup for testing the \textsc{opencl} implementation of the track 
+fitting algorithm. On this setup, we use a 3.07 GHz Intel Core i7 CPU 950, which has four cores 
+and up to eight computation threads. We can run the algorithm serially on this CPU (one core), or 
+run the \textsc{opencl} algorithm and make use of the multi-core architecture. Additionally, we have an 
+AMD Radeon HD 7970 GPU in a PCIe slot in this setup, on which we can also run the \textsc{opencl} algorithm.
 
@@
 we show the results of algorithm latency measurements for three different modes: running the serial algorithm  
-on the CPU (single-core), running the \textsc{opencl} algorithm on the CPU (8-core), and running the \textsc{opencl} algorithm
+on the CPU (single-core), running the \textsc{opencl} algorithm on the CPU (multi-core), and running the \textsc{opencl} algorithm
 on the AMD Radeon GPU. The \textsc{opencl} implementation of the algorithm on the CPU provides a significant speedup---about 
@@
 with GPUs and a more complete event-level parallelization using
-\textsc{cuda}.  We have updated latency studies and shown that for larger packet sizes 
+\textsc{cuda}.  Additionally, we have implemented a parallelized algorithm using 
+\textsc{opencl} and we have run that algorithm on a multi-core CPU and GPU, showing a boost
+in performance over serial CPU computation for even small numbers of events. 
+We have also updated latency studies and shown that for larger packet sizes 
 (greater than 600 kB), \textsc{cuda}-aware MPI outperforms P2P. This result fully agrees with expectations 

### 2013-11-14 papers/2013/NSS/NSS.tex
interim edits

@@
 \begin{abstract}
-  Interest in parallel architectures applied to real time selections is 
+  Interest in many-core architectures applied to real time selections is 
   growing in High Energy Physics (HEP) experiments. In this paper we describe
   performance measurements of many-core devices 
- %Graphic Processing Units (GPUs) and Intel Many
-  %Integrated Core architecture (MIC) 
   when applied to a typical HEP online task: 
@@
   algorithms to many-core devices. We measure performance of
-different architectures (Intel Xeon Phi and AMD GPU's, in addition 
-to NVidia GPU's) and different software
+different architectures (Intel Xeon Phi and AMD GPUs, in addition 
+to NVidia GPUs) and different software
 environments (OpenCL, in addition to NVidia CUDA).
@@
 considering different I/O strategies 
-  to/from the parallel devices. \end{abstract}
+  to/from the many-core devices. \end{abstract}
 
@@
 Physics (HEP) experiments at hadron colliders.  Reducing the rate of
-data to be saved on tape from millions to hundreds of events per
-second is critical. To increase the purity of the collected samples,
+data to be saved on tape is critical. To increase the purity of the collected samples,
 rate reduction has to be coupled with an initial selection of the most
@@
 event rate has to be reduced from tens of MHz to a few kHz.  The
-selection system (trigger) is usually organized in multiple levels,
+selection system (trigger) is usually organized in successive levels,
 each capable of performing a finer selection on more complex physics
@@
 environments ($\approx 100~\mu$s) in previous
-papers~\cite{TIPP2011,NSS2012}.  In this paper we extend those studies
-to include other many-core architectures (Intel MIC in addition to
-GPUs).  The algorithm run on the parallel architecture is a complete
-version of the fast track-fitting algorithm of the Silicon Vertex
-Tracker (SVT) system at CDF~\cite{SVT1}.
-We also investigate the performance of the algorithm on a GPU manufactured by AMD using a different
-computing environment: \textsc{opencl}.
-Starting with a serial algorithm
-implemented on a CPU, we test an \textit{embarrassingly parallel}
-algorithm on the Intel MIC environment. In this case each event is
-handled independently by a core on the accelerator, and the
-parallelization is achieved with only minor changes to the legacy code
-and is only possible in the Intel MIC environment. Next we consider an
-algorithm where we unroll three internal nested loops and run these in
-parallel on a GPU, using the \textsc{cuda} environment. This second
-approach is programmatically more complicated and less trivial to
-implement. In neither case have we re-thought the basic algorithms or
-the data structures used. To achieve optimal performance, these steps
-would have to be taken.  As one might expect, the improvement from the
-first approach is rather modest, and the second approach shows larger
-performance gains. For GPUs, we also test different strategies to transfer
-data to and from the device.
+papers~\cite{TIPP2011,NSS2012,CHEP2013}.  In this paper we extend
+those studies to include other many-core architectures (Intel MIC in
+addition to GPUs) and other programming toolsets (\textsc{opencl}).
+The algorithm run on the parallel architecture is a complete version
+of the fast track-fitting algorithm of the Silicon Vertex Tracker
+(SVT) system at CDF~\cite{SVT1}.  We also investigate the performance
+of the algorithm on a GPU manufactured by AMD using a different
+computing environment: \textsc{opencl}.  Starting with a serial
+algorithm implemented on a CPU, we test an \textit{embarrassingly
+  parallel} algorithm on the Intel MIC environment. In this case each
+event is handled independently by a core on the accelerator, and the
+parallelization is achieved with only minor changes to the legacy
+code.  This approach is only possible in the Intel MIC
+environment. Next we consider an algorithm where we unroll three
+internal nested loops and run these in parallel on a GPU, using the
+\textsc{cuda} and \textsc{opencl} environments. This second approach
+is programmatically more complicated and less trivial to implement. In
+neither case have we re-thought the basic algorithms or the data
+structures used. To achieve optimal performance, these steps would
+have to be taken.  As one might expect, the improvement from the first
+approach is rather modest, albeit easier to implement, and the second
+approach shows larger performance gains. For GPUs, we also test
+different strategies to transfer data to and from the device.
 
@@
 
-To be run on NVidia GPUs, the code has been ported to \textsc{cuda}: each step --
-unpack, combine and track fit -- is performed by a specific kernel; the
-three nested loops are unrolled so that each GPU thread processes a
-single combination of hits.  The \textsc{cuda} implementation makes use of 
-\textsc{thrust}~\cite{bib_thrust}, a C++ template library for \textsc{cuda}, in the unpacking step.
+To be run on NVidia GPUs, the code has been ported to \textsc{cuda}:
+each step -- unpack, combine and track fit -- is performed by a
+specific kernel; the three nested loops are unrolled so that each GPU
+thread processes a single combination of hits.  The \textsc{cuda}
+implementation makes use of \textsc{thrust}~\cite{bib_thrust}, a C++
+template library for \textsc{cuda}, in the unpacking step. The
+existence of template libraries such as \textsc{thrust} is an
+advantage of the \textsc{cuda} environment.
 
-To implement the algorithm to run on an AMD GPU, we have ported the \textit{combine} and \textit{track fit}
-\textsc{cuda} kernels to \textsc{opencl}, which requires minimal changes. Because the \textsc{thrust} template libraries
-can only be used with \textsc{cuda}, we resort to unpacking serially on the CPU.
+To implement the algorithm to run on an AMD GPU, we have ported the
+\textit{combine} and \textit{track fit} \textsc{cuda} kernels to
+\textsc{opencl}, which requires minimal changes. Because the
+\textsc{thrust} template libraries can only be used with
+\textsc{cuda}, we resort to unpacking serially on the CPU.
 
-To run on MIC, where cores are more powerful 
-but fewer in number, we adopted the so-called
-\textit{embarrassingly parallel} approach and used \textsc{pragma
-  OpenMP} for statements to unroll only the external loop on the events, so
-that each core processes a single event: the porting requires much
-less effort compared to \textsc{cuda}, but the level of parallelism is
-limited.
+To run on MIC, where cores are more powerful but fewer in number, we
+adopted the so-called \textit{embarrassingly parallel} approach and
+used \textsc{pragma OpenMP} for statements to unroll only the external
+loop on the events, so that each core processes a single event: the
+porting requires much less effort compared to \textsc{cuda}, but the
+level of parallelism is limited.
 
@@
 Table~\ref{tab_hwspecs}. The GPUs include a less expensive gaming
-class GPUs (the NVidia GTX and AMD Radeon cards) and ones optimized for scientific computing
-(Tesla). The MIC corresponds to a Xeon Phi introduced in November 2012.
+class GPUs (the NVidia GTX and AMD Radeon cards) and ones optimized
+for scientific computing (Tesla). The MIC corresponds to a Xeon Phi
+introduced in November 2012.
 
@@
     according to the manufacturer's specifications. The first three
-    are \textsc{nvidia} GPUs, the MIC 5110P is an Intel Xeon Phi, and the final one is an AMD GPU.}
+    are \textsc{nvidia} GPUs, the MIC 5110P is an Intel Xeon Phi, and
+    the final one is an AMD GPU.  For Xeon Phi, the ``cores'' column
+    counts the HW threads per core as equivalent to a GPU core.}
   \label{tab_hwspecs}
@@
 
-We have an additional setup for testing the \textsc{opencl} implementation of the track 
-fitting algorithm. On this setup, we use a 3.07 GHz Intel Core i7 CPU 950, which has four cores 
-and up to eight computation threads. We can run the algorithm serially on this CPU (one core), or 
-run the \textsc{opencl} algorithm and make use of the multi-core architecture. Additionally, we have an 
-AMD Radeon HD 7970 GPU in a PCIe slot in this setup, on which we can also run the \textsc{opencl} algorithm.
+We have an additional setup for testing the \textsc{opencl}
+implementation of the track fitting algorithm. Here, we use a 3.07 GHz
+Intel Core i7 CPU 950, which has four cores and up to eight
+computation threads. We run the algorithm serially on this CPU (one
+core), and also run a CPU-based \textsc{opencl} algorithm which makes
+use of the multi-core architecture. Additionally, we have an AMD
+Radeon HD 7970 GPU in a PCIe slot in this setup, on which we also run
+the \textsc{opencl} algorithm.
 
@@
 explore different data-taking conditions, the number of events ranges
-from one to 3000, \emph{i.e.,} from 2048 to about six millions of combinations
-to fit.
+from one to 3000, \emph{i.e.,} between 2048 to about six millions of
+combinations to fit.
 
@@
 \end{figure}
-Accelerators performance drop with decreasing number of fits, as can be seen 
-in Fig.~\ref{fig:algo_only_timing_zoom}.
+The accelerator card's performance drop with decreasing number of fits, as can be seen 
+in Fig.~\ref{fig:algo_only_timing_zoom}, due to overhead.
 Figure~\ref{fig:algo_only_speedup} shows the speed-up with respect to the serial 

### 2013-11-14 papers/2013/NSS/NSS.tex
some more english updates

@@
 \subsection{Data processing in \textsc{opencl}}
-We also measure the track fitting algorithm latency in an implementation  
-using \textsc{opencl}. Due to limitations on the available space for storing data on the GPU, we only test the 
-algorithm on up to 1000 events (\textit{i.e.} about 2 million combinations to fit). In Fig.~\ref{fig:opencl_timing} 
-we show the results of algorithm latency measurements for three different modes: running the serial algorithm  
-on the CPU (single-core), running the \textsc{opencl} algorithm on the CPU (multi-core), and running the \textsc{opencl} algorithm
-on the AMD Radeon GPU. The \textsc{opencl} implementation of the algorithm on the CPU provides a significant speedup---about 
-a factor of 5---over running the algorithm serially on the CPU. Running the same \textsc{opencl} algorithm on the GPU provides an 
-even greater speedup due to the increased number of cores and parallel threads that can be run, though there is additional latency to copy data 
-into and out of the GPU that makes running on the GPU take longer for small numbers of combinations to fit.
+We also measure the track fitting algorithm latency in an
+implementation using \textsc{opencl}. The \textsc{opencl} tools have
+the advantage of being an open standard, while the \textsc{cuda} tools
+are only compatible with Nvidia GPUs. Additionally, \textsc{opencl}
+can work also on CPUs and on the Xeon Phi. For our tests, due to
+limitations on the available space for storing data on the GPU, we
+only test the algorithm on up to 1000 events (\textit{i.e.} about 2
+million combinations to fit). In Fig.~\ref{fig:opencl_timing} we show
+the results of algorithm latency measurements for three different
+modes: running the serial algorithm on the CPU (single-core), running
+the \textsc{opencl} algorithm on the CPU (multi-core), and running the
+\textsc{opencl} algorithm on the AMD Radeon GPU. The \textsc{opencl}
+implementation of the algorithm on the CPU provides a significant
+speedup---about a factor of 5---over running the algorithm serially on
+the CPU. Running the same \textsc{opencl} algorithm on the GPU
+provides an even greater speedup due to the increased number of cores
+and parallel threads that can be run, though there is additional
+latency to copy data into and out of the GPU that makes running on the
+GPU take longer for small numbers of combinations to fit.
 \begin{figure}[!t]
@@
 on GPUs and Intel MIC. We examined a staged approach to using
-accelerator cards in a hadron collider trigger application. We have
-demonstrated that in this application, significant gains can be
-achieved with the `embarrassingly parallel' approach on an Intel MIC
-architecture, with the smallest amount of required changes to an
-existing serial code base. However, better performance is achieved
-with GPUs and a more complete event-level parallelization using
-\textsc{cuda}.  Additionally, we have implemented a parallelized algorithm using 
-\textsc{opencl} and we have run that algorithm on a multi-core CPU and GPU, showing a boost
-in performance over serial CPU computation for even small numbers of events. 
-We have also updated latency studies and shown that for larger packet sizes 
-(greater than 600 kB), \textsc{cuda}-aware MPI outperforms P2P. This result fully agrees with expectations 
-since the P2P mechanism is optimized (and effective) to reduce the transfer latency for 
-small size packets.
-%We have updated latency studies and shown that for
-%larger packet sizes (greater than 600~kB), \textsc{cuda}-aware MPI
-%outperforms P2P, contrary to naive expectation.  
-Even at large packet
-size, the data transfer takes an appreciable fraction of the total
-algorithm time (about 20-25\%).
+accelerator cards in a hadron collider trigger application and
+compared additional software tool sets. We have demonstrated that in
+this application, significant gains can be achieved with the
+`embarrassingly parallel' approach on an Intel MIC architecture, with
+the smallest amount of required changes to an existing serial code
+base. Additionally, we have implemented a parallelized algorithm using
+\textsc{cuda} and \textsc{opencl}. Better performance is achieved with
+GPUs and a more complete event-level parallelization using these
+tools.  We have run the parallelized algorithm on a multi-core CPU and
+GPU, showing a boost in performance over serial CPU computation for
+even small numbers of events.  We have also updated latency studies
+and shown that for larger packet sizes (greater than 600 kB),
+\textsc{cuda}-aware MPI outperforms P2P.  Even at large packet size,
+the data transfer takes an appreciable fraction of the total algorithm
+time (about 20-25\%).
 

### 2013-11-18 papers/2013/NSS/NSS.tex
add comment about opencl failing on nvidia

@@
 \end{figure}
+Surprisingly, though \textsc{opencl} is an open standard, we find that
+we are unable to run the same \textsc{opencl} code on Nvidia GPUs. On two
+different installations with two different video cards, we find
+incorrect results running the code that ran successfully on the AMD
+card. At present it is unclear what the root cause  is.
 

### 2013-11-19 papers/2013/NSS/NSS.tex
Fixed some typos

@@
 \subsection{Code implementation}
-The starting point of of our studies is the SVT track fitting
+The starting point of our studies is the SVT track fitting
 simulation code, written in the \textsc{c} language. SVT track fitting
@@
 the source of the data, or an upstream trigger processor, as
-the ultimate sink of the data, while the the receiver is the
+the ultimate sink of the data, while the receiver is the
 trigger system: the time to transfer data to the receiver is thus a
@@
     MIC (embarrassingly parallel), and GPUs (fully parallel), in
-    blue, greed, and red, respectively. The GPUs exhibit the best
+    blue, green, and red, respectively. The GPUs exhibit the best
     performance due to the full parallelization. }

### 2013-11-20 papers/2013/NSS/NSS.tex
corrected a few typos

@@
   P.~Wittich
-\thanks{Manuscript received November 15, 2013.
+\thanks{Manuscript received November 22, 2013.
 This work was supported by the  US National Science Foundation, the US 
@@
 \thanks{M.~Corvo is with University of Ferrara.}% 
-\thanks{S.~Amerio is with University of Padova}% 
+\thanks{S.~Amerio is with University of Padova.}% 
 \thanks{W.~Ketchum is with Los Alamos National Laboratory.}%
@@
   algorithms to many-core devices. We measure performance of
-different architectures (Intel Xeon Phi and AMD GPUs, in addition 
+different architectures (Intel Xeon Phi and \textsc{amd} GPUs, in addition 
 to NVidia GPUs) and different software
@@
 papers~\cite{TIPP2011,NSS2012,CHEP2013}.  In this paper we extend
-those studies to include other many-core architectures (Intel MIC in
-addition to GPUs) and other programming toolsets (\textsc{opencl}).
+those studies to include other many-core architectures (Intel MIC 
+and \textsc{amd} GPUs in
+addition to \textsc{nvidia} GPUs) and other programming 
+toolsets (\textsc{opencl} in addition to \textsc{cuda}).
 The algorithm run on the parallel architecture is a complete version
 of the fast track-fitting algorithm of the Silicon Vertex Tracker
-(SVT) system at CDF~\cite{SVT1}.  We also investigate the performance
-of the algorithm on a GPU manufactured by AMD using a different
-computing environment: \textsc{opencl}.  Starting with a serial
-algorithm implemented on a CPU, we test an \textit{embarrassingly
+(SVT) system at CDF~\cite{SVT1}.  
+%We also investigate the performance
+%of the algorithm on a GPU manufactured by AMD using a different
+%computing environment: \textsc{opencl}.  
+Starting with a serial algorithm implemented on a CPU, 
+we test an \textit{embarrassingly
   parallel} algorithm on the Intel MIC environment. In this case each
@@
 
-To be run on NVidia GPUs, the code has been ported to \textsc{cuda}:
+To be run on \textsc{nvidia} GPUs, the code has been ported to \textsc{cuda}:
 each step -- unpack, combine and track fit -- is performed by a
@@
 
-To implement the algorithm to run on an AMD GPU, we have ported the
+To implement the algorithm to run on an \textsc{amd} GPU, we have ported the
 \textit{combine} and \textit{track fit} \textsc{cuda} kernels to
@@
 Table~\ref{tab_hwspecs}. The GPUs include a less expensive gaming
-class GPUs (the NVidia GTX and AMD Radeon cards) and ones optimized
+class GPUs (the \textsc{nvidia} GTX and \textsc{amd} Radeon cards) and ones optimized
 for scientific computing (Tesla). The MIC corresponds to a Xeon Phi
@@
     are \textsc{nvidia} GPUs, the MIC 5110P is an Intel Xeon Phi, and
-    the final one is an AMD GPU.  For Xeon Phi, the ``cores'' column
+    the final one is an \textsc{amd} GPU.  For Xeon Phi, the ``cores'' column
     counts the HW threads per core as equivalent to a GPU core.}
@@
 core), and also run a CPU-based \textsc{opencl} algorithm which makes
-use of the multi-core architecture. Additionally, we have an AMD
+use of the multi-core architecture. Additionally, we have an \textsc{amd}
 Radeon HD 7970 GPU in a PCIe slot in this setup, on which we also run
@@
 the advantage of being an open standard, while the \textsc{cuda} tools
-are only compatible with Nvidia GPUs. Additionally, \textsc{opencl}
+are only compatible with \textsc{nvidia} GPUs. Additionally, \textsc{opencl}
 can work also on CPUs and on the Xeon Phi. For our tests, due to
@@
 the \textsc{opencl} algorithm on the CPU (multi-core), and running the
-\textsc{opencl} algorithm on the AMD Radeon GPU. The \textsc{opencl}
+\textsc{opencl} algorithm on the \textsc{amd} Radeon GPU. The \textsc{opencl}
 implementation of the algorithm on the CPU provides a significant
@@
     number of track fits. We compare timing running the algorithm serially on a CPU, 
-    using \textsc{opencl} on the CPU, and on an AMD GPU.}
+    using \textsc{opencl} on the CPU, and on an \textsc{amd} GPU.}
   \label{fig:opencl_timing}
@@
 Surprisingly, though \textsc{opencl} is an open standard, we find that
-we are unable to run the same \textsc{opencl} code on Nvidia GPUs. On two
+we are unable to run the same \textsc{opencl} code on \textsc{nvidia} GPUs. On two
 different installations with two different video cards, we find
-incorrect results running the code that ran successfully on the AMD
+incorrect results running the code that ran successfully on the \textsc{amd}
 card. At present it is unclear what the root cause  is.

### 2013-12-13 papers/2013/TNS/GPU_TNS.tex
TNS paper

%% bare_jrnl.tex
%% V1.4
%% 2012/12/27
%% by Michael Shell
%% see http://www.michaelshell.org/
%% for current contact information.
%%
%% This is a skeleton file demonstrating the use of IEEEtran.cls
%% (requires IEEEtran.cls version 1.8 or later) with an IEEE journal paper.
%%
%% Support sites:
%% http://www.michaelshell.org/tex/ieeetran/
%% http://www.ctan.org/tex-archive/macros/latex/contrib/IEEEtran/
%% and
%% http://www.ieee.org/

% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
% *** with production work. IEEE's font choices can trigger bugs that do  ***
% *** not appear when using other class files.                            ***
% The testflow support page is at:
% http://www.michaelshell.org/tex/testflow/

%%*************************************************************************
%% Legal Notice:
%% This code is offered as-is without any warranty either expressed or
%% implied; without even the implied warranty of MERCHANTABILITY or
%% FITNESS FOR A PARTICULAR PURPOSE! 
%% User assumes all risk.
%% In no event shall IEEE or any contributor to this code be liable for
%% any damages or losses, including, but not limited to, incidental,
%% consequential, or any other damages, resulting from the use or misuse
%% of any information contained here.
%%
%% All comments are the opinions of their respective authors and are not
%% necessarily endorsed by the IEEE.
%%
%% This work is distributed under the LaTeX Project Public License (LPPL)
%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
%% distributed and modified. A copy of the LPPL, version 1.3, is included
%% in the base LaTeX documentation of all distributions of LaTeX released
%% 2003/12/01 or later.
%% Retain all contribution notices and credits.
%% ** Modified files should be clearly indicated as such, including  **
%% ** renaming them and changing author support contact information. **
%%
%% File list of work: IEEEtran.cls, IEEEtran_HOWTO.pdf, bare_adv.tex,
%%                    bare_conf.tex, bare_jrnl.tex, bare_jrnl_compsoc.tex,
%%                    bare_jrnl_transmag.tex
%%*************************************************************************

% Note that the a4paper option is mainly intended so that authors in
% countries using A4 can easily print to A4 and see how their papers will
% look in print - the typesetting of the document will not typically be
% affected with changes in paper size (but the bottom and side margins will).
% Use the testflow package mentioned above to verify correct handling of
% both paper sizes by the user's LaTeX system.
%
% Also note that the "draftcls" or "draftclsnofoot", not "draft", option
% should be used if it is desired that the figures are to be displayed in
% draft mode.
%
\documentclass[journal]{IEEEtran}
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[journal]{../sty/IEEEtran}

\usepackage{graphicx}
\usepackage{subfigure}
\usepackage[cmex10]{amsmath}
\usepackage{cite}

\usepackage{hyperref}
\usepackage{multirow}
\usepackage[all]{hypcap} % fix links to figs/tabs

\renewcommand{\arraystretch}{1.1} % make tables more readable

% Some very useful LaTeX packages include:
% (uncomment the ones you want to load)

% *** MISC UTILITY PACKAGES ***
%
%\usepackage{ifpdf}
% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
% compilation based on whether the output is pdf or dvi.
% usage:
% \ifpdf
%   % pdf code
% \else
%   % dvi code
% \fi
% The latest version of ifpdf.sty can be obtained from:
% http://www.ctan.org/tex-archive/macros/latex/contrib/oberdiek/
% Also, note that IEEEtran.cls V1.7 and later provides a builtin
% \ifCLASSINFOpdf conditional that works the same way.
% When switching from latex to pdflatex and vice-versa, the compiler may
% have to be run twice to clear warning/error messages.

% *** CITATION PACKAGES ***
%
%\usepackage{cite}
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off
% such as if a citation ever needs to be enclosed in parenthesis.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 4.0 (2003-05-27) and later if using hyperref.sty. cite.sty does
% not currently provide for hyperlinked citations.
% The latest version can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/cite/
% The documentation is contained in the cite.sty file itself.

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
  % \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  % \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation
% can be obtained at: 
% http://www.ctan.org/tex-archive/macros/latex/required/graphics/
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found at:
% http://www.ctan.org/tex-archive/info/epslatex/
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

% *** MATH PACKAGES ***
%
%\usepackage[cmex10]{amsmath}
% A popular package from the American Mathematical Society that provides
% many useful and powerful commands for dealing with mathematics. If using
% it, be sure to load this package with the cmex10 option to ensure that
% only type 1 fonts will utilized at all point sizes. Without this option,
% it is possible that some math symbols, particularly those within
% footnotes, will be rendered in bitmap form which will result in a
% document that can not be IEEE Xplore compliant!
%
% Also, note that the amsmath package sets \interdisplaylinepenalty to 10000
% thus preventing page breaks from occurring within multiline equations. Use:
%\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks as IEEEtran.cls normally
% does. amsmath.sty is already installed on most LaTeX systems. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/amslatex/math/

% *** SPECIALIZED LIST PACKAGES ***
%
%\usepackage{algorithmic}
% algorithmic.sty was written by Peter Williams and Rogerio Brito.
% This package provides an algorithmic environment fo describing algorithms.
% You can use the algorithmic environment in-text or within a figure
% environment to provide for a floating algorithm. Do NOT use the algorithm
% floating environment provided by algorithm.sty (by the same authors) or
% algorithm2e.sty (by Christophe Fiorio) as IEEE does not use dedicated
% algorithm float types and packages that provide these will not provide
% correct IEEE style captions. The latest version and documentation of
% algorithmic.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithms/
% There is also a support site at:
% http://algorithms.berlios.de/index.html
% Also of interest may be the (relatively newer and more customizable)
% algorithmicx.sty package by Szasz Janos:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithmicx/

% *** ALIGNMENT PACKAGES ***
%
%\usepackage{array}
% Frank Mittelbach's and David Carlisle's array.sty patches and improves
% the standard LaTeX2e array and tabular environments to provide better
% appearance and additional user controls. As the default LaTeX2e table
% generation code is lacking to the point of almost being broken with
% respect to the quality of the end results, all users are strongly
% advised to use an enhanced (at the very least that provided by array.sty)
% set of table tools. array.sty is already installed on most systems. The
% latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/tools/

% IEEEtran contains the IEEEeqnarray family of commands that can be used to
% generate multiline equations as well as matrices, tables, etc., of high
% quality.

% *** SUBFIGURE PACKAGES ***
%\ifCLASSOPTIONcompsoc
%  \usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
%\else
%  \usepackage[caption=false,font=footnotesize]{subfig}
%\fi
% subfig.sty, written by Steven Douglas Cochran, is the modern replacement
% for subfigure.sty, the latter of which is no longer maintained and is
% incompatible with some LaTeX packages including fixltx2e. However,
% subfig.sty requires and automatically loads Axel Sommerfeldt's caption.sty
% which will override IEEEtran.cls' handling of captions and this will result
% in non-IEEE style figure/table captions. To prevent this problem, be sure
% and invoke subfig.sty's "caption=false" package option (available since
% subfig.sty version 1.3, 2005/06/28) as this is will preserve IEEEtran.cls
% handling of captions.
% Note that the Computer Society format requires a larger sans serif font
% than the serif footnote size font used in traditional IEEE formatting
% and thus the need to invoke different subfig.sty package options depending
% on whether compsoc mode has been enabled.
%
% The latest version and documentation of subfig.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/subfig/

% *** FLOAT PACKAGES ***
%
%\usepackage{fixltx2e}
% fixltx2e, the successor to the earlier fix2col.sty, was written by
% Frank Mittelbach and David Carlisle. This package corrects a few problems
% in the LaTeX2e kernel, the most notable of which is that in current
% LaTeX2e releases, the ordering of single and double column floats is not
% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
% single column figure to be placed prior to an earlier double column
% figure. The latest version and documentation can be found at:
% http://www.ctan.org/tex-archive/macros/latex/base/

%\usepackage{stfloats}
% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
% the ability to do double column floats at the bottom of the page as well
% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command:
%\fnbelowfloat
% to enable the placement of footnotes below bottom floats (the standard
% LaTeX2e kernel puts them above bottom floats). This is an invasive package
% which rewrites many portions of the LaTeX2e float routines. It may not work
% with other packages that modify the LaTeX2e float routines. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/sttools/
% Do not use the stfloats baselinefloat ability as IEEE does not allow
% \baselineskip to stretch. Authors submitting work to the IEEE should note
% that IEEE rarely uses double column equations and that authors should try
% to avoid such use. Do not be tempted to use the cuted.sty or midfloat.sty
% packages (also by Sigitas Tolusis) as IEEE does not format its papers in
% such ways.
% Do not attempt to use stfloats with fixltx2e as they are incompatible.
% Instead, use Morten Hogholm'a dblfloatfix which combines the features
% of both fixltx2e and stfloats:
%
% \usepackage{dblfloatfix}
% The latest version can be found at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/dblfloatfix/

%\ifCLASSOPTIONcaptionsoff
%  \usepackage[nomarkers]{endfloat}
% \let\MYoriglatexcaption\caption
% \renewcommand{\caption}[2][\relax]{\MYoriglatexcaption[#2]{#2}}
%\fi
% endfloat.sty was written by James Darrell McCauley, Jeff Goldberg and 
% Axel Sommerfeldt. This package may be useful when used in conjunction with 
% IEEEtran.cls'  captionsoff option. Some IEEE journals/societies require that
% submissions have lists of figures/tables at the end of the paper and that
% figures/tables without any captions are placed on a page by themselves at
% the end of the document. If needed, the draftcls IEEEtran class option or
% \CLASSINPUTbaselinestretch interface can be used to increase the line
% spacing as well. Be sure and use the nomarkers option of endfloat to
% prevent endfloat from "marking" where the figures would have been placed
% in the text. The two hack lines of code above are a slight modification of
% that suggested by in the endfloat docs (section 8.4.1) to ensure that
% the full captions always appear in the list of figures/tables - even if
% the user used the short optional argument of \caption[]{}.
% IEEE papers do not typically make use of \caption[]'s optional argument,
% so this should not be an issue. A similar trick can be used to disable
% captions of packages such as subfig.sty that lack options to turn off
% the subcaptions:
% For subfig.sty:
% \let\MYorigsubfloat\subfloat
% \renewcommand{\subfloat}[2][\relax]{\MYorigsubfloat[]{#2}}
% However, the above trick will not work if both optional arguments of
% the \subfloat command are used. Furthermore, there needs to be a
% description of each subfigure *somewhere* and endfloat does not add
% subfigure captions to its list of figures. Thus, the best approach is to
% avoid the use of subfigure captions (many IEEE journals avoid them anyway)
% and instead reference/explain all the subfigures within the main caption.
% The latest version of endfloat.sty and its documentation can obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/endfloat/
%
% The IEEEtran \ifCLASSOPTIONcaptionsoff conditional can also be used
% later in the document, say, to conditionally put the References on a 
% page by themselves.

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
%\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/url/
% Basically, \url{my_url_here}.

% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.
% (Unless specifically asked to do so by the journal or conference you plan
% to submit to, of course. )

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
% Do not put math or special symbols in the title.
\title{Applications of Many-Core Technologies to On-line Event Reconstruction in High Energy Physics
Experiments}
%
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
%

% note the % following the last \IEEEmembership and also \thanks - 
% these prevent an unwanted space from occurring between the last author name
% and the end of the author line. i.e., if you had this:
% 
% \author{....lastname \thanks{...} \thanks{...} }
%                     ^------------^------------^----Do not want these spaces!
%
% a space would be appended to the last name and could cause every name on that
% line to be shifted left slightly. This is one of those "LaTeX things". For
% instance, "\textbf{A} \textbf{B}" will typeset as "A B" not "AB". To get
% "AB" then you have to do: "\textbf{A}\textbf{B}"
% \thanks is no different in this regard, so shield the last } of each \thanks
% that ends a line with a % and do not let a space in before the next \thanks.
% Spaces after \IEEEmembership other than the last one are OK (and needed) as
% you are supposed to have spaces between the names. For what it is worth,
% this is a minor point as most people would not even notice if the said evil
% space somehow managed to creep in.

% paper title
\title{Applications of Many-Core Technologies to On-line Event Reconstruction in High Energy Physics
Experiments}
% Author list
\author{A.~Gianelle, 
  S.~Amerio, 
  D.~Bastieri, 
  M.~Corvo, 
  W.~Ketchum,
  T.~Liu, 
  A.~Lonardo, 
  D.~Lucchesi,
  S.~Poprocki, 
  R.~Rivera, 
  L.~Tosoratto,
  P.~Vicini
  and 
  P.~Wittich
\thanks{Manuscript received XXX, XXX.
This work was supported by the  US National Science Foundation, the US 
Department of Energy Office of Science and the Italian
Istituto Nazionale di Fisica Nucleare. This work was partially supported by the 
EU Framework Programme 7 project EURETILE under grant number 247846.}% <-this % stops a space
\thanks{D.~Bastieri and D.~Lucchesi are with University of Padova and INFN Padova.}% 
\thanks{A.~Gianelle is with INFN Padova.}% 
\thanks{M.~Corvo is with University of Ferrara.}% 
\thanks{S.~Amerio is with University of Padova.}% 
\thanks{W.~Ketchum is with Los Alamos National Laboratory.}%
\thanks{T.~Liu and R.~Rivera are with Fermi National Accelerator Laboratory.}%
\thanks{A.~Lonardo, L.~Tosoratto and P.~Vicini are with INFN Roma.}% 
\thanks{S.~Poprocki and P.~Wittich are with Cornell University.}
}

% The paper headers
%\markboth{Journal of \LaTeX\ Class Files,~Vol.~11, No.~4, December~2012}%
%{Shell \MakeLowercase{\textit{et al.}}: Bare Demo of IEEEtran.cls for Journals}
% The only time the second header will appear is for the odd numbered pages
% after the title page when using the twoside option.
% 
% *** Note that you probably will NOT want to include the author's ***
% *** name in the headers of peer review papers.                   ***
% You can use \ifCLASSOPTIONpeerreview for conditional compilation here if
% you desire.

% If you want to put a publisher's ID mark on the page you can do it like
% this:
%\IEEEpubid{0000--0000/00\$00.00~\copyright~2012 IEEE}
% Remember, if you use this you must call \IEEEpubidadjcol in the second
% column for its text to clear the IEEEpubid mark.

% use for special paper notices
%\IEEEspecialpapernotice{(Invited Paper)}

% make the title area
\maketitle

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
The abstract goes here.
\end{abstract}

% Note that keywords are not normally used for peerreview papers.
%\begin{IEEEkeywords}
%IEEEtran, journal, \LaTeX, paper, template.
%\end{IEEEkeywords}

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

\section{Introduction}

\section{SVT track fitting algorithm}
\subsection{Code implementation}
\subsection{Hardware used}

\section{Data processing results}

\section{Data processing vs power consumption}

\section{Data transfer}

\section{Conclusion}
The conclusion goes here.

% if have a single appendix:
%\appendix[Proof of the Zonklar Equations]
% or
%\appendix  % for no appendix heading
% do not use \section anymore after \appendix, only \section*
% is possibly needed

% use appendices with more than one appendix
% then use \section to start each appendix
% you must declare a \section before using any
% \subsection or using \label (\appendices by itself
% starts a section numbered zero.)
%

% use section* for acknowledgement
\section*{Acknowledgment}

The authors would like to thank...

% Can use something like this to put references on a page
% by themselves when using endfloat and the captionsoff option.
\ifCLASSOPTIONcaptionsoff
  \newpage
\fi

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
%\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
%\bibliography{IEEEabrv,../bib/paper}
%
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)

\bibliographystyle{IEEEtran}

\bibliography{gpu}

% that's all folks
\end{document}

### 2014-02-14 papers/2013/TNS/GPU_TNS.tex
Added text from NSS proceedings and kayla measurements

@@
 \section{Introduction}
+\IEEEPARstart{R}{eal}-time event reconstruction plays a fundamental role in High Energy
+Physics (HEP) experiments at hadron colliders.  Reducing the rate of
+data to be saved on tape is critical. To increase the purity of the collected samples,
+rate reduction has to be coupled with an initial selection of the most
+interesting events.  In a typical hadron collider experiment, the
+event rate has to be reduced from tens of MHz to a few kHz.  The
+selection system (trigger) is usually organized in successive levels,
+each capable of performing a finer selection on more complex physics
+objects describing the event. Trigger systems usually comprise a first
+level based on custom hardware, followed by one or two levels usually
+based on farms of general purpose processors.  At all levels, latency
+is a concern: for a fixed processing time, the faster a decision is
+rendered about accepting or rejecting an event improves the purity of
+the collected data sample. The possibility of using commercial devices
+at a low trigger level is very appealing: they are subject to
+continuous performance improvements driven by the consumer market, are
+less expensive than dedicated hardware, and are easier to support.
+Among the commercial devices, many-core architectures such as Graphic
+Processing Units (GPUs)~\cite{bib_gpu} and Intel Many Integrated Core
+(MIC)~\cite{bib_intelMIC} are of particular interest for online
+selections given their great computing power: the latest
+\textsc{nvidia}~\cite{bib_nvidia} GPU architecture, Kepler, exceeds
+Teraflop computing power. Moreover, high-level programming
+architectures based on \textsc{c/c++} such as
+\textsc{cuda}~\cite{bib_cuda} and \textsc{opencl}~\cite{bib_opencl}
+make programming these devices more accessible to the general
+physicist user.  The goal of this study is to investigate the
+strengths and weaknesses of many-core devices when applied in a low
+latency environment, with particular emphasis on the data transfer
+latency to/from the device and the algorithm latency for processing on
+the device in a manner similar to a typical HEP trigger application,
+and to understand the cost/complexity ratio of porting legacy serial
+code to many-core devices.
+
+We showed initial studies on GPU performance in low-latency
+environments ($\approx 100~\mu$s) in previous
+papers~\cite{TIPP2011,NSS2012,CHEP2013}.  In this paper we extend
+those studies to include other many-core architectures (Intel MIC 
+and \textsc{amd} GPUs in
+addition to \textsc{nvidia} GPUs) and other programming 
+toolsets (\textsc{opencl} in addition to \textsc{cuda}).
+The algorithm run on the parallel architecture is a complete version
+of the fast track-fitting algorithm of the Silicon Vertex Tracker
+(SVT) system at CDF~\cite{SVT1}.  
+%We also investigate the performance
+%of the algorithm on a GPU manufactured by AMD using a different
+%computing environment: \textsc{opencl}.  
+Starting with a serial algorithm implemented on a CPU, 
+we test an \textit{embarrassingly
+  parallel} algorithm on the Intel MIC environment. In this case each
+event is handled independently by a core on the accelerator, and the
+parallelization is achieved with only minor changes to the legacy
+code.  This approach is only possible in the Intel MIC
+environment. Next we consider an algorithm where we unroll three
+internal nested loops and run these in parallel on a GPU, using the
+\textsc{cuda} and \textsc{opencl} environments. This second approach
+is programmatically more complicated and less trivial to implement. In
+neither case have we re-thought the basic algorithms or the data
+structures used. To achieve optimal performance, these steps would
+have to be taken.  As one might expect, the improvement from the first
+approach is rather modest, albeit easier to implement, and the second
+approach shows larger performance gains. For GPUs, we also test
+different strategies to transfer data to and from the device.
 
 \section{SVT track fitting algorithm}
+The Silicon Vertex Trigger (SVT)~\cite{SVT1,SVT2} is a track
+reconstruction processor used in the CDF experiment at Tevatron
+accelerator. It reconstructs tracks in about 20 $\mu s$ in two steps:
+first, low resolution tracks (\textit{roads}) are found in each event
+among the energy deposits left in the tracking detector by charged
+particles; second, track fitting is performed on all possible
+combinations of hits inside a road.  This algorithm uses a linearized
+approximation to track-fitting as implemented in hardware (described
+in greater detail in~\cite{SVT3}).  With the linearized track fit of
+the SVT approach, the determination of the track parameters ($p_i$) is
+reduced to a simple scalar product:
+\[
+p_i = \vec{f_i} \cdot \vec{x_i} + q_i,
+\]
+where $\vec{x_i}$ are input silicon hits, and $\vec{f_i}$ and $q_i$ are 
+pre-defined constant sets. For each set of hits, the algorithm
+computes the impact parameter $d_0$, the azimuthal angle $\phi$, 
+the transverse momentum $p_\mathrm{T}$ , and the $\chi^2$ of the
+fitted track by using simple operations such as memory lookup and 
+integer addition and multiplication.
+
+We ported the track fitting as it is well suited to parallelization -
+each track can be handled independently.
+
 \subsection{Code implementation}
-\subsection{Hardware used}
+The starting point of our studies is the SVT track fitting
+simulation code, written in the \textsc{c} language. SVT track fitting
+is divided into three main functions: first, the unpacking of input
+data and filling of all the necessary data structures; second, the
+computation of all possible combinations of hits in each road and
+third, the linearized track fit of each combination of hits. Three
+main loops are present - on events, roads and hit combinations.
 
-\section{Data processing results}
+To be run on \textsc{nvidia} GPUs, the code has been ported to \textsc{cuda}:
+each step -- unpack, combine and track fit -- is performed by a
+specific kernel; the three nested loops are unrolled so that each GPU
+thread processes a single combination of hits.  The \textsc{cuda}
+implementation makes use of \textsc{thrust}~\cite{bib_thrust}, a C++
+template library for \textsc{cuda}, in the unpacking step. The
+existence of template libraries such as \textsc{thrust} is an
+advantage of the \textsc{cuda} environment.
 
-\section{Data processing vs power consumption}
+To implement the algorithm to run on an \textsc{amd} GPU, we have ported the
+\textit{combine} and \textit{track fit} \textsc{cuda} kernels to
+\textsc{opencl}, which requires minimal changes. Because the
+\textsc{thrust} template libraries can only be used with
+\textsc{cuda}, we resort to unpacking serially on the CPU.
 
-\section{Data transfer}
+To run on MIC, where cores are more powerful but fewer in number, we
+adopted the so-called \textit{embarrassingly parallel} approach and
+used \textsc{pragma OpenMP} for statements to unroll only the external
+loop on the events, so that each core processes a single event: the
+porting requires much less effort compared to \textsc{cuda}, but the
+level of parallelism is limited.
+
+\subsection{Experimental setups}
+The many-core devices used in this study are listed in
+Table~\ref{tab_hwspecs}. The GPUs include a less expensive gaming
+class GPUs (the \textsc{nvidia} GTX and \textsc{amd} Radeon cards) and ones optimized
+for scientific computing (Tesla). The MIC corresponds to a Xeon Phi
+introduced in November 2012.
+
+\begin{table*}[tbp]
+  \caption{Capabilities of the many-core devices used in this study,
+    according to the manufacturer's specifications. The first three
+    are \textsc{nvidia} GPUs, the MIC 5110P is an Intel Xeon Phi, and
+    the final one is an \textsc{amd} GPU.  For Xeon Phi, the ``cores'' column
+    counts the HW threads per core as equivalent to a GPU core.}
+  \label{tab_hwspecs}
+  \centering
+  \begin{tabular}{|l|c|c|c|c|c|}
+    \hline
+    Model & Tesla M2050 & Tesla K20m & GeForce GTX  Titan & MIC 5110P & Radeon HD 7970 \\
+    \hline
+     \hline
+    Performance (SP, GFlops) & 1030 & 3520 & 4500 & 2022 & 3790\\
+    Memory bandwidth  (GB/s) & 148 & 208 & 288  & 320 & 264\\   
+    Memory size (GB) & 3 & 5 & 6 & 8 & 3\\
+    Number of cores & 448 & 2496 & 2688 & 240 & 2048\\
+    Clock speed (GHz) & 1.15 & 0.706 & 0.837 & 1.053 & 1.375\\
+    \hline
+  \end{tabular}
+\end{table*}
+
+To measure the data transfer latency we use a computing cluster composed of 
+12 identical nodes.  Each node contains a Intel Xeon E6520 2.4 GHz 
+CPU and two Tesla M2075 GPU cards. The nodes are connected by InfiniBand 
+communication links using Connect-X2 Mellanox or APEnet+ adapters. 
+APEnet+ is an FPGA-based PCIe board supporting peer-to-peer communication with 
+Tesla and Kepler cards~\cite{apenet2010}.
+Two nodes of this cluster are used to measure data transfer latency,
+one acting as a transmitter and the
+other as a receiver.  Data are transferred from the transmitter to the 
+receiver, processed on the GPU and sent back to the receiver (see
+Fig.~\ref{fig:data_flow}).  
+The latency for a complete loop is measured on the transmitter using 
+standard \textsc{c} libraries.
+\begin{figure}[tbp]
+\centering
+\includegraphics[width=3.5in]{figures/SetUp-general}
+\caption{Data flow. Data is sent from the transmitter PC to the
+  receiver PC, where it is processed by the GPU before being returned
+  to the transmitter PC. The transmitter plays the role of the
+  detector as the source of the data and as an upstream trigger
+  processor as the data's ultimate sink. The receiver PC plays the
+  role of a component in the trigger system. }
+\label{fig:data_flow}
+\end{figure}
+In this setup, the transmitter can represent the detector, as
+the source of the data, or an upstream trigger processor, as
+the ultimate sink of the data, while the receiver is the
+trigger system: the time to transfer data to the receiver is thus a
+rough estimate of the latency to transfer the data from the detector
+front-end to the trigger system.
+
+We have an additional setup for testing the \textsc{opencl}
+implementation of the track fitting algorithm. Here, we use a 3.07 GHz
+Intel Core i7 CPU 950, which has four cores and up to eight
+computation threads. We run the algorithm serially on this CPU (one
+core), and also run a CPU-based \textsc{opencl} algorithm which makes
+use of the multi-core architecture. Additionally, we have an \textsc{amd}
+Radeon HD 7970 GPU in a PCIe slot in this setup, on which we also run
+the \textsc{opencl} algorithm.
+
+The power measurements are performed on a \textsc{cuda} development kit~\cite{bib_kayla} 
+based on the kayla MITX platform. This kit features a  
+\textsc{nvidia} Tegra 3 \textsc{arm} Cortex A9 Quad-Core and a Quadro K2000 GPU, both low
+energy consumption devices.
+
+\section{Results}
+The input data consists of events with a fixed number of roads and
+combinations: each event has 2048 combinations to be fitted. To
+explore different data-taking conditions, the number of events ranges
+from one to 3000, \emph{i.e.,} between 2048 to about six millions of
+combinations to fit.
+
+\subsection{Data processing results}
+Each data sample is processed 100 times by the track fitting algorithm. 
+The average latency as a function of the number of fits is presented in 
+Fig.~\ref{fig:algo_only_timing} for the serial,
+embarrassingly parallel and parallel algorithms. We see that the
+embarrassingly parallel algorithm gives a modest increase with respect
+to the serial (CPU) algorithm. Switching to a fully parallel algorithm
+affords a much more significant speed improvement. 
+\begin{figure}[!t]
+  \centering
+  \includegraphics[width=0.9\linewidth]{figures/TimeComp_MIC}
+  \caption{Algorithm-only comparison for timing as a function of the
+    number of track fits. We compare timing on CPUs (serial), Intel
+    MIC (embarrassingly parallel), and GPUs (fully parallel), in
+    blue, green, and red, respectively. The GPUs exhibit the best
+    performance due to the full parallelization. }
+  \label{fig:algo_only_timing}
+\end{figure}
+\begin{figure}[!t]
+  \centering
+  \includegraphics[width=0.9\linewidth]{figures/TimeCompZoom_MIC.png} 
+  \caption{Algorithm-only comparison for timing as a function of the
+    number of track fits: zoom in the low number of fits region. At
+    low number of fits, the CPU performs better, due to start-up
+    costs associated with data transfers to the accelerator card.}
+  \label{fig:algo_only_timing_zoom}
+\end{figure}
+\begin{figure}[!t]
+  \centering
+  \includegraphics[width=0.9\linewidth]{figures/Speedup_MIC}
+  \caption{Speed-up with respect to a standard CPU (Intel Xeon
+    E5630). The speed-ups plateau after about two million fits.}
+  \label{fig:algo_only_speedup}
+\end{figure}
+The accelerator card's performance drop with decreasing number of fits, as can be seen 
+in Fig.~\ref{fig:algo_only_timing_zoom}, due to overhead.
+Figure~\ref{fig:algo_only_speedup} shows the speed-up with respect to the serial 
+algorithm run on a standard CPU (Intel Xeon E5630): the maximum gain is 
+obtained processing at least 500 events. This means that to fully exploit 
+parallel architectures millions of fits have to be performed in parallel.
+
+\subsubsection{Breakdown of computing time}
+In Fig.~\ref{fig:breakdown} we show the fractional time spent in
+various parts of the algorithm for the embarrassingly parallel algorithm 
+(on Intel MIC) and the parallel algorithm (on \textsc{nvidia} Titan GPU), as a 
+function of the number of fits. On both accelerator cards the fractional times
+are constant for more than 500 input events, where computing resources are saturated. 
+Unlike the MIC, the fit stage takes most of the time on the GPU: this 
+could be caused by the intense memory access frequency intrinsic to
+this part of the algorithm.  
+\begin{figure*}[!t]
+\centering
+\subfigure[]
+{\label{fig:breakdown_MIC}
+\includegraphics[width=0.45\linewidth]{figures/Mic5_perc_nk}}
+\subfigure[]
+{\label{fig:breakdown_GPU}
+\includegraphics[width=0.45\linewidth]{figures/titan_perc}}
+\caption{Breakdown of computing time for MIC (a) and the GTX Titan GPU
+  (b). White corresponds to unpacking, green hash corresponds to
+  generating hit combinations, solid blue is the linearized track fit,
+  and magenta cross-hatch corresponds to offloading (MIC only). For
+  MIC, combinations and fitting take the same amount of time for large
+  number of events. For GPU, fitting dominates. }
+\label{fig:breakdown}
+\end{figure*}
+
+\subsubsection{Data processing in \textsc{opencl}}
+We also measure the track fitting algorithm latency in an
+implementation using \textsc{opencl}. The \textsc{opencl} tools have
+the advantage of being an open standard, while the \textsc{cuda} tools
+are only compatible with \textsc{nvidia} GPUs. Additionally, \textsc{opencl}
+can work also on CPUs and on the Xeon Phi. For our tests, due to
+limitations on the available space for storing data on the GPU, we
+only test the algorithm on up to 1000 events (\textit{i.e.} about 2
+million combinations to fit). In Fig.~\ref{fig:opencl_timing} we show
+the results of algorithm latency measurements for three different
+modes: running the serial algorithm on the CPU (single-core), running
+the \textsc{opencl} algorithm on the CPU (multi-core), and running the
+\textsc{opencl} algorithm on the \textsc{amd} Radeon GPU. The \textsc{opencl}
+implementation of the algorithm on the CPU provides a significant
+speedup---about a factor of 5---over running the algorithm serially on
+the CPU. Running the same \textsc{opencl} algorithm on the GPU
+provides an even greater speedup due to the increased number of cores
+and parallel threads that can be run, though there is additional
+latency to copy data into and out of the GPU that makes running on the
+GPU take longer for small numbers of combinations to fit.
+\begin{figure}[!t]
+  \centering
+  \includegraphics[width=0.9\linewidth]{figures/openclamd.pdf}
+  \caption{The timing of the \textsc{opencl} algorithm as a function of the
+    number of track fits. We compare timing running the algorithm serially on a CPU, 
+    using \textsc{opencl} on the CPU, and on an \textsc{amd} GPU.}
+  \label{fig:opencl_timing}
+\end{figure}
+Surprisingly, though \textsc{opencl} is an open standard, we find that
+we are unable to run the same \textsc{opencl} code on \textsc{nvidia} GPUs. On two
+different installations with two different video cards, we find
+incorrect results running the code that ran successfully on the \textsc{amd}
+card. At present it is unclear what the root cause  is.
+
+\subsection{Data processing vs power consumption}
+
+\begin{table*}[tbp]
+  \caption{}
+  \label{tab_kayla}
+  \centering
+  \begin{tabular}{|l|c|c|c|}
+    \hline
+         & \multicolumn{3}{|c|}{Absorbed current (A)} \\
+    \hline
+    Mode & ARM & GPU & Motherboard  \\
+    \hline    
+    \hline
+    idle & 0.29 & 0.60-0.70 & 0.60 \\
+    \hline
+    read & 0.29 & 1.01  & 0.60 \\
+    \hline
+    Host $\rightarrow$ Device & 0.33 & 1.08  & 1.17 \\
+    \hline
+    Device $\rightarrow$ Host & 0.33 & 1.06  & 1.14 \\
+    \hline
+    Full algorithm & 0.33 & 0.98  & 2.24 \\
+%    \hline
+%    Full algorithm (no Thrust) & 0.33 & 1.06-1.08  & 1.15 \\
+    \hline
+  \end{tabular}
+\end{table*}
+
+\subsection{Data transfer}
+The experimental setup described in Fig.~\ref{fig:data_flow} allows us to 
+test different data transfer  strategies to the GPU. The standard data transfer 
+strategy is via the system memory, where the 
+PCIe adapter card and the GPU allocate \emph{separate} buffers on the
+system memory for the copy (as shown in Fig.~\ref{fig:standardDT}).
+This is inefficient, as the data are copied twice in the 
+system memory before being transferred to the GPU/PCIe card. 
+Data may also be transferred using Direct Memory Access (DMA, 
+GPUDirect~\cite{bib_GPUDirect}) to the CPU memory:
+the PCIe card and the GPU share the \textit{same} buffer on the CPU memory; as 
+a result the data are copied only once in the CPU memory 
+(Fig.~\ref{fig:GPUDirectV1}). 
+With our experimental setup two additional copy strategies can be tested
+which are the results of different levels of 
+optimization of the GPUDirect protocol:
+\begin{itemize}
+\item \textsc{cuda}-Aware MPI, where the copy latency is further reduced
+by automatically allocating the buffer on the CPU memory;
+\item peer-to-peer (P2P) strategy, when data are transferred 
+directly to the GPU, without any  
+intermediate copy to the CPU (Fig.~\ref{fig:GPUDirectV2}).
+\end{itemize}
+
+\begin{figure*}[!t]
+\centering
+\subfigure[]
+{\label{fig:standardDT}
+\includegraphics[width=0.3\linewidth]{figures/noGPUDirect}}
+\hspace{1mm}
+\subfigure[]
+{\label{fig:GPUDirectV1}
+\includegraphics[width=0.3\linewidth]{figures/GPUDirect}}
+\subfigure[]
+{\label{fig:GPUDirectV2}
+\includegraphics[width=0.3\linewidth]{figures/GPUDirect-APE}}
+\caption{Standard data transfer (a), via GPUDirect (b) and via
+  GPUDirect with P2P support (c). In (a), two buffers are required in
+  the main memory. In GPUDirect (b), one of the main memory buffers is
+  eliminated. In GPUDirect with P2P support, data is sent directly
+  from the APEnet+ transceiver to the GPU memory.}
+\end{figure*}
+
+\begin{figure}[!t]
+  \centering
+  \includegraphics[width=0.85\linewidth]{figures/datatransfer}
+  \caption{Total latency (data transfer, copy to/from the GPU and data
+    processing on the GPU) as a function of data buffer sizes, for
+    three different levels of optimization of GPUDirect: v1.0,
+    \textsc{cuda}-aware MPI and P2P. The smallest transferred data
+    packed is 600 kB.  \textsc{cuda}-aware MPI shows the best
+    performance for larger packet size.}
+  \label{fig:xferlatency}
+\end{figure}
+
+In Fig.~\ref{fig:xferlatency} we show the total latency (data
+transfer, copy to/from the GPU and data processing on the GPU) as a
+function of data packet size when data are transferred using GPUDirect
+v1.0, \textsc{cuda}-aware MPI and P2P. For the packet sizes considered in this test 
+ \textsc{cuda}-aware MPI gives the best performance. This is expected as 
+P2P is optimized for small packet sizes
+ (see also~\cite{NSS2012} and~\cite{bib_mvapich}). As a matter of fact, for larger packet size, the channel 
+throughput becomes dominant: the shortest transfer 
+time of CUDA aware-MPI system is easily explained comparing the link bandwidth of Mellanox board (40 Gb/s) 
+with the smaller throughput of a APEnet+ single link (30 Gb/s).  
+The data transfer latency accounts for a significant part of the total
+latency, as can be seen in Fig.~\ref{fig:transferOnly}: about 20-25\%
+of total latency is due to moving the data to and from the GPU.
+
+\begin{figure}[!t]
+  \centering
+  \includegraphics[width=0.85\linewidth]{figures/cudaware}
+  \caption{Time per fit, in msec. The two curves show total timing
+    with and without calculations performed on the GPU, thereby
+    showing the considerable time spent in data transfer. About
+    20-25\% of the time is spent in data transfer.}
+  \label{fig:transferOnly}
+\end{figure}
 

### 2014-02-14 src_opencl/svt_gpu_opencl.cpp
opencl implementation in place ... found some strange bugs though, so nothing going yet

@@ -12,6 +12,7 @@
 #include <fstream>
 #include <iterator>
 #include <stdio.h>
+#include <malloc.h>
 
 const int gf_maskdata[] = {
   0x00000000,
@@ -189,13 +190,43 @@ int main(int argc, char* argv[]) {
   int ow;
 
 
-  struct evt_arrays *evt = new evt_arrays;
+  std::cout << "get_page_size return is " << getpagesize() << std::endl;
+
   int totEvts;
+  
+  struct evt_arrays *evt = new evt_arrays;
   struct fep_arrays *fep_dev = new fep_arrays;
   struct extra_data *edata_dev = new extra_data;
   struct fit_arrays *fit_dev = new fit_arrays;
   struct fout_arrays *fout_dev = new fout_arrays;
-
+  
+  int n_words=k;
+  
+  int N_THREADS_PER_BLOCK = 32;
+  int *ids, *out1, *out2, *out3;
+  int tEvts = 0;
+  // unsigned int *d_data_in;
+  long sizeW = sizeof(int) * n_words;
+  
+  ids  = (int *)malloc(sizeW);
+  out1 = (int *)malloc(sizeW);
+  out2 = (int *)malloc(sizeW);
+  out3 = (int *)malloc(sizeW);
+  
+  /*
+  struct evt_arrays *evt = (struct evt_arrays*)memalign(getpagesize(),sizeof(struct evt_arrays));
+  struct fep_arrays *fep_dev = (struct fep_arrays*)memalign(getpagesize(),sizeof(struct fep_arrays));
+  struct extra_data *edata_dev = (struct extra_data*)memalign(getpagesize(),sizeof(struct extra_data));
+  struct fit_arrays *fit_dev = (struct fit_arrays*)memalign(getpagesize(),sizeof(struct fit_arrays));
+  struct fout_arrays *fout_dev = (struct fout_arrays*)memalign(getpagesize(),sizeof(struct fout_arrays));
+  */
+  /*
+  struct evt_arrays *evt = (struct evt_arrays*)memalign(pow(2,ceil(log(sizeof(struct evt_arrays))/log(2))),sizeof(struct evt_arrays));
[…]

### 2014-02-14 src_opencl/svtsim_defines.h

@@ -23,7 +23,7 @@
 
 
 //GF
-#define NEVTS 200
+#define NEVTS 1500
 #define NROADS 64
 #define EVT 0 //all words from the input file are considered belonging to the same event
 
